### 2018-12-18 README.Rmd
Initial commit: Setup repository

---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```
# fortebaseline

This repository contains the source code and raw manuscript files for the Hector permafrost emissions paper.

## Installation

You can install the development version of `hector_permafrost_emit` from GitHub with:

```r
# install.packages("devtools")
devtools::install_github("ashiklom/hector_permafrost_emit")
```

### 2018-12-18 analysis/paper/paper.Rmd
Initial commit: Setup repository

---
title: "Hector Permafrost"
author:
  - Alexey N. Shiklomanov
  - Ben Bond-Lamberty
  - Corinne Hartin
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: 
    bookdown::html_document2:
      fig_caption: yes
      reference_docx: "../templates/template.docx"
bibliography: references.bib
csl: "../templates/biogeosciences.csl"
abstract: |
  Text of abstract
keywords: |
  keyword 1; keyword 2; keyword 3
highlights: |
  These are the highlights. 
---

```{r, setup, echo = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  warning = FALSE,
  message = FALSE,
  echo = FALSE,
  comment = "#>",
  fig.path = "../figures/"
)

library(hector_permafrost_emit) # Or use devtools::load_all('.', quiet = T) if your code is in script files, rather than as functions in the `/R` diretory
```

# Introduction

Permafrost carbon sink:

- [@zimov_2006_climate_change]
- [@treat_2013_permafrost_carbon_bomb]
- [@schuur_2015_climate; @schuur_2011_climate_change; @schaefer_2014_impact; @schaefer_2011_amount; @lee_2011_rate; @kuhry_2010_potential; @kessler_2017_estimating; @hope_2015_economic; @drake_2015_ancient; @burke_2017_quantifying]

# Methods

## Model description: Hector

## Permafrost emissions scenarios

Digitized scenarios from [@schaefer_2011_amount].

Back-of-the-envelope estimates from [@zimov_2006_climate_change]:
- 500 Gt C in loess that could be completely emitted by 2100 (plus other C sources).
- 10-40 g C m-3 day-1 off the bat, slowing down to equilibrium (?) rate of 0.5-5 g C m-3 day-1 for several years. Combine with data on permafrost spatial extent, density, etc. to generate estimates (but can back-calculate from 500 Gt C above?)

# Results

```{r get-data, eval = FALSE}
# Note the path that we need to use to access our data files when rendering this document
my_data <- readr::read_csv("../data/raw_data/my_csv_file.csv")
```

# Discussion

# Conclusion

# Acknowledgements

<!-- The following line inserts a page break when the output is MS Word. For page breaks in PDF, use \newpage on its own line.  -->
##### pagebreak

# References 
<!-- The following line ensures the references appear here for the MS Word or HTML output files, rather than right at the end of the document (this will not work for PDF files):  -->
<div id="refs"></div>

##### pagebreak

### Colophon

This report was generated on `r Sys.time()` using the following computational environment and dependencies:

```{r colophon, cache = FALSE}
# which R packages and versions?
devtools::session_info()
```

The current Git commit details are:

```{r}
# what commit is this file at? You may need to change the path value
# if your Rmd is not in analysis/paper/
git2r::repository("../..")
```

### 2018-12-18 analysis/scratch.R

@@ -1,0 +1,50 @@
+# Read CRUNCEP wind inputs. Is the same wind time series looped
+# from 1902 to 1948? 
+winds <- list()
+files <- glue::glue("http://localhost:7999/thredds/dodsC/dbfiles/CRUNCEP_site_1-33/CRUNCEP.{1903:1910}.nc")
+for (yr in 1903:1951) {
+  file <- glue::glue("http://localhost:7999/thredds/dodsC/dbfiles/CRUNCEP_site_1-33/CRUNCEP.{yr}.nc")
+  nc <- ncdf4::nc_open(file)
+  winds[[as.character(yr)]] <- tibble::tibble(
+    year = yr,
+    north = ncdf4::ncvar_get(nc, "northward_wind"),
+    east = ncdf4::ncvar_get(nc, "eastward_wind"),
+    rn = seq_along(north)
+  )
+}
+wind_df <- dplyr::bind_rows(winds) %>%
+  dplyr::mutate(datetime = ISOdate(year, 1, 1, 0, 0, 0, tz = "UTC") + 4 * lubridate::hours(rn))
+
+wind_df %>%
+  dplyr::filter(year > 1940) %>%
+  ggplot() +
+  aes(x = rn, y = north) +
+  geom_line() +
+  facet_grid(year ~ .)
+
+ggplot(wind_df) +
+  aes(x = rn, y = north, color = year <= 1950) +
+  geom_line()
+
+# Yes, yes it is.
+############################################################
+# Is read.output actually loading everything it can from the ED output?
+# No! 
+
+filename <- "http://localhost:7999/thredds/dodsC/outputs/PEcAn_99000000032/out/99000000030/analysis-T-1932-00-00-000000-g01.h5"
+hf <- ncdf4::nc_open(filename)
+soil_moist <- ncdf4::ncvar_get(hf, "FMEAN_SOIL_WATER_PY")
+ncdf4::nc_close(hf)
+
+tsoil_moist <- tibble::as_tibble(t(soil_moist)) %>%
+  dplyr::mutate(t = dplyr::row_number()) %>%
+  tidyr::gather(layer, value, -t) %>%
+  dplyr::mutate(layer = as.factor(layer))
+
+ggplot(tsoil_moist) +
+  aes(x = t, y = value, color = layer) +
+  geom_line() +
+  scale_color_viridis_d()
+
+tot_soil_moist <- colSums(soil_moist)
+plot(tot_soil_moist, type = "l")

### 2018-12-18 tests/testthat.R

@@ -1,0 +1,4 @@
+library(testthat)
+library(hector_permafrost_emit)
+
+test_check("hector_permafrost_emit")

### 2018-12-18 README.Rmd
Set up drake. Add Schaefer 2011 CO2 emissions

@@
 ```
+
+## Data description
+
+- `analysis/data/raw_data/schaefer_teb_rpc_(min/max/mean).csv` -- Raw, digitized versions of Figure 5c (Ensemble mean and uncertainty of permafrost C respiration) in [@schaefer_2011_amount].
+Original figure is `analysis/data/raw_data/schaefer_teb_timeseries.gif`.
+These results were digitized using WebPlotDigitizer (https://apps.automeris.io/wpd/), and interpolated to annual resolution (using R `stats::spline` function).
+    - [@schaefer_2011_amount] assume only CO2 emissions. So we add this as an extra anthropogenic CO2 flux.

### 2018-12-18 analysis/paper/paper.Rmd
Set up drake. Add Schaefer 2011 CO2 emissions

@@
 
+For all scenarios, we added the emission as a combination of fossil fuel CO2 emissions (`ffi_emissions` column in Hector) and methane emissions (`CH4_emissions` column in Hector).
+Where these were explicitly separated, we added them accordingly.
+Where they were presented only in terms of C, we did ???
+
 Digitized scenarios from [@schaefer_2011_amount].
+SiBCASA model predictions of CO2 emissions (permafrost respiration; $R_{pc}$; note -- no methane!) through 2300.
+These results were digitized using WebPlotDigitizer (https://apps.automeris.io/wpd/), and interpolated to annual resolution (using R `stats::spline` function).
+
+Digitized scenarios from [@hope_2015_economic].
+CO2 and CH4 emissions from SiBCASA model.
+
+[@schuur_2009_effect] -- Estimate 0.8 - 1.1 Pg C yr-1.
 

### 2018-12-18 analysis/drake.R

@@ -1,0 +1,56 @@
+library(drake)
+library(magrittr)
+library(ggplot2)
+pkgconfig::set_config("drake::strings_in_dots" = "literals")
+
+read_schaefer <- function(file, data_type) {
+  dat <- readr::read_csv(file, col_names = FALSE, col_types = "dd") %>%
+    dplyr::select(year = X1, r_pc = X2)
+  ymin <- ceiling(min(dat[["year"]]))
+  ymax <- floor(max(dat[["year"]]))
+  new_years <- seq(ymin, ymax)
+  dat_interp <- spline(dat[["year"]], dat[["r_pc"]], xout = new_years)
+  tibble::as_tibble(dat_interp) %>%
+    dplyr::select(year = x, !!data_type := y)
+}
+
+plan <- drake_plan(
+  schaefer_raw_min_file = file_in("analysis/data/raw_data/schaefer_teb_rpc_min.csv"),
+  schaefer_raw_max_file = file_in("analysis/data/raw_data/schaefer_teb_rpc_max.csv"),
+  schaefer_raw_mean_file = file_in("analysis/data/raw_data/schaefer_teb_rpc_mean.csv"),
+  schaefer_dat = read_schaefer(schaefer_raw_min_file, "r_pc_min") %>%
+    dplyr::full_join(read_schaefer(schaefer_raw_max_file, "r_pc_max")) %>%
+    dplyr::full_join(read_schaefer(schaefer_raw_mean_file, "r_pc_mean")) %>%
+    dplyr::arrange(year) %>%
+    dplyr::mutate(
+      r_pc_mean = dplyr::case_when(
+        year < 2025 ~ 0,
+        TRUE ~ r_pc_mean
+      ),
+      r_pc_min = dplyr::case_when(
+        year <= 2060 ~ r_pc_mean,
+        TRUE ~ r_pc_min
+      ),
+      r_pc_max = dplyr::case_when(
+        year <= 2060 ~ r_pc_mean,
+        TRUE ~ r_pc_max
+      )
+    ),
+  hector_rcp45 = readr::read_csv(
+    system.file("input/emissions/RCP45_emissions.csv", package = "hector"),
+    skip = 3
+  )
+)
+
+config <- drake_config(plan)
+make(plan)
+
+loadd(schaefer_dat)
+readd(hector_rcp45)
+
[…]

### 2018-12-18 R/create_scenario.R
Add code for creating scenarios. Create scenarios for Hope, Schaefer

@@ -1,0 +1,65 @@
+#' Create a Hector emissions scenario
+#'
+#' @param cc `data.frame` of total carbon (C) emissions, with columns
+#'   `year` and `value` (emissions in GtC/yr).
+#' @param ch4_frac Fraction of carbon emitted as methane. Length must
+#'   be either 1 or `nrow(cc)`.
+#' @param co2 `data.frame` of CO2 emissions (same structure as `cc`)
+#' @param ch4 `data.frame` of CH4 emissions (same structure as `cc`)
+#' @param base_scenario Character describing base scenario (default = "RCP45").
+#' @param base_scenario_file Path to base emissions file (default is
+#'   `hector` package RCP 4.5 emissions file).  
+#' @return `data.frame` containing Hector emissions, with additional
+#'   CO2 and CH4 emissions added to target years.
+#' @author Alexey Shiklomanov
+#' @export
+create_scenario <- function(cc = NULL, ch4_frac = NULL,
+                            co2 = NULL, ch4 = NULL,
+                            base_scenario = "RCP45",
+                            base_scenario_file = system.file(
+                              paste0("input/emissions/",
+                                     base_scenario,
+                                     "_emissions.csv"),
+                              package = "hector"
+                            )) {
+  if (!file.exists(base_scenario_file)) {
+    stop("Unable to find base scenario file: ", base_scenario_file)
+  }
+
+  if (is.null(co2) && is.null(ch4)) {
+    if (is.null(cc)) {
+      stop(
+        "Missing C, CO2, and CH4. ",
+        "You must provide either C and CH4 fraction, or both CO2 and CH4."
+      )
+    }
+    if (is.null(ch4_frac)) {
+      stop("Provided C, but missing CH4 fraction.")
+    }
+    if (!(length(ch4_frac) %in% c(1, nrow(cc)))) {
+      stop(
+        "CH4 fraction must be length 1 or `nrow(cc)`. ",
+        "Input is length ", length(ch4_frac), "."
+      )
+    }
+    co2 <- ch4 <- cc
+    co2[["value"]] <- cc[["value"]] * (1 - ch4_frac)
+    ch4[["value"]] <- cc[["value"]] * ch4_frac
+  }
+
+  if (is.null(co2)) stop("Missing CO2 flux.")
[…]

### 2018-12-18 R/read_and_interpolate.R

@@ -1,0 +1,49 @@
+#' Read raw emissions data file and interpolate to target years
+#'
+#' @param file Raw emissions data file
+#' @param start_year Target start year for interpolation. If `NULL`
+#'   (default), use smallest year in input data.
+#' @param end_year Target end year for interpolation. If `NULL`
+#'   (default), use largest year in input data.
+#' @param scale Numeric factor by which to multiply values, e.g. to
+#'   convert units (default = 1; i.e. no conversion).
+#' @param ... Additional arguments to [readr::read_csv()].
+#' @return `data.frame` containing interpolated data, with columns
+#'   `year` and `value`.
+#' @author Alexey Shiklomanov
+#' @export
+read_and_interpolate <- function(file,
+                                 start_year = NULL,
+                                 end_year = 2100,
+                                 scale = 1,
+                                 ...) {
+  stopifnot(file.exists(file))
+
+  dat <- readr::read_csv(file, col_names = FALSE, col_types = "dd", ...)
+  colnames(dat) <- c("year", "value")
+
+  ymin <- ceiling(min(dat[["year"]]))
+  if (is.null(start_year)) start_year <- ymin
+  if (start_year < ymin) {
+    stop(
+      "Start year ", start_year,
+      " less than min available year ", ymin, "."
+    )
+  }
+
+  ymax <- floor(max(dat[["year"]]))
+  if (is.null(end_year)) end_year <- ymax
+  if (end_year > ymax) {
+    stop(
+      "End year ", end_year,
+      " greater than max available year ", ymax, "."
+    )
+  }
+
+  years <- seq(start_year, end_year)
+  values <- spline(x = dat[["year"]], y = dat[["value"]] * scale, xout = years)
+
+  names(values) <- c("year", "value")
+
+  tibble::as_tibble(values)
+}

### 2018-12-18 analysis/drake.R

@@ -3,6 +3,8 @@ library(magrittr)
 library(ggplot2)
 pkgconfig::set_config("drake::strings_in_dots" = "literals")
 
+devtools::load_all(".")
+
 read_schaefer <- function(file, data_type) {
   dat <- readr::read_csv(file, col_names = FALSE, col_types = "dd") %>%
     dplyr::select(year = X1, r_pc = X2)
@@ -14,43 +16,52 @@ read_schaefer <- function(file, data_type) {
     dplyr::select(year = x, !!data_type := y)
 }
 
-plan <- drake_plan(
-  schaefer_raw_min_file = file_in("analysis/data/raw_data/schaefer_teb_rpc_min.csv"),
-  schaefer_raw_max_file = file_in("analysis/data/raw_data/schaefer_teb_rpc_max.csv"),
-  schaefer_raw_mean_file = file_in("analysis/data/raw_data/schaefer_teb_rpc_mean.csv"),
-  schaefer_dat = read_schaefer(schaefer_raw_min_file, "r_pc_min") %>%
-    dplyr::full_join(read_schaefer(schaefer_raw_max_file, "r_pc_max")) %>%
-    dplyr::full_join(read_schaefer(schaefer_raw_mean_file, "r_pc_mean")) %>%
-    dplyr::arrange(year) %>%
-    dplyr::mutate(
-      r_pc_mean = dplyr::case_when(
-        year < 2025 ~ 0,
-        TRUE ~ r_pc_mean
-      ),
-      r_pc_min = dplyr::case_when(
-        year <= 2060 ~ r_pc_mean,
-        TRUE ~ r_pc_min
-      ),
-      r_pc_max = dplyr::case_when(
-        year <= 2060 ~ r_pc_mean,
-        TRUE ~ r_pc_max
-      )
-    ),
-  hector_rcp45 = readr::read_csv(
-    system.file("input/emissions/RCP45_emissions.csv", package = "hector"),
-    skip = 3
+schaefer_template <- drake_plan(
+  schaefer_file = file_in("analysis/data/raw_data/schaefer_teb_rpc_ZZZ.csv"),
+  scahefer = read_and_interpolate(schaefer_file_ZZZ, 2050, 2100) %>%
+    # Convert cumulative emissions to annual
+    dplyr::mutate(value = c(0, diff(value))) %>%
[…]

### 2018-12-18 analysis/drake.R
Fix typos in Schaefer plans. Remove unnecessary CH4 unit conversion

@@ -18,7 +18,7 @@ read_schaefer <- function(file, data_type) {
 
 schaefer_template <- drake_plan(
   schaefer_file = file_in("analysis/data/raw_data/schaefer_teb_rpc_ZZZ.csv"),
-  scahefer = read_and_interpolate(schaefer_file_ZZZ, 2050, 2100) %>%
+  schaefer = read_and_interpolate(schaefer_file_ZZZ, 2050, 2100) %>%
     # Convert cumulative emissions to annual
     dplyr::mutate(value = c(0, diff(value))) %>%
     create_scenario(cc = ., ch4_frac = 0)
@@ -37,7 +37,7 @@ hope_template <- drake_plan(
   hope_co2_file = file_in("analysis/data/raw_data/hope_2016_MtCO2_ZZZ.csv"),
   hope = create_scenario(
     co2 = read_and_interpolate(hope_co2_file_ZZZ, 2012, 2100, scale = mtco2_gtc),
-    ch4 = read_and_interpolate(hope_ch4_file_ZZZ, 2012, 2100, scale = mtch4_gtc)
+    ch4 = read_and_interpolate(hope_ch4_file_ZZZ, 2012, 2100, scale = 1)
   )
 )
 
@@ -47,7 +47,14 @@ hope_plan <- evaluate_plan(
   values = c("lo", "mean", "hi")
 )
 
-scenarios_plan <- bind_plans(schaefer_plan, hope_plan)
+baseline_plan <- drake_plan(
+  no_permafrost_rcp45 = readr::read_csv(
+    system.file("input/emissions/RCP45_emissions.csv", package = "hector"),
+    skip = 3
+  )
+)
+
+scenarios_plan <- bind_plans(schaefer_plan, hope_plan, baseline_plan)
 combined_plan <- scenarios_plan %>%
   dplyr::filter(!grepl("_file_", target)) %>%
   gather_plan(target = "scenario_list") %>%
@@ -60,8 +67,21 @@ plan <- bind_plans(scenarios_plan, combined_plan)
 config <- drake_config(plan)
 make(plan)
 
+########################
+# An exploratory plot.
+# TODO: This will eventually be merged into drake plan above
+
 readd(all_scenarios) %>%
-  dplyr::filter(Date > 1900, Date <= 2100, !grepl("schaefer", scenario)) %>%
+  dplyr::filter(
+    Date > 1900,
+    Date <= 2100,
+    !grepl("schaefer", scenario)
+  ) %>%
[…]

### 2018-12-18 README.Rmd
Add Hope & Schaefer 2016 Nature CC scenario data
Update README

@@
 ```
-# fortebaseline
+# Hector permafrost emissions analysis
 
@@
     - [@schaefer_2011_amount] assume only CO2 emissions. So we add this as an extra anthropogenic CO2 flux.
+- `analysis/data/raw_data/hope_2016_Mt{CO2/CH4}_{lo/mean/hi}.csv` -- Raw, digitized versions of Figures 1 (permafrost CO2 emissions from SiBCASA) and 2 (same for CH4) from [@hope_2015_economic].
+Same digitization and interpolation procedure.
+Original figures are `analysis/data/raw_data/hope_2016_nclimate_{co2,ch4}.jpg`.

### 2018-12-19 README.Rmd
Replace `_` with `.` in package name to make it valid

@@
 
-You can install the development version of `hector_permafrost_emit` from GitHub with:
+You can install the development version of `hector.permafrost.emit` from GitHub with:
 
@@
 # install.packages("devtools")
-devtools::install_github("ashiklom/hector_permafrost_emit")
+devtools::install_github("ashiklom/hector.permafrost.emit")
 ```

### 2018-12-19 analysis/paper/paper.Rmd
Replace `_` with `.` in package name to make it valid

@@
 
-library(hector_permafrost_emit) # Or use devtools::load_all('.', quiet = T) if your code is in script files, rather than as functions in the `/R` diretory
+library(hector.permafrost.emit) # Or use devtools::load_all('.', quiet = T) if your code is in script files, rather than as functions in the `/R` diretory
 ```

### 2018-12-19 tests/testthat.R

@@ -1,4 +1,4 @@
 library(testthat)
-library(hector_permafrost_emit)
+library(hector.permafrost.emit)
 
-test_check("hector_permafrost_emit")
+test_check("hector.permafrost.emit")

### 2018-12-19 R/create_scenario.R
Add code for running Hector with scenarios, and processing results

@@ -6,26 +6,29 @@
 #'   be either 1 or `nrow(cc)`.
 #' @param co2 `data.frame` of CO2 emissions (same structure as `cc`)
 #' @param ch4 `data.frame` of CH4 emissions (same structure as `cc`)
+#' @param base_scenario_data `data.frame` containing baseline
+#'   scenario, onto which CO2 and CH4 emissions will be added. If
+#'   `NULL` (default), read from `base_scenario_file`
 #' @param base_scenario Character describing base scenario (default = "RCP45").
 #' @param base_scenario_file Path to base emissions file (default is
 #'   `hector` package RCP 4.5 emissions file).  
+#' @param ... Additional arguments to [readr::read_csv()] for reading
+#'   scenario file
 #' @return `data.frame` containing Hector emissions, with additional
 #'   CO2 and CH4 emissions added to target years.
 #' @author Alexey Shiklomanov
 #' @export
 create_scenario <- function(cc = NULL, ch4_frac = NULL,
                             co2 = NULL, ch4 = NULL,
+                            base_scenario_data = NULL,
                             base_scenario = "RCP45",
                             base_scenario_file = system.file(
                               paste0("input/emissions/",
                                      base_scenario,
                                      "_emissions.csv"),
                               package = "hector"
-                            )) {
-  if (!file.exists(base_scenario_file)) {
-    stop("Unable to find base scenario file: ", base_scenario_file)
-  }
-
+                            ),
+                            ...) {
   if (is.null(co2) && is.null(ch4)) {
     if (is.null(cc)) {
       stop(
@@ -50,8 +53,21 @@ create_scenario <- function(cc = NULL, ch4_frac = NULL,
   if (is.null(co2)) stop("Missing CO2 flux.")
   if (is.null(ch4)) stop("Missing CH4 flux.")
 
-  base_dat <- readr::read_csv(base_scenario_file, skip = 3)
-  new_scenario <- base_dat
+  if (is.null(base_scenario_data)) {
+    if (!file.exists(base_scenario_file)) {
+      stop("Unable to find base scenario file: ", base_scenario_file)
+    }
+    readr_args <- modifyList(list(skip = 3), list(...))
+    base_scenario_data <- do.call(readr::read_csv, readr_args)
+  }
+
+  stopifnot(
+    is.data.frame(base_scenario_data),
+    "ffi_emissions" %in% colnames(base_scenario_data),
[…]

### 2018-12-19 R/run_hector_emissions.R

@@ -1,0 +1,58 @@
+#' Run Hector for scenario, given as `data.frame`
+#'
+#' @inheritParams emissions2ini
+#' @inheritDotParams emissions2ini
+#' @return
+#' @author Alexey Shiklomanov
+#' @export
+run_hector_emissions <- function(emissions, ...) {
+  ini <- emissions2ini(emissions, ...)
+  result <- hector::runscenario(ini)
+  result
+}
+
+#' Create Hector INI file from emissions `data.frame`
+#'
+#' @param emissions `data.frame` containing Hector emissions data
+#' @param base_ini Path to base INI file. Default is Hector RCP 4.5
+#'   INI file.
+#' @param replace_string Regular expression for emissions string to
+#'   replace. Default is `"emissions/RCP45_emissions\\.csv"`.
+#' @param emissions_outfile Target output emissions CSV file. Default
+#'   = `tempfile(fileext = ".csv")`.
+#' @param ini_outfile Target output INI file. Default =
+#'   `tempfile(fileext = ".ini")`.
+#' @return Path to INI file
+#' @author Alexey Shiklomanov
+emissions2ini <- function(emissions,
+                          base_ini = system.file(
+                            "input/hector_rcp45.ini",
+                            package = "hector"
+                          ),
+                          replace_string = "emissions/RCP45_emissions\\.csv",
+                          emissions_outfile = tempfile(fileext = ".csv"),
+                          ini_outfile = tempfile(fileext = ".ini")) {
+
+  stopifnot(file.exists(base_ini))
+  ini_base <- readLines(base_ini)
+
+  readr::write_csv(emissions, emissions_outfile)
+
+  if (!any(grepl(replace_string, ini_base))) {
+    stop(
+      "Target replacement string '", replace_string, "' ",
+      "not found in INI file '", base_ini, "'."
+    )
+  }
+  ini_new <- gsub(replace_string, emissions_outfile, ini_base)
+
+  # Use full path, not relative path, for volcanic emissions
+  ini_new <- gsub(
+    "emissions/volcanic_RF\\.csv",
+    system.file("input/emissions/volcanic_RF.csv", package = "hector"),
+    ini_new
+  )
+  writeLines(ini_new, ini_outfile)
+
+  ini_outfile
+}

### 2018-12-19 analysis/drake.R

@@ -5,56 +5,71 @@ pkgconfig::set_config("drake::strings_in_dots" = "literals")
 
 devtools::load_all(".")
 
-read_schaefer <- function(file, data_type) {
-  dat <- readr::read_csv(file, col_names = FALSE, col_types = "dd") %>%
-    dplyr::select(year = X1, r_pc = X2)
-  ymin <- ceiling(min(dat[["year"]]))
-  ymax <- floor(max(dat[["year"]]))
-  new_years <- seq(ymin, ymax)
-  dat_interp <- spline(dat[["year"]], dat[["r_pc"]], xout = new_years)
-  tibble::as_tibble(dat_interp) %>%
-    dplyr::select(year = x, !!data_type := y)
-}
+rcps <- paste0("RCP", c("26", "45", "6", "85"))
+
+mtco2_gtc <- (12 / 48) * (1 / 1000)
+mtch4_gtc <- (12 / 16) * (1 / 1000)
+
+baseline_template <- drake_plan(
+  no_permafrost = readr::read_csv(
+    system.file("input/emissions/ZZZ_emissions.csv", package = "hector"),
+    skip = 3
+  )
+)
+
+baseline_plan <- evaluate_plan(
+  baseline_template,
+  wildcard = "ZZZ",
+  values = rcps
+)
 
 schaefer_template <- drake_plan(
   schaefer_file = file_in("analysis/data/raw_data/schaefer_teb_rpc_ZZZ.csv"),
   schaefer = read_and_interpolate(schaefer_file_ZZZ, 2050, 2100) %>%
     # Convert cumulative emissions to annual
     dplyr::mutate(value = c(0, diff(value))) %>%
-    create_scenario(cc = ., ch4_frac = 0)
+    create_scenario(
+      cc = .,
+      ch4_frac = 0,
+      base_scenario_data = no_permafrost_YYY
+    )
 )
 
 schaefer_plan <- evaluate_plan(
   schaefer_template,
-  wildcard = "ZZZ", values = c("min", "mean", "max")
+  rules = list(
+    ZZZ = c("min", "mean", "max"),
+    YYY = rcps
+  )
 )
-
-mtco2_gtc <- (12/48) * (1 / 1000)
-mtch4_gtc <- (12/16) * (1 / 1000)
 
 hope_template <- drake_plan(
   hope_ch4_file = file_in("analysis/data/raw_data/hope_2016_MtCH4_ZZZ.csv"),
[…]

### 2018-12-20 analysis/paper/paper.Rmd
Add some notes and references to introduction.

@@
 
-Permafrost carbon sink:
+Climate change is a problem (CITE: IPCC, NCA, etc.).
+Important to reduce carbon emissions to meet temperature targets.
+The extent of anthropogenic emissions reduction depends on what how much help (or push-back) we get from the physical and biological Earth systems.
 
+Permafrost is an important C reservoir.
+There are an estimated 500 Gt C in organic matter buried by frozen loess in the north plains of Siberia and Central Alaska [@zimov_2006_climate_change], and an additional 400 Gt C in other permafrost (CITE; see ref in Zimov), and XXX more in peat (CITE).
+This frozen C can be released into the atmosphere by several processes related to warming, including aerobic respiration in thawed soil (REF) and anaeribic respiration in thermokarst lakes (REF).
+These impacts are exacerbated by the fact that the Arctic is warming roughly 2.5 (???) times faster than the global average (REF).
+Projections of permafrost C emissions vary.
+[@schuur_2009_effect] estimate 0.8 - 1.1 Pg C yr-1.
+Back-of-the-envelope estimates from [@zimov_2006_climate_change]: 10-40 g C m-3 day-1 off the bat, slowing down to equilibrium (?) rate of 0.5-5 g C m-3 day-1 for several years.
+
+Several studies have attempted to evaluate the potential economic impacts of current and future permafrost C emissions.
+@hope_2015_economic projected permafrost C emissions using the SiBCASA (TODO: write in full) land surface model and used these projections as inputs to the PAGE09 integrated assessment model (IAM).
+They found that XXX.
+Similarly, @gonzalez-eguino_2016_significant used the DICE (TODO: write in full) model to XXX.
+
+(Paragraph on simple climate models.)
+Hector [@hartin_2015_simple].
+However, Hector does not have an explicit representation of permafrost C emissions.
+In this study, we investigate whether the additional complexity and parametric and structural uncertainty of an explicit representation of permafrost may be warranted in Hector.
+To do this, we evaluate the sensitivity of climate variables (as predicted by Hector) to several different exogenous scenarios of permafrost C emissions.
+
+Some more relevant references:
 - [@zimov_2006_climate_change]
@@
 ## Model description: Hector
+
+Hector [@hartin_2015_simple,@hartin_2016_ocean].
+Simple climate model.
 

### 2019-01-04 analysis/paper/paper.Rmd
Add more notes to paper

@@
 Permafrost is an important C reservoir.
-There are an estimated 500 Gt C in organic matter buried by frozen loess in the north plains of Siberia and Central Alaska [@zimov_2006_climate_change], and an additional 400 Gt C in other permafrost (CITE; see ref in Zimov), and XXX more in peat (CITE).
-This frozen C can be released into the atmosphere by several processes related to warming, including aerobic respiration in thawed soil (REF) and anaeribic respiration in thermokarst lakes (REF).
-These impacts are exacerbated by the fact that the Arctic is warming roughly 2.5 (???) times faster than the global average (REF).
+Total northern soil C pool is estimated at 1672 PgC, of which approximately 1466 Pg (88%) is in permafrost [@tarnocai_2009_soil].
+This frozen C can be released into the atmosphere by several processes related to warming, including aerobic respiration in thawed soil (REF) and anaeribic respiration in thermokarst lakes and wetlands [@turetsky_2002_boreal; @wickland_2006_effects].
+But emissions from wet soils may be offset by high organic matter accumulation rates [@camill_2001_changes].
+Permafrost thaw in boreal peatlands in north-central Saskatchewan increased CO2 and CH4 fluxes from soil to atmosphere by 1.6 and 30 times, respectively [@turetsky_2002_boreal].
+These impacts are exacerbated by the fact that the Arctic is warming roughly 2.5 (TODO ???) times faster than the global average (REF).
 Projections of permafrost C emissions vary.
@@
 Several studies have attempted to evaluate the potential economic impacts of current and future permafrost C emissions.
-@hope_2015_economic projected permafrost C emissions using the SiBCASA (TODO: write in full) land surface model and used these projections as inputs to the PAGE09 integrated assessment model (IAM).
+@lawrence_2012_simulation project declines in near-surface permafrost by 33% and 72% in RCP 2.6 and 8.5, respectively, and permafrost C feedback will stabilize at 2100 in RCP 2.6.
+More recently, @hope_2015_economic projected permafrost C emissions using the SiBCASA (TODO: write in full) land surface model and used these projections as inputs to the PAGE09 integrated assessment model (IAM).
 They found that XXX.
@@
 
+Rate of permafrost C release also depends on soil moisture conditions -- drier soils release C much faster ("carbon bomb") than wetter soils ("carbon fizz") [@elberling_2013_long].
+Moisture will also affect the balance of aerobic (CO2 release) vs.
+anaerobic (CH4) C release [@turetsky_2002_boreal], to the extent that unclear if anaerobic (wet) areas are C sources or sinks [@wickland_2006_effects].
+Effects of permafrost thaw on soil moisture are a complex hydrological problem -- drainage very sensitive to local (micro-)topography [@wickland_2006_effects].
+So will vegetation cover [@wickland_2006_effects].
+
 # Conclusion
 
-# Acknowledgements
+# Acknowledgments
+
+Funded by EPA grant XXX.
+Cyberinfrastructure support from Pacific Northwest National Laboratory (PNNL).
 

### 2019-01-04 R/make_ini.R
Add functions for writing GCAM configuration files

@@ -1,0 +1,34 @@
+#' Add exogenous CSV data to a Hector INI file
+#'
+#' @param exo_csv Path to exogenous emissions CSV file
+#' @param ini_in Path to baseline Hector INI file
+#' @param ini_out Target path for output INI file
+#' @return `ini_out`, invisibly
+#' @author Alexey Shiklomanov
+#' @export
+make_ini <- function(exo_csv, ini_in, ini_out) {
+  stopifnot(file.exists(exo_csv), file.exists(ini_in))
+  exo_csvn <- normalizePath(exo_csv, mustWork = TRUE)
+  input <- readLines(ini_in)
+  output <- input
+  # First, remove any lines already containing exo_emissions
+  output <- grep("exo_emissions", output, value = TRUE, invert = TRUE)
+  # Then, insert the line at the beginning of the simpleNbox section
+  nbox_start <- grep("\\[simpleNbox\\]", output)
+  stopifnot(length(nbox_start) == 1)
+  output <- append(
+    output,
+    paste0("exo_emissions = csv:", exo_csvn),
+    nbox_start
+  )
+  # Do the same thing for exo_ch4_emissions
+  ch4_start <- grep("\\[CH4\\]", output)
+  stopifnot(length(ch4_start) == 1)
+  output <- append(
+    output,
+    paste0("exo_ch4_emissions = csv:", exo_csvn),
+    ch4_start
+  )
+  writeLines(output, ini_out)
+  invisible(ini_out)
+}

### 2019-01-04 R/make_xml.R

@@ -1,0 +1,49 @@
+#' Process GCAM input XML files
+#'
+#' @param hector_ini Path to Hector INI file to use with scenario 
+#' @param hector_xml_in Path to reference Hector XML file to modify
+#' @param gcam_xml_in Path to reference GCAM XML file to modify
+#' @param hector_xml_out Path to target Hector XML file
+#' @param gcam_xml_out Path to target GCAM XML file
+#' @return Named `character(2)` containing paths to Hector and GCAM
+#'   output files, invisibly.
+#' @author Alexey Shiklomanov
+#' @export
+make_xmls <- function(hector_ini,
+                      hector_xml_in,
+                      gcam_xml_in,
+                      hector_xml_out,
+                      gcam_xml_out) {
+  stopifnot(
+    file.exists(hector_ini),
+    file.exists(hector_xml_in),
+    file.exists(gcam_xml_in)
+  )
+  hector_ini <- normalizePath(hector_ini, mustWork = TRUE)
+  hector_xml_out <- suppressWarnings(normalizePath(hector_xml_out))
+
+  xmlh <- readLines(hector_xml_in)
+  xmlh_out <- gsub(
+    "<hector-ini-file>.*</hector-ini-file>",
+    paste0("<hector-ini-file>", hector_ini, "</hector-ini-file>"),
+    xmlh
+  )
+  writeLines(xmlh_out, hector_xml_out)
+
+  xmlg <- readLines(gcam_xml_in)
+  xmlg_out <- gsub(
+    "<Value name *= *\"climate\">.*</Value>",
+    paste0("<Value name = \"climate\">", hector_xml_out, "</Value>"),
+    xmlg
+  )
+  writeLines(xmlg_out, gcam_xml_out)
+
+  invisible(c(hector = hector_xml_out, gcam = gcam_xml_out))
+}
+
+if (FALSE) {
+  hector_ini <- file.path(gcam_climate, "hector-gcam.ini")
+  hector_xml_in <- file.path(gcam_root, "input", "gcamdata", "xml", "hector.xml")
+  gcam_xml_in <- file.path(gcam_exe, "configuration_ref.xml")
+  hector_xml_out <- hector_xml_in
+}

### 2019-01-04 R/run_hector_emissions.R

@@ -8,9 +8,9 @@
 #'   `exo_ch4_emissions` (for CH4).
 #' @inheritParams hector::newcore
 #' @inheritDotParams hector::newcore
+#' @return Hector output, as `data.frame` (see [hector::fetchvars()]). 
 #' @author Alexey Shiklomanov
 #' @export
-#' @return
 run_hector_emissions <- function(rcp, exo_emissions_df = NULL,
                                  suppresslogging = TRUE,
                                  ...) {

### 2019-01-04 analysis/drake.R

@@ -2,11 +2,15 @@ library(drake)
 library(magrittr)
 library(ggplot2)
 pkgconfig::set_config("drake::strings_in_dots" = "literals")
-
 devtools::load_all(".")
 expose_imports("hector.permafrost.emit")
 
 rcps <- paste0("RCP", c("26", "45", "6", "85"))
+
+gcam_root <- getOption("gcam_root")
+stopifnot(!is.null(gcam_root), file.exists(gcam_root))
+gcam_climate <- file.path(gcam_root, "input", "climate")
+gcam_exe <- file.path(gcam_root, "exe")
 
 mtco2_gtc <- (12 / 48) * (1 / 1000)
 mtch4_gtc <- (12 / 16) * (1 / 1000)
@@ -62,10 +66,10 @@ scenarios_plan <- bind_plans(schaefer_plan, hope_plan)
 combined_plan <- scenarios_plan %>%
   dplyr::filter(!grepl("_file_", target)) %>%
   gather_plan(target = "scenario_list") %>%
-  dplyr::add_row(
-    target = "all_scenarios",
-    command = "dplyr::bind_rows(scenario_list, .id = 'scenario')"
-  )
+  bind_plans(drake_plan(
+    scenario_names = names(scenario_list),
+    all_scenarios = dplyr::bind_rows(scenario_list, .id = 'scenario')
+  ))
 
 run_template <- drake_plan(
   results = run_hector_emissions(
@@ -78,10 +82,7 @@ run_template <- drake_plan(
 run_plan <- evaluate_plan(
   run_template,
   list(
-    emissions__ = c(
-      "NULL", # Baseline -- no exogenous emissions
-      grep("file", scenarios_plan[["target"]], value = TRUE, invert = TRUE)
-    ),
+    emissions__ = c("NULL", readd(scenario_names)),
     rcp__ = c("26", "45", "60", "85")
   )
 )
@@ -100,6 +101,27 @@ results_plan <- bind_plans(
       )
   )
 )
+
+gcam_plan <- drake_plan(
+  csv = readr::write_csv(ZZZ, file_out("CLIM/ZZZ.csv")),
+  ini = make_ini(
+    file_in("CLIM/ZZZ.csv"),
+    file_in("CLIM/hector-gcam.ini"),
+    file_out("CLIM/ZZZ.ini")
+  ),
+  xml = make_xmls(
+    ini_ZZZ,
[…]

### 2019-01-04 analysis/scratch.R

@@ -13,6 +13,12 @@ mtch4_gtc <- (12 / 16) * (1 / 1000)
 
 #########################################
 
+make(bind_plans(scenarios_plan, combined_plan))
+
+readd(scenario_names)
+
+loadd(hope_lo)
+
 loadd(all_results)
 loadd(all_scenarios)
 

### 2019-01-04 R/make_xml.R
Add GCAM scenario name. Remove extraneous code.

@@ -1,10 +1,12 @@
 #' Process GCAM input XML files
 #'
-#' @param hector_ini Path to Hector INI file to use with scenario 
+#' @param hector_ini Path to Hector INI file to use with scenario
 #' @param hector_xml_in Path to reference Hector XML file to modify
 #' @param gcam_xml_in Path to reference GCAM XML file to modify
 #' @param hector_xml_out Path to target Hector XML file
 #' @param gcam_xml_out Path to target GCAM XML file
+#' @param gcam_scenario_name Desired GCAM scenario name, as character.
+#'   Default = `"hector_permafrost"`.
 #' @return Named `character(2)` containing paths to Hector and GCAM
 #'   output files, invisibly.
 #' @author Alexey Shiklomanov
@@ -13,7 +15,8 @@ make_xmls <- function(hector_ini,
                       hector_xml_in,
                       gcam_xml_in,
                       hector_xml_out,
-                      gcam_xml_out) {
+                      gcam_xml_out,
+                      gcam_scenario_name = "hector_permafrost") {
   stopifnot(
     file.exists(hector_ini),
     file.exists(hector_xml_in),
@@ -36,14 +39,12 @@ make_xmls <- function(hector_ini,
     paste0("<Value name = \"climate\">", hector_xml_out, "</Value>"),
     xmlg
   )
+  xmlg_out <- gsub(
+    "<Value name *= *\"scenarioName\">.*</Value>",
+    paste0("<Value name = \"scenarioName\">", gcam_scenario_name, "</Value>"),
+    xmlg_out
+  )
   writeLines(xmlg_out, gcam_xml_out)
 
   invisible(c(hector = hector_xml_out, gcam = gcam_xml_out))
-}
-
-if (FALSE) {
-  hector_ini <- file.path(gcam_climate, "hector-gcam.ini")
-  hector_xml_in <- file.path(gcam_root, "input", "gcamdata", "xml", "hector.xml")
-  gcam_xml_in <- file.path(gcam_exe, "configuration_ref.xml")
-  hector_xml_out <- hector_xml_in
 }

### 2019-01-04 analysis/drake.R

@@ -114,7 +114,8 @@ gcam_plan <- drake_plan(
     file_in("GCAM/input/gcamdata/xml/hector.xml"),
     file_in("EXE/configuration_ref.xml"),
     file_out("GCAM/input/gcamdata/xml/hector_ZZZ.xml"),
-    file_out("EXE/config_ZZZ.xml")
+    file_out("EXE/config_ZZZ.xml"),
+    gcam_scenario_name = "ZZZ"
   )
 ) %>%
   evaluate_plan(rules = list(ZZZ = readd(scenario_names))) %>%

### 2019-01-10 analysis/drake.R
Add preliminary GCAM output, and clean up climate figure

@@ -94,7 +94,7 @@ scenario_names <- c(
 run_plan <- evaluate_plan(
   run_template,
   list(
-    emissions__ = c("NULL", scenario_names),
+    emissions__ = scenario_names,
     rcp__ = c("26", "45", "60", "85")
   )
 )
@@ -107,8 +107,7 @@ results_plan <- bind_plans(
       tibble::as_tibble() %>%
       tidyr::separate(scenario, c("permafrost", "RCP"), sep = "\\.") %>%
       dplyr::mutate(
-        permafrost = forcats::fct_inorder(permafrost) %>%
-          forcats::fct_recode("baseline" = "NULL"),
+        permafrost = forcats::fct_inorder(permafrost),
         RCP = forcats::fct_inorder(RCP)
       )
   )
@@ -137,48 +136,78 @@ gcam_plan <- drake_plan(
   )
 
 plot_plan <- drake_plan(
-  tidy_scenarios = all_scenarios %>%
-    tidyr::gather(variable, value, -Date, -scenario) %>%
-    dplyr::rename(year = Date) %>%
-    dplyr::mutate(datatype = "emissions"),
-  tidy_results = all_results %>%
-    tibble::as_tibble() %>%
-    dplyr::select(-scenario) %>%
-    dplyr::rename(scenario = id) %>%
+  gcam_project = rgcam::loadProject(file_in("gcam-output/permafrost.dat")),
+  gcam_results = c(Ca = "CO2 concentrations",
+                   Ftot = "Climate forcing",
+                   Tgav = "Global mean temperature") %>%
+    purrr::imap_dfr(function(x, n) purrr::map_dfr(gcam_project, x) %>%
+                                     dplyr::mutate(variable = n, RCP = "GCAM")) %>%
+    dplyr::rename(permafrost = scenario, units = Units),
+  combined_results = suppressWarnings(dplyr::bind_rows(all_results, gcam_results)),
+  climate_plot = combined_results %>%
+    dplyr::filter(year > 2000, year <= 2100,
+                  !grepl("schaefer", permafrost),
+                  variable == c("Tgav")) %>%
     dplyr::mutate(
-      scenario = gsub("results_", "", scenario),
-      datatype = "results"
-    ),
-  tidy_all = dplyr::bind_rows(tidy_scenarios, tidy_results) %>%
-    dplyr::mutate(
-      RCP = gsub(".*(RCP.*)", "\\1", scenario) %>%
-        forcats::fct_inorder(),
[…]

### 2019-01-10 analysis/read-gcam-output.R

@@ -1,0 +1,68 @@
+library(ggplot2)
+
+import::from(magrittr, "%>%")
+
+gcam_db_path <- "~/Projects/hector_project/permafrost_emit/gcam-output/"
+
+conn <- rgcam::localDBConn(gcam_db_path, "database_basexdb")
+project_file <- file.path(gcam_db_path, "permafrost.dat")
+
+scenarios <- c("no_permafrost", paste0("hope_", c("lo", "mean", "hi"))) %>%
+  purrr::walk(~rgcam::addScenario(conn, project_file, scenario = .x))
+
+prj <- rgcam::loadProject(project_file)
+
+getvar <- function(x, var) {
+  dplyr::mutate(x[[var]], variable = !!var)
+}
+
+climate <- c("CO2 concentrations", "Climate forcing", "Global mean temperature") %>%
+  purrr::map_dfr(~purrr::map_dfr(prj, getvar, var = .x, .id = "variable")) %>%
+  dplyr::mutate(
+    scenario = factor(scenario, scenarios),
+    variable = forcats::fct_inorder(variable)
+  ) %>%
+  dplyr::filter(year > 1975)
+
+climate %>%
+  dplyr::filter(year == max(year)) %>%
+  tidyr::spread(scenario, value) %>%
+  dplyr::mutate(diff = hope_hi - hope_lo) %>%
+  dplyr::select(variable, diff)
+
+ggplot(climate) +
+  aes(x = year, y = value, linetype = scenario) +
+  geom_line() +
+  facet_wrap(~variable, scales = "free_y")
+
+gdp <- prj %>%
+  purrr::map_dfr("GDP by region") %>%
+  dplyr::mutate(
+    scenario = factor(scenarios),
+    region = forcats::fct_inorder(region)
+  )
+
+gdp %>%
+  dplyr::filter(scenario %in% c("hope_lo", "hope_hi")) %>%
+  tidyr::spread(scenario, value) %>%
+  dplyr::mutate(diff = hope_hi - hope_lo)
+
+ggplot(dat) +
+  aes(x = year, y = value, linetype = scenario) +
+  geom_line()
+
+drake::readd(all_scenarios) %>%
+  dplyr::group_by(scenario) %>%
+    dplyr::summarize_at(
+      dplyr::vars(exo_emissions, exo_ch4_emissions),
+      sum
+    )
+
+drake::readd(all_results) %>%
+  dplyr::filter()
+
[…]

### 2019-01-10 analysis/scratch.R

@@ -32,3 +32,18 @@ all_results %>%
   geom_line() +
   facet_wrap(~variable, scales = "free_y") +
   scale_linetype(drop = TRUE)
+
+#########################################
+no_permafrost <- tibble::tibble(
+  Date = 2000:2100,
+  exo_emissions = 0,
+  exo_ch4_emissions = 0
+)
+
+run_hector_emissions("45", no_permafrost)
+
+loadd(all_results)
+loadd(gcam_results)
+## loadd(tidy_scenarios)
+readd(combined_results) %>%
+  dplyr::count(variable)

### 2019-01-10 analysis/paper/paper.Rmd
PAPER: Add more notes and references

@@
 Back-of-the-envelope estimates from [@zimov_2006_climate_change]: 10-40 g C m-3 day-1 off the bat, slowing down to equilibrium (?) rate of 0.5-5 g C m-3 day-1 for several years.
+
+(Previous work on modeling permafrost)
+[@burke_2017_quantifying] -- JULES and ORCHIDEE, with new permafrost scheme, combined with intermediate complexity climate-ocean model (IMOGEN), to look at climate sensitivity to permafrost C emissions.
 
@@
 
+Temperature amplification of permafrost carbon feedback (by 2100) 0.02 to 0.36 °C [@burke_2013_estimating; @schneider_2012_estimating; @schneider_2015_observation], or 0.1 to 0.8 °C in [@macdougall_2012_significant; @macdougall_2013_if], or 10-40% of peak temperature change [@crichton_2016_permafrost], or 0.2 to 12% [@burke_2017_quantifying].
+
+Permafrost carbon has greater impact at lower emissions scenarios [@burke_2017_quantifying; @macdougall_2012_significant; @macdougall_2013_if] .
+
 # Conclusion

### 2019-01-17 analysis/paper/paper.Rmd
Add more notes and references to paper

@@
   - Ben Bond-Lamberty
+  - Ben Kravitz?
+  - Kate Calvin?
   - Corinne Hartin
@@
 Permafrost thaw in boreal peatlands in north-central Saskatchewan increased CO2 and CH4 fluxes from soil to atmosphere by 1.6 and 30 times, respectively [@turetsky_2002_boreal].
-These impacts are exacerbated by the fact that the Arctic is warming roughly 2.5 (TODO ???) times faster than the global average (REF).
+These impacts are exacerbated by the fact that the Arctic is warming roughly 2.5 (TODO ???) times faster than the global average (TODO REF).
 Projections of permafrost C emissions vary.
@@
 
+On the other hand, warming and CO2 fertilization may increase vegetation productivity, which could increase soil C storage through enhanced litterfall; the balance of these two processes is uncertain [@jones_2005_global].
+Several modeling studies generally predict increases in soil C in high latitudes [@burke_2017_quantifying; @qian_2010_enhanced; @ito_2016_impacts].
+However, this increase is dampened by including permafrost C [@burke_2017_quantifying].
+
 (Previous work on modeling permafrost)
 [@burke_2017_quantifying] -- JULES and ORCHIDEE, with new permafrost scheme, combined with intermediate complexity climate-ocean model (IMOGEN), to look at climate sensitivity to permafrost C emissions.
+Models are highly sensitive to representation of soil processes, which can be more important than differences in scenario and/or climate drivers [@burke_2017_quantifying].
 
@@
 
+@hope_2015_economic (and others?) rely on climate damage functions of the general form $\delta GDP = f(\delta T)$.
+Such functions are, at best, highly uncertain (TODO REF), and possibly conceptually flawed (TODO REF).
+(TODO More on this).
+GCAM (TODO full name) is an alternative (TODO REF).
+(TODO More on GCAM).
+
 (Paragraph on simple climate models.)
@@
 To do this, we evaluate the sensitivity of climate variables (as predicted by Hector) to several different exogenous scenarios of permafrost C emissions.
+(TODO Modify this to be about economic impact)
 
@@
 to generate estimates (but can back-calculate from 500 Gt C above?)
+
+## Modeling permafrost emissions
+
+Frozen carbon residence time (FCRt) from [@burke_2017_quantifying]:
+
+$$ FCRt = FCRt0 * exp(-ΔT / Γ) (for ΔT > 0.2 °C) $$
+
+- $FCRt_0$ (years) reflects the stability of permafrost C (length of time that permafrost C is stable when $\Delta T = 0$)
+- $\Gamma$ -- decay term (°C); temperature change at which "the number of years taken for all of the old permafrost C to be emitted reduces to 1/e of its initial value"
 

### 2019-02-06 analysis/paper/paper.Rmd
Add permafrost emissions cases to paper outline

@@
 
+Default heterotrophic respiration ($R$) scheme for a pool $p$ (detritus or soil) in Hector:
+
+$$ R_{p} = C_p \times f_p \times Q_{10} ^ \frac{T}{10} $$
+
+This is case 1.
+
+Case 2 uses the same formulation, but splits permafrost C into its own "biome", with its own photosynthesis ($\Beta$) and respiration ($Q_{10}$) parameters.
+
+Case 3 extends Case 2 by calculating CO~2~ and CH~4~ emissions as a fraction of the total respiration, controlled by two new parameters $\alpha$ and $\phi$:
+
+$$ R_{p,tot} = C_p \times f_p \times Q_{10} ^ \frac{T}{10} $$
+
+$$ R_{p, CH_4} = \alpha R_{p, tot} $$
+
+$$ R_{p, CO_2} = (1 - \alpha) ^ \phi R_{p, tot} $$
+
+## Other notes
+
 Frozen carbon residence time (FCRt) from [@burke_2017_quantifying]:

### 2019-02-06 analysis/paper/paper.Rmd
Add package citations to getcitations script

@@
       reference_docx: "../templates/template.docx"
-bibliography: references.bib
+bibliography: ["references.bib", "packages.bib"]
 csl: "../templates/biogeosciences.csl"

### 2019-02-06 analysis/paper/paper.Rmd
Add lots of methods notes to paper

@@
 
+## Parameter calibration
+
+We used the `BayesianTools` R package [@R-BayesianTools] for all parameter calibration.
+The outputs of these calibrations are joint posterior distributions of parameters and their covariances, from which we sample for the sensitivity analysis.
+
+For global parameters, we used the following likelihood:
+
+$$ \log(L) = \sum_s Normal(Hector(\Beta, Q_{10}, s) | CMIP5(s), \sigma) $$
+
+where $s$ is one of the four representative carbon pathways (RCPs), $CMIP5(s)$ are the CMIP5 global mean outputs for the corresponding variables, and $\sigma$ is the model error (estimated during the fit).
+We also used the resulting distributions for $\Beta$ and $Q_{10}$ for the non-permafrost biome in cases 2 and 3.
+We feel this is appropriate because the CMIP5 models against which these parameters are calibrated do not include permafrost C feedbacks.
+
+For case 2, we calibrated the permafrost-specific $\Beta$ and $Q_{10}$ against various literature sources, including:
+
+- Land surface model simulations [@hope_2015_economic; @schaefer_2011_amount; @burke_2017_quantifying].
+    - Try to calibrate against NPP and soil respiration if possible
+- Literature surveys [@schaefer_2014_impact] 
+- Warming experiments [@wickland_2006_effects]
+- TODO: Others?
+
+Some of these are time series, while others are individual estimates at particular points in time.
+To give them equal weight in the likelihood, we down-weight the time series likelihoods by the number of time steps.
+
+We derived a distribution for the Arctic warming factor from TODO.
+
+TODO: Table and multi-panel figure of input datasets.
+
+For the $\alpha$ and $\phi$ parameters in case 3, we looked at the literature on permafrost methane emissions [e.g., @wickland_2006_effects].
+
+## Initial conditions
+
+For case 1, we use the Hector defaults [@hartin_2015_simple].
+
+For cases 2 and 3, we assumed the same total pool sizes as case 1, but divided them into the permafrost and non-permafrost biomes according to TODO.
+
 ## Other notes
@@
 
+Other Hector parameters to consider.
+
+| Variable | INI name | Description                                 | Value |
+|----------|----------|---------------------------------------------|-------|
+| $f_{nv}$ | `f_nppv` | Fraction of NPP C transferred to vegetation | 0.35  |
+| $f_{nd}$ | `f_nppd` | Fraction of NPP C transferred to detritus   | 0.60  |
+| $f_{nd}$ |          | Fraction of NPP C transferred to soil       | 0.05  |
+| $f_{lv}$ | `f_lucv` | Fraction of LUC change flux from vegetation | 0.10  |
+| $f_{ld}$ | `f_lucd` | Fraction of LUC change flux from detritus   | 0.01  |
+| $f_{ls}$ |          | Fraction of LUC change flux from soil       | 0.89  |
+| $f_{ds}$ |          | Fraction of detritus C that goes to soil    | 0.60  |
+| $f_{rd}$ |          | Fraction of respiration C to detritus       | 0.25  |
+| $f_{rs}$ |          | Fraction of respriation C to soil           | 0.02  |
+
+According to [@hartin_2015_simple], these were selected to be "generally consistent with previous simple earth system models [e.g., @meinshausen_2011_emulating]".
+
 # Results
+
+**Figure**: Parameter distributions from calibration.
+
+**Figure**: Hector projections of parameter sensitivity (CI ribbon, or light/transparent lines), colored (faceted?) by case.
+
+**Figure**: PEcAn-like variance decomposition of parameters (for each parameter: sensitivity, uncertainty, and partial variance)
 

### 2019-02-06 analysis/cmip5-calibrate.R
Very rough CMIP5 calibration script

@@ -1,0 +1,111 @@
+library(hector)
+library(hectortools)
+library(tidyverse)
+library(BayesianTools)
+
+GMT_REF <- 13 # Hector reference temperature, degrees C
+
+hector_dir <- "~/Projects/hector_project/hector"
+cmip5_dir <- file.path(
+  hector_dir,
+  "scripts",
+  "outputplotter",
+  "comparison_data"
+)
+cmip5_files <- list.files(cmip5_dir, recursive = TRUE, full.names = TRUE)
+
+model_file <- file.path(cmip5_dir, "GMD_2014", "IPCC", "IPCC2005.csv")
+model_file <- file.path(cmip5_dir, "GMD_2015", "CMIP5",
+                        "Hector_models_tas", "rcp45", "ensemble_mean_model.csv")
+
+model_data <- read_csv(model_file)
+
+hc <- newcore(system.file("input", "hector_rcp45.ini", package = "hector"),
+              name = "hector",
+              suppresslogging = TRUE)
+
+run(hc)
+result <- fetchvars(hc, 2000:2100, "Tgav")
+
+model_data %>%
+  filter(
+    year <= 2100,
+    vtag == "tgav",
+    value > 13
+  ) %>%
+  ggplot() +
+  aes(x = year, y = value - min(value), color = model) +
+  geom_line() +
+  geom_line(aes(color = "hector"), data = result, linetype = "dashed")
+
+model_wide <- model_data %>%
+  filter(year >= 2006, year <= 2100, vtag == "tgav") %>%
+  select(model, year, value) %>%
+  spread(model, value)
+
+model_tgav <- model_wide %>%
+  select(-year) %>%
+  as.matrix() %>%
+  `-`(GMT_REF)
+
+# Define test likelihood
+run_hector_params <- function(params, pb = NULL) {
+  if (!is.null(pb)) pb$tick()
+  beta <- params[1]
+  q10 <- params[2]
+  reset(hc)
+  setvar(hc, NA, BETA(), beta, NA)
+  setvar(hc, NA, Q10_RH(), q10, NA)
+  run(hc, 2100)
+  result <- fetchvars(hc, 2006:2100, "Tgav")
+  result
+}
+
+hc <- newcore(system.file("input", "hector_rcp45.ini", package = "hector"),
+              name = "hector",
+              suppresslogging = TRUE)
+
+matplot(model_tgav, type = "l")
[…]

### 2019-05-30 analysis/drake.R
Store parameters in CSV file

@@ -1,39 +1,22 @@
 #!/usr/bin/env Rscript
-library(drake)
-library(ggplot2)
+library(drake, exclude = c("expand", "gather", "plan"))
+library(ggplot2, exclude = "ggsave")
+library(dplyr)
+library(tidyr)
+library(readr)
+library(parallel, include.only = "detectCores")
+library(cowplot)
+library(future)
 
-requireNamespace("future", quietly = TRUE)
-
-# begin imports
-import::from("tibble", "tibble", "as_tibble", .into = "")
-import::from("dplyr", "bind_rows", "bind_cols", "filter", "group_by",
-             "summarize", "mutate", "ungroup", .into = "")
-import::from("cowplot", "plot_grid", "theme_cowplot", "save_plot",
-             .into = "")
-import::from("parallel", "detectCores", .into = "")
-import::from("magrittr", "%>%", .into = "")
-import::from("tidyr", "gather", .into = "")
-# end imports
+requireNamespace("ggpairs", quietly = TRUE)
 
 devtools::load_all(".")
 expose_imports("hector.permafrost.emit")
 
 # To make this reproducible
-set.seed(8675309)
-ndraws <- 500
-beta_draws <- runif(ndraws, 0, 1)
-q10_draws <- runif(ndraws, 0, 10)
-
-npp_alpha <- c(f_nppv = 0.35, f_nppd = 0.60, f_npps = 0.05)
-npp_draws <- rdirichlet(ndraws, npp_alpha)[, 1:2]
-
-draws <- tibble(
-  beta = runif(ndraws, 0, 1),
-  q10_rh = runif(ndraws, 0, 10),
-  f_litterd = rbeta(ndraws, 0.98 * 4, 0.02 * 4)
-) %>%
-  bind_cols(as_tibble(npp_draws))
-
+draws <- read_csv(file.path("analysis", "data",
+                            "derived_data", "parameter-draws.csv"))
+draws <- head(draws, 100)
 params <- rlang::syms(colnames(draws))
 
 plan <- drake_plan(
@@ -117,12 +100,10 @@ plan <- drake_plan(
   )
 )
 
-is_callr <- Sys.getenv("CALLR") == "true"
-
 dconf <- drake_config(
   plan,
   parallelism = "future",
-  jobs = 1,
+  jobs = parallel::detectCores(),
   prework = quote({
     devtools::load_all(".", quiet = TRUE)
   })
@@ -132,15 +113,13 @@ dconf <- drake_config(
 dout <- outdated(dconf)
[…]

### 2019-05-30 analysis/scripts/draw_params.R

@@ -1,0 +1,21 @@
+library(dplyr)
+library(readr)
+library(hector.permafrost.emit)
+
+set.seed(8675309)
+ndraws <- 1000
+beta_draws <- runif(ndraws, 0, 1)
+q10_draws <- runif(ndraws, 0, 10)
+
+npp_alpha <- c(f_nppv = 0.35, f_nppd = 0.60, f_npps = 0.05)
+npp_draws <- rdirichlet(ndraws, npp_alpha)[, 1:2]
+
+draws <- tibble(
+  beta = runif(ndraws, 0, 1),
+  q10_rh = runif(ndraws, 0, 10),
+  f_litterd = rbeta(ndraws, 0.98 * 4, 0.02 * 4)
+) %>%
+  bind_cols(as_tibble(npp_draws))
+
+write_csv(draws, file.path("analysis", "data",
+                           "derived_data", "parameter-draws.csv"))

### 2019-05-30 analysis/drake.R
Add flag to control if all draws are run

@@ -16,7 +16,9 @@ expose_imports("hector.permafrost.emit")
 # To make this reproducible
 draws <- read_csv(file.path("analysis", "data",
                             "derived_data", "parameter-draws.csv"))
-draws <- head(draws, 100)
+if (!("--all" %in% commandArgs(trailingOnly = TRUE))) {
+  draws <- head(draws, 100)
+}
 params <- rlang::syms(colnames(draws))
 
 plan <- drake_plan(
@@ -100,26 +102,27 @@ plan <- drake_plan(
   )
 )
 
-dconf <- drake_config(
-  plan,
-  parallelism = "future",
-  jobs = parallel::detectCores(),
-  prework = quote({
-    devtools::load_all(".", quiet = TRUE)
-  })
-)
-
-# Set number of cores depending on number of outdated tasks
-dout <- outdated(dconf)
-if (length(dout) > 10) {
-  dconf[["jobs"]] <- parallel::detectCores()
-  message("Number of outdated targets (", length(dout), ") ",
-          "is greater than 10.",
-          "Running in parallel across ", dconf[["jobs"]], " cores.")
-}
-
 if (interactive()) {
+  ## callr::rscript(here::here("analysis", "drake.R"), cmdargs = "--all")
   callr::rscript(here::here("analysis", "drake.R"))
 } else {
+  dconf <- drake_config(
+    plan,
+    parallelism = "future",
+    jobs = parallel::detectCores(),
+    prework = quote({
+      devtools::load_all(".", quiet = TRUE)
+    })
+  )
+
+  # Set number of cores depending on number of outdated tasks
+  dout <- outdated(dconf)
+  if (length(dout) > 10) {
+    dconf[["jobs"]] <- parallel::detectCores()
+    message("Number of outdated targets (", length(dout), ") ",
+            "is greater than 10.",
+            "Running in parallel across ", dconf[["jobs"]], " cores.")
+  }
+
   make(config = dconf)
 }

### 2019-05-30 analysis/scripts/draw_params.R
Add Makefile

@@ -1,3 +1,4 @@
+#!/usr/bin/env Rscript
 library(dplyr)
 library(readr)
 library(hector.permafrost.emit)

### 2019-05-30 analysis/paper/paper.Rmd
Add more notes and refs to paper
Render manuscript (and fix some LaTeX issues)

@@
 They found that XXX.
+@chen_2019_economic also used PAGE09.
 Similarly, @gonzalez-eguino_2016_significant used the DICE (TODO: write in full) model to XXX.
+@kessler_2017_estimating also used DICE.
 
@@
 
-Case 2 uses the same formulation, but splits permafrost C into its own "biome", with its own photosynthesis ($\Beta$) and respiration ($Q_{10}$) parameters.
+Case 2 uses the same formulation, but splits permafrost C into its own "biome", with its own photosynthesis ($\beta$) and respiration ($Q_{10}$) parameters.
 
@@
 
-$$ \log(L) = \sum_s Normal(Hector(\Beta, Q_{10}, s) | CMIP5(s), \sigma) $$
+$$ \log(L) = \sum_s Normal(Hector(\beta, Q_{10}, s) | CMIP5(s), \sigma) $$
 
 where $s$ is one of the four representative carbon pathways (RCPs), $CMIP5(s)$ are the CMIP5 global mean outputs for the corresponding variables, and $\sigma$ is the model error (estimated during the fit).
-We also used the resulting distributions for $\Beta$ and $Q_{10}$ for the non-permafrost biome in cases 2 and 3.
+We also used the resulting distributions for $\beta$ and $Q_{10}$ for the non-permafrost biome in cases 2 and 3.
 We feel this is appropriate because the CMIP5 models against which these parameters are calibrated do not include permafrost C feedbacks.
 
-For case 2, we calibrated the permafrost-specific $\Beta$ and $Q_{10}$ against various literature sources, including:
+For case 2, we calibrated the permafrost-specific $\beta$ and $Q_{10}$ against various literature sources, including:
 
@@
 
-$$ FCRt = FCRt0 * exp(-ΔT / Γ) (for ΔT > 0.2 °C) $$
+$$ FCRt = FCRt0 * exp(-\Delta T / \Gamma) (for \Delta T > 0.2 °C) $$
 
@@
 
+```{r fig-parameter-distribution, fig.cap = param_dist_cap}
+param_dist_cap <- paste(
+  "Input parameter distributions."
+)
+readd(draws_plot)
+```
+
 # Results
-
-**Figure**: Parameter distributions from calibration.
 
@@
 **Figure**: PEcAn-like variance decomposition of parameters (for each parameter: sensitivity, uncertainty, and partial variance)
-
-```{r get-data, eval = FALSE}
-# Note the path that we need to use to access our data files when rendering this document
-my_data <- readr::read_csv("../data/raw_data/my_csv_file.csv")
-```
 

### 2019-05-30 analysis/drake.R

@@ -7,6 +7,7 @@ library(readr)
 library(parallel, include.only = "detectCores")
 library(cowplot)
 library(future)
+library(here)
 
 requireNamespace("ggpairs", quietly = TRUE)
 
@@ -22,6 +23,9 @@ if (!("--all" %in% commandArgs(trailingOnly = TRUE))) {
 params <- rlang::syms(colnames(draws))
 
 plan <- drake_plan(
+  paper_file = knitr_in(!!here("analysis", "paper", "paper.Rmd")),
+  paper_md = rmarkdown::render(paper_file, "github_document"),
+  paper_pdf = rmarkdown::render(paper_file, "pdf_document"),
   draws_plot = GGally::ggpairs(draws),
   sims = target(
     hector_with_params(

### 2019-05-30 analysis/drake.R
Set up paper deployment on gh-pages

@@ -22,10 +22,11 @@ if (!("--all" %in% commandArgs(trailingOnly = TRUE))) {
 }
 params <- rlang::syms(colnames(draws))
 
+paper_file <- here("analysis", "paper", "paper.Rmd")
+
 plan <- drake_plan(
-  paper_file = knitr_in(!!here("analysis", "paper", "paper.Rmd")),
-  paper_md = rmarkdown::render(paper_file, "github_document"),
-  paper_pdf = rmarkdown::render(paper_file, "pdf_document"),
+  paper_md = rmarkdown::render(knitr_in(!!paper_file), "github_document"),
+  paper_pdf = rmarkdown::render(knitr_in(!!paper_file), "pdf_document"),
   draws_plot = GGally::ggpairs(draws),
   sims = target(
     hector_with_params(

### 2019-05-30 analysis/scripts/deploy-paper.sh

@@ -1,0 +1,32 @@
+#!/usr/bin/env bash
+set -o errexit
+
+PROJDIR=$(git rev-parse --show-toplevel)
+cd "$PROJDIR"
+
+DEPLOY_DIR=analysis/paper/_deploy
+mkdir -p "$DEPLOY_DIR"
+cp -Rf analysis/paper/paper{.md,.html,.pdf,_files} "$DEPLOY_DIR"
+
+cd "$DEPLOY_DIR"
+
+if [[ ! -f index.html ]]; then
+    ln -sf paper.html index.html
+fi
+
+BRANCH=$(git rev-parse --abbrev-ref HEAD)
+if [[ "$BRANCH" != "gh-pages" ]]; then
+    echo "Creating the gh-pages branch."
+    git init
+    git checkout -b gh-pages
+    git remote add origin git@github.com:ashiklom/hector_permafrost_emit
+fi
+
+if [[ -n $(git status --short) ]]; then
+    git add .
+    MSG=`date "+Update paper <%Y-%m-%d %H:%M>"`
+    git commit -a -m "$MSG"
+    git push --force origin gh-pages
+else
+    echo "Nothing to commit."
+fi

### 2019-05-30 analysis/paper/paper.Rmd
Set up paper deployment on gh-pages
Add true HTML export (including MathJax)

@@
       reference_docx: "../templates/template.docx"
+    rmarkdown::github_document:
+      html_preview: false
 bibliography: ["references.bib", "packages.bib"]
@@
   echo = FALSE,
-  comment = "#>",
-  fig.path = "../figures/"
+  comment = "#>"
 )

### 2019-05-30 analysis/drake.R

@@ -27,6 +27,7 @@ paper_file <- here("analysis", "paper", "paper.Rmd")
 plan <- drake_plan(
   paper_md = rmarkdown::render(knitr_in(!!paper_file), "github_document"),
   paper_pdf = rmarkdown::render(knitr_in(!!paper_file), "pdf_document"),
+  paper_html = rmarkdown::render(knitr_in(!!paper_file), "html_document"),
   draws_plot = GGally::ggpairs(draws),
   sims = target(
     hector_with_params(

### 2019-05-30 analysis/scripts/deploy-paper.sh
Use git hash in paper update message

@@ -8,6 +8,8 @@ DEPLOY_DIR=analysis/paper/_deploy
 mkdir -p "$DEPLOY_DIR"
 cp -Rf analysis/paper/paper{.md,.html,.pdf,_files} "$DEPLOY_DIR"
 
+HASH=$(git rev-parse --short HEAD)
+
 cd "$DEPLOY_DIR"
 
 if [[ ! -f index.html ]]; then
@@ -24,7 +26,7 @@ fi
 
 if [[ -n $(git status --short) ]]; then
     git add .
-    MSG=`date "+Update paper <%Y-%m-%d %H:%M>"`
+    MSG="Update paper <$HASH>"
     git commit -a -m "$MSG"
     git push --force origin gh-pages
 else

### 2019-05-30 analysis/paper/paper.Rmd
Add current figures to paper

@@
 param_dist_cap <- paste(
-  "Input parameter distributions."
+  "Input parameter distributions for global Hector."
 )
@@
 # Results
+
+## Global Hector
+
+```{r fig-global-timeseries, fig.cap = global_ts_cap, fig.dim = c(10, 8)}
+global_ts_cap <- paste(
+  "Time series of Hector outputs for different parameter combinations."
+)
+readd(ts_both)
+```
+
+```{r fig-global-scatter, fig.cap = global_scatter_cap, fig.dim = c(8, 5)}
+global_scatter_cap <- paste(
+  "Hector outputs at 2100 as a function of input parameter values."
+)
+readd(scatter_both)
+```
+
+```{r fig-global-sensitivity, fig.cap = global_sensitivity_cap, fig.dim = c(5, 5)}
+global_sensitivity_cap <- paste(
+  "Sensitivity and variance decomposition analysis for global Hector."
+)
+readd(sensitivity_plot)
+```
+
+## Permafrost as a biome
+
+**Figure**: Hector projections of parameter sensitivity (CI ribbon, or light/transparent lines), colored (faceted?) by case.
+
+**Figure**: PEcAn-like variance decomposition of parameters (for each parameter: sensitivity, uncertainty, and partial variance)
+
+## Permafrost methane emissions
 

### 2019-05-30 analysis/drake.R
Refactor drake plan into separate files

@@ -24,89 +24,10 @@ params <- rlang::syms(colnames(draws))
 
 paper_file <- here("analysis", "paper", "paper.Rmd")
 
-plan <- drake_plan(
-  paper_md = rmarkdown::render(knitr_in(!!paper_file), "github_document"),
-  paper_pdf = rmarkdown::render(knitr_in(!!paper_file), "pdf_document"),
-  paper_html = rmarkdown::render(knitr_in(!!paper_file), "html_document"),
-  draws_plot = GGally::ggpairs(draws),
-  sims = target(
-    hector_with_params(
-      beta = beta,
-      q10_rh = q10_rh,
-      f_nppv = f_nppv,
-      f_nppd = f_nppd,
-      f_litterd = f_litterd
-    ),
-    transform = map(!!!draws)
-  ),
-  all_sims = target(
-    as_tibble(bind_rows(sims, .id = "id")),
-    transform = combine(sims)
-  ),
-  ts_common = ggplot(all_sims) +
-    aes(x = year, y = value, group = id) +
-    geom_line() +
-    facet_wrap(vars(variable), scales = "free_y") +
-    scale_color_gradient(low = "grey", high = "red") +
-    ylab("") +
-    theme_bw() +
-    theme(
-      legend.position = "bottom"
-    ),
-  ts = target(
-    ts_common + aes(color = .y),
-    transform = map(.y = !!params)
-  ),
-  ts_both = target(
-    plot_grid(ts),
-    transform = combine(ts)
-  ),
-  ts_both_png = save_plot(
-    file_out(!!here::here("analysis", "figures", "ts_both.png")),
-    ts_both
-  ),
-  # Scatter plot of values in 2100
-  lastyear = filter(all_sims, year == 2100),
-  sensitivity = lastyear %>%
-    group_by(variable) %>%
-    sensitivity_analysis(!!!params) %>%
-    tidy_sensitivity(),
-  sensitivity_plot = sensitivity %>%
-    filter(
-      parameter != "total",
-      stat != "sens",
-      stat != "var"
-    ) %>%
-    ggplot() +
-    aes(x = parameter, y = value) +
-    geom_segment(aes(x = parameter, y = 0, xend = parameter, yend = value)) +
-    geom_point() +
-    coord_flip() +
-    facet_grid(cols = vars(stat), rows = vars(variable), scales = "free_x") +
-    theme_bw(),
-  sensitivity_plot_png = save_plot(
[…]

### 2019-05-30 analysis/drake/markdown.R

@@ -1,0 +1,5 @@
+plan <- bind_plans(plan, drake_plan(
+  paper_md = rmarkdown::render(knitr_in(!!paper_file), "github_document"),
+  paper_pdf = rmarkdown::render(knitr_in(!!paper_file), "pdf_document"),
+  paper_html = rmarkdown::render(knitr_in(!!paper_file), "html_document")
+))

### 2019-05-30 analysis/drake/scatter_plot.R

@@ -1,0 +1,22 @@
+plan <- bind_plans(plan, drake_plan(
+  # Scatter plot of values in 2100
+  lastyear = filter(all_sims, year == 2100),
+  scatter_common = ggplot(lastyear) +
+    aes(y = value) +
+    facet_wrap(vars(variable), scales = "free_y") +
+    geom_point() +
+    ylab("") +
+    theme_bw(),
+  scatter = target(
+    scatter_common + aes(x = .x),
+    transform = map(.x = !!params)
+  ),
+  scatter_both = target(
+    plot_grid(scatter),
+    transform = combine(scatter)
+  ),
+  scatter_both_png = save_plot(
+    file_out(!!here::here("analysis", "figures", "scatter_both.png")),
+    scatter_both
+  )
+))

### 2019-05-30 analysis/drake/sensitivity.R

@@ -1,0 +1,23 @@
+plan <- bind_plans(plan, drake_plan(
+  sensitivity = lastyear %>%
+    group_by(variable) %>%
+    sensitivity_analysis(!!!params) %>%
+    tidy_sensitivity(),
+  sensitivity_plot = sensitivity %>%
+    filter(
+      parameter != "total",
+      stat != "sens",
+      stat != "var"
+    ) %>%
+    ggplot() +
+    aes(x = parameter, y = value) +
+    geom_segment(aes(x = parameter, y = 0, xend = parameter, yend = value)) +
+    geom_point() +
+    coord_flip() +
+    facet_grid(cols = vars(stat), rows = vars(variable), scales = "free_x") +
+    theme_bw(),
+  sensitivity_plot_png = save_plot(
+    file_out(!!here::here("analysis", "figures", "sensitivity_plot.png")),
+    sensitivity_plot
+  )
+))

### 2019-05-30 analysis/drake/simulations.R

@@ -1,0 +1,17 @@
+plan <- bind_plans(plan, drake_plan(
+  draws_plot = GGally::ggpairs(draws),
+  sims = target(
+    hector_with_params(
+      beta = beta,
+      q10_rh = q10_rh,
+      f_nppv = f_nppv,
+      f_nppd = f_nppd,
+      f_litterd = f_litterd
+    ),
+    transform = map(!!!draws)
+  ),
+  all_sims = target(
+    as_tibble(bind_rows(sims, .id = "id")),
+    transform = combine(sims)
+  )
+))

### 2019-05-30 analysis/drake/ts_plot.R

@@ -1,0 +1,24 @@
+plan <- bind_plans(plan, drake_plan(
+  ts_common = ggplot(all_sims) +
+    aes(x = year, y = value, group = id) +
+    geom_line() +
+    facet_wrap(vars(variable), scales = "free_y") +
+    scale_color_gradient(low = "grey", high = "red") +
+    ylab("") +
+    theme_bw() +
+    theme(
+      legend.position = "bottom"
+    ),
+  ts = target(
+    ts_common + aes(color = .y),
+    transform = map(.y = !!params)
+  ),
+  ts_both = target(
+    plot_grid(ts),
+    transform = combine(ts)
+  ),
+  ts_both_png = save_plot(
+    file_out(!!here::here("analysis", "figures", "ts_both.png")),
+    ts_both
+  )
+))

### 2019-05-30 R/split_biome.R
Draft of split_biome function

@@ -1,0 +1,56 @@
+split_biome <- function(core, biome_name,
+                        frac_veg = 0.5,
+                        frac_soil = frac_veg,
+                        frac_detritus = frac_soil,
+                        frac_npp_flux0 = frac_veg) {
+  frac_tbl <- tibble::tribble(
+    ~variable, ~frac,
+    hector::VEGC(), frac_veg,
+    hector::SOILC(), frac_soil,
+    hector::DETRITUSC(), frac_detritus,
+    hector::NPP_FLUX0(), frac_npp_flux0
+  )
+  inits <- hector::fetchvars(core, NA, frac_tbl[["variable"]])
+  new_biome <- inits %>%
+    dplyr::left_join(frac_tbl, by = "variable") %>%
+    dplyr::mutate(variable = gsub("global", biome_name, variable),
+                  value = value * frac) %>%
+    dplyr::select(-frac)
+  orig_biome <- inits %>%
+    dplyr::left_join(frac_tbl, by = "variable") %>%
+    dplyr::mutate(value = value * (1 - frac)) %>%
+    dplyr::select(-frac)
+  # Also, clone parameters
+  param_names <- c(
+    hector::BETA(),
+    hector::Q10_RH(),
+    hector::F_NPPD(),
+    hector::F_NPPV(),
+    hector::F_LITTERD(),
+    hector::WARMINGFACTOR()
+  )
+  params <- hector::fetchvars(core, NA, param_names)
+  new_params <- params %>%
+    dplyr::mutate(variable = gsub("global", biome_name, variable)) %>%
+    dplyr::bind_rows(params)
+  new_values <- dplyr::bind_rows(orig_biome, new_biome, new_params)
+  set_values <- new_values %>%
+    dplyr::select(dates = year, var = variable, value, unit = units) %>%
+    purrr::pwalk(hector::setvar, core = core)
+  invisible(hector::reset(core))
+  new_values
+}
+
+if (FALSE) {
+
+  ini <- system.file("input", "hector_rcp45.ini", package = "hector")
+  core <- hector::newcore(ini, suppresslogging = FALSE)
+  biome_name <- "boreal"
+  frac_veg <- 0.5
+  frac_soil <- frac_veg
+  frac_detritus <- frac_soil
+  frac_npp_flux0 <- frac_veg
+  v <- split_biome(core, biome_name)
+  ## hector::run(core)
+
+}

### 2019-05-30 R/hector_with_params.R
Add split_biome function, and Hector function tests

@@ -3,27 +3,31 @@
 #' To specify units, set the `unit` attribute of the parameter. 
 #'
 #' @param ... Named list of parameter values (e.g. `beta = 0.3`).
-#'   Names must match Hector parameter names. 
+#'   Names must match Hector parameter names.
 #' @param .dots A named list of arguments. Provides an alternative
 #'   specification to `...`
-#' @param rcp
-#' @return 
+#' @param rcp Representative carbon pathway (RCP) to use. One of
+#'   `"26"`, `"45"`, `"60"`, `"85"`
+#' @param core (Optional) An existing Hector core to modify.
+#' @return Hector output, as a `data.frame`
 #' @author Alexey Shiklomanov
 #' @examples
-#' hector_with_params(beta = 0.5, q10_rh = )
+#' hector_with_params(beta = 0.5, q10_rh = 1.8)
 #' @export
-hector_with_params <- function(..., .dots = list(), rcp = "45") {
+hector_with_params <- function(..., .dots = list(), rcp = "45", core = NULL) {
   params <- modifyList(.dots, list(...))
-  ini_file <- system.file(
-    "input",
-    paste0("hector_rcp", rcp, ".ini"),
-    package = "hector"
-  )
-  core <- hector::newcore(
-    ini_file,
-    name = "sensitivity",
-    suppresslogging = TRUE
-  )
+  if (is.null(core)) {
+    ini_file <- system.file(
+      "input",
+      paste0("hector_rcp", rcp, ".ini"),
+      package = "hector"
+    )
+    core <- hector::newcore(
+      ini_file,
+      name = "sensitivity",
+      suppresslogging = TRUE
+    )
+  }
   purrr::iwalk(
     params,
     ~hector::setvar(core, NA, .y, .x, NA)

### 2019-05-30 R/split_biome.R

@@ -1,8 +1,26 @@
-split_biome <- function(core, biome_name,
+#' Create a Hector core with a split biome
+#'
+#' @param biome_name (Character) Name of new biome
+#' @param frac_veg Fraction of global vegetation C in new biome.
+#'   Default is 0.5.
+#' @param frac_soil Fraction of global soil C in new biome. Default is `frac_veg`.
+#' @param frac_detritus Fraction of global detritus C in new biome.
+#'   Default is `frac_soil` (which in turn defaults to `frac_veg`)
+#' @param frac_npp_flux0 Fraction if global initial NPP flux in new
+#'   biome. Default is `frac_veg`.
+#' @inheritParams hector_with_params
+#' @return Hector core with biomes `"global"` and `biome_name`
+#' @author Alexey Shiklomanov
+#' @export
+split_biome <- function(biome_name,
                         frac_veg = 0.5,
                         frac_soil = frac_veg,
                         frac_detritus = frac_soil,
-                        frac_npp_flux0 = frac_veg) {
+                        frac_npp_flux0 = frac_veg,
+                        rcp = "45") {
+  ini_file <- system.file("input", paste0("hector_rcp", rcp, ".ini"),
+                          package = "hector")
+  ini <- hectortools::read_ini(ini_file)
   frac_tbl <- tibble::tribble(
     ~variable, ~frac,
     hector::VEGC(), frac_veg,
@@ -10,7 +28,10 @@ split_biome <- function(core, biome_name,
     hector::DETRITUSC(), frac_detritus,
     hector::NPP_FLUX0(), frac_npp_flux0
   )
-  inits <- hector::fetchvars(core, NA, frac_tbl[["variable"]])
+  inits <- ini[["simpleNbox"]][gsub("global\\.", "", frac_tbl[["variable"]])] %>%
+    unlist() %>%
+    tibble::enframe("variable", "value") %>%
+    dplyr::mutate(variable = paste0("global.", variable))
   new_biome <- inits %>%
     dplyr::left_join(frac_tbl, by = "variable") %>%
     dplyr::mutate(variable = gsub("global", biome_name, variable),
@@ -29,28 +50,19 @@ split_biome <- function(core, biome_name,
     hector::F_LITTERD(),
     hector::WARMINGFACTOR()
   )
-  params <- hector::fetchvars(core, NA, param_names)
-  new_params <- params %>%
[…]

### 2019-05-30 tests/testthat/test.hector_with_params.R

@@ -1,0 +1,12 @@
+context("hector_with_params function")
+
+test_that("Function works", {
+  r1 <- hector_with_params(beta = 0.4)
+  r2 <- hector_with_params(beta = 0.7)
+  expect_s3_class(r1, "data.frame")
+  expect_s3_class(r2, "data.frame")
+  expect_true(all(
+    subset(r1, variable == "Ca")[["value"]] >
+      subset(r2, variable == "Ca")[["value"]]
+  ))
+})

### 2019-05-30 tests/testthat/test.split_biome.R

@@ -1,0 +1,11 @@
+context("Run Hector with split biome")
+
+test_that("Hector split_biome works.", {
+  result <- hector_with_params(
+    global.beta = 0.9,
+    global.q10_rh = 1.8,
+    core = split_biome("boreal", frac_veg = 0.2)
+  )
+  expect_s3_class(result, "data.frame")
+  expect_true(!any(is.na(result[["value"]])))
+})

### 2019-05-30 R/hector_with_params.R
Pass arguments from `hector_with_params` to `split_biome`

@@ -3,7 +3,8 @@
 #' To specify units, set the `unit` attribute of the parameter. 
 #'
 #' @param ... Named list of parameter values (e.g. `beta = 0.3`).
-#'   Names must match Hector parameter names.
+#'   Names must match Hector parameter names, or arguments to
+#'   [split_biome()].
 #' @param .dots A named list of arguments. Provides an alternative
 #'   specification to `...`
 #' @param rcp Representative carbon pathway (RCP) to use. One of
@@ -13,21 +14,38 @@
 #' @author Alexey Shiklomanov
 #' @examples
 #' hector_with_params(beta = 0.5, q10_rh = 1.8)
+#' hector_with_params(
+#'   global.beta = 0.6,
+#'   permafrost.beta = 0.8,
+#'   biome_name = "permafrost",
+#'   frac_veg = 0.2
+#' )
 #' @export
 hector_with_params <- function(..., .dots = list(), rcp = "45", core = NULL) {
-  params <- modifyList(.dots, list(...))
+  raw_params <- modifyList(.dots, list(...))
+  all_names <- names(raw_params)
+  sb_names_all <- names(formals(split_biome))
+  sb_names <- intersect(all_names, sb_names_all)
   if (is.null(core)) {
     ini_file <- system.file(
       "input",
       paste0("hector_rcp", rcp, ".ini"),
       package = "hector"
     )
-    core <- hector::newcore(
-      ini_file,
-      name = "sensitivity",
-      suppresslogging = TRUE
-    )
+    if (length(sb_names) > 0) {
+      sb_params <- raw_params[sb_names]
+      if (is.null(sb_params[["biome_name"]])) biome_name <- "permafrost"
+      core <- do.call(split_biome, sb_params)
+    } else {
+      core <- hector::newcore(
+        ini_file,
+        name = "sensitivity",
+        suppresslogging = TRUE
+      )
+    }
   }
+  param_names <- setdiff(all_names, sb_names_all)
+  params <- raw_params[param_names]
   purrr::iwalk(
     params,
     ~hector::setvar(core, NA, .y, .x, NA)

### 2019-05-30 R/split_biome.R

@@ -64,5 +64,5 @@ split_biome <- function(biome_name,
     params, new_params, remove_values
   ))
   new_ini <- modifyList(ini, new_values)
-  hectortools::newcore_ini(new_ini)
+  hectortools::newcore_ini(new_ini, name = "split_biome")
 }

### 2019-06-03 R/hector_with_params.R
Submit Hector runs separate from workflow

@@ -20,6 +20,7 @@
 #'   biome_name = "permafrost",
 #'   frac_veg = 0.2
 #' )
+#' @importFrom magrittr %>%
 #' @export
 hector_with_params <- function(..., .dots = list(), rcp = "45", core = NULL) {
   raw_params <- modifyList(.dots, list(...))

### 2019-06-03 analysis/drake/biome-simulations.R

@@ -1,30 +1,36 @@
 ## Print the hector_with_params call args to console:
 ## writeLines(sprintf("%1$s = %1$s,", as.character(biome_params)))
 
+## plan <- bind_plans(plan, drake_plan(
+##   biome_draws_plot = GGally::ggpairs(draws),
+##   biome_sims = target(
+##     hector_with_params(
+##       global.beta = global.beta,
+##       global.q10_rh = global.q10_rh,
+##       global.f_litterd = global.f_litterd,
+##       permafrost.beta = permafrost.beta,
+##       permafrost.q10_rh = permafrost.q10_rh,
+##       permafrost.f_litterd = permafrost.f_litterd,
+##       global.f_nppv = global.f_nppv,
+##       global.f_nppd = global.f_nppd,
+##       permafrost.f_nppv = permafrost.f_nppv,
+##       permafrost.f_nppd = permafrost.f_nppd,
+##       frac_veg = frac_veg,
+##       frac_soil = frac_soil,
+##       frac_detritus = frac_detritus,
+##       biome_name = "permafrost"
+##     ),
+##     transform = map(!!!biome_draws)
+##   ),
+##   biome_all_sims = target(
+##     as_tibble(bind_rows(biome_sims, .id = "id")),
+##     transform = combine(biome_sims)
+##   ),
+##   biome_sims_out = write_csv(biome_all_sims, file_out("analysis/data/output/biome-sims.csv"))
+## ))
+
 plan <- bind_plans(plan, drake_plan(
-  biome_draws_plot = GGally::ggpairs(draws),
-  biome_sims = target(
-    hector_with_params(
-      global.beta = global.beta,
-      global.q10_rh = global.q10_rh,
-      global.f_litterd = global.f_litterd,
-      permafrost.beta = permafrost.beta,
-      permafrost.q10_rh = permafrost.q10_rh,
-      permafrost.f_litterd = permafrost.f_litterd,
-      global.f_nppv = global.f_nppv,
-      global.f_nppd = global.f_nppd,
-      permafrost.f_nppv = permafrost.f_nppv,
-      permafrost.f_nppd = permafrost.f_nppd,
-      frac_veg = frac_veg,
-      frac_soil = frac_soil,
-      frac_detritus = frac_detritus,
-      biome_name = "permafrost"
-    ),
-    transform = map(!!!biome_draws)
-  ),
-  biome_all_sims = target(
[…]

### 2019-06-03 analysis/drake/simulations.R

@@ -1,18 +1,18 @@
-plan <- bind_plans(plan, drake_plan(
-  draws_plot = GGally::ggpairs(draws),
-  sims = target(
-    hector_with_params(
-      beta = beta,
-      q10_rh = q10_rh,
-      f_nppv = f_nppv,
-      f_nppd = f_nppd,
-      f_litterd = f_litterd
-    ),
-    transform = map(!!!draws)
-  ),
-  all_sims = target(
-    as_tibble(bind_rows(sims, .id = "id")),
-    transform = combine(sims)
-  ),
-  all_sims_out = write_csv(all_sims, file_out("analysis/data/output/s1_sims.csv"))
-))
+## plan <- bind_plans(plan, drake_plan(
+##   draws_plot = GGally::ggpairs(draws),
+##   sims = target(
+##     hector_with_params(
+##       beta = beta,
+##       q10_rh = q10_rh,
+##       f_nppv = f_nppv,
+##       f_nppd = f_nppd,
+##       f_litterd = f_litterd
+##     ),
+##     transform = map(!!!draws)
+##   ),
+##   all_sims = target(
+##     as_tibble(bind_rows(sims, .id = "id")),
+##     transform = combine(sims)
+##   ),
+##   all_sims_out = write_csv(all_sims, file_out("analysis/data/output/s1_sims.csv"))
+## ))

### 2019-06-03 analysis/scripts/biome-draw-params.R

@@ -1,10 +1,18 @@
 #!/usr/bin/env Rscript
 library(dplyr)
 library(readr)
+library(clustermq)
+library(fst)
+library(here)
 library(hector.permafrost.emit)
 
+logdir <- here("logs")
+dir.create(logdir, recursive = TRUE, showWarnings = FALSE)
+outdir <- here("analysis", "data", "output")
+dir.create(outdir, recursive = TRUE, showWarnings = FALSE)
+
 set.seed(8675309)
-ndraws <- 1000
+ndraws <- 10000
 
 npp_alpha <- c(f_nppv = 0.35, f_nppd = 0.60, f_npps = 0.05)
 gl_npp_draws <- rdirichlet(ndraws, npp_alpha)[, 1:2] %>%
@@ -26,5 +34,20 @@ draws <- tibble(
 ) %>%
   bind_cols(as_tibble(gl_npp_draws), as_tibble(pf_npp_draws))
 
-write_csv(draws, file.path("analysis", "data",
-                           "derived_data", "biome-parameter-draws.csv"))
+write_csv(draws, here("analysis", "data",
+                      "derived_data", "biome-parameter-draws.csv"))
+
+hector_fun <- function(...) {
+  library(hector.permafrost.emit)
+  hector_with_params(...)
+}
+
+result <- Q_rows(draws, hector_fun,
+                 const = list(biome_name = "permafrost"),
+                 n_jobs = 50,
+                 template = list(log_file = file.path(logdir, "biome.log")))
+result_df <- result %>%
+  bind_rows() %>%
+  as_tibble()
+
+write_fst(result_df, file.path(outdir, "biome-sims.fst"))

### 2019-06-03 analysis/scripts/draw_params.R

@@ -1,10 +1,18 @@
 #!/usr/bin/env Rscript
 library(dplyr)
 library(readr)
+library(clustermq)
+library(fst)
+library(here)
 library(hector.permafrost.emit)
 
+logdir <- here("logs")
+dir.create(logdir, recursive = TRUE, showWarnings = FALSE)
+outdir <- here("analysis", "data", "output")
+dir.create(outdir, recursive = TRUE, showWarnings = FALSE)
+
 set.seed(8675309)
-ndraws <- 1000
+ndraws <- 5000
 
 npp_alpha <- c(f_nppv = 0.35, f_nppd = 0.60, f_npps = 0.05)
 npp_draws <- rdirichlet(ndraws, npp_alpha)[, 1:2]
@@ -16,5 +24,18 @@ draws <- tibble(
 ) %>%
   bind_cols(as_tibble(npp_draws))
 
-write_csv(draws, file.path("analysis", "data",
-                           "derived_data", "parameter-draws.csv"))
+write_csv(draws, here("analysis", "data",
+                      "derived_data", "parameter-draws.csv"))
+
+hector_fun <- function(...) {
+  library(hector.permafrost.emit)
+  hector_with_params(...)
+}
+
+result <- Q_rows(draws, hector_fun, n_jobs = 50,
+                 template = list(log_file = file.path(logdir, "global.log")))
+result_df <- result %>%
+  bind_rows() %>%
+  as_tibble()
+
+write_fst(result_df, file.path(outdir, "global-sims.fst"))

### 2019-06-03 analysis/paper/paper.Rmd
Add biome simulation figures

--- 2019-05-30 analysis/paper/paper.Rmd
@@
 # Results
+
+## Global Hector
+
+```{r fig-global-timeseries, fig.cap = global_ts_cap, fig.dim = c(10, 8)}
+global_ts_cap <- paste(
+  "Time series of Hector outputs for different parameter combinations."
+)
+readd(ts_global_sims)
+```
+
+```{r fig-global-scatter, fig.cap = global_scatter_cap, fig.dim = c(8, 5)}
+global_scatter_cap <- paste(
+  "Hector outputs at 2100 as a function of input parameter values."
+)
+readd(scatter_lastyear_global_sims)
+```
+
+```{r fig-global-sensitivity, fig.cap = global_sensitivity_cap, fig.dim = c(5, 5)}
+global_sensitivity_cap <- paste(
+  "Sensitivity and variance decomposition analysis for global Hector."
+)
+readd(sensitivity_plot_sensitivity_global_params_lastyear_global_sims)
+```
+
+## Permafrost as a biome
+
+```{r fig-biome-timeseries, fig.cap = biome_ts_cap, fig.dim = c(10, 8)}
+biome_ts_cap <- paste(
+  "Overall parameter uncertainty of Hector simulations with permafrost biome."
+)
+readd(ts_biome_sims)
+```
+
+```{r fig-biome-scatter, fig.cap = biome_scatter_cap, fig.dim = c(8, 8)}
+biome_scatter_cap <- paste(
+  "Hector outputs (with permafrost biome) at 2100 as a function of input parameter values."
+)
+readd(scatter_lastyear_biome_sims)
+```
+
+```{r fig-biome-sensitivity, fig.cap = biome_sensitivity_cap, fig.dim = c(5, 7)}
+biome_sensitivity_cap <- paste(
+  "Sensitivity and variance decomposition analysis for Hector with permafrost biome."
+)
+readd(sensitivity_plot_sensitivity_biome_params_lastyear_biome_sims)
+```
+
+## Permafrost methane emissions
 

### 2019-06-03 analysis/drake/01-global-setup.R

@@ -1,6 +1,28 @@
-draws <- read_csv(file.path("analysis", "data",
-                            "derived_data", "parameter-draws.csv"))
-if (!run_all) {
-  draws <- head(draws, 100)
-}
-params <- rlang::syms(colnames(draws))
+global_sims_head <- fst::read_fst(here::here(
+  "analysis", "data", "output", "global-sims.fst"
+), to = 1)
+global_params <- global_sims_head %>%
+  select(-(scenario:units)) %>%
+  colnames() %>%
+  rlang::syms()
+
+biome_sims_head <- here::here("analysis", "data", "output", "biome-sims.fst") %>%
+  fst::read_fst(to = 1)
+biome_params <- biome_sims_head %>%
+  select(-(scenario:units)) %>%
+  colnames() %>%
+  rlang::syms()
+
+param_types <- rlang::syms(c("global_params", "biome_params"))
+
+biomes_params <- bind_rows(
+  tibble(model = list(rlang::sym("global_sims")), params = global_params),
+  tibble(model = list(rlang::sym("biome_sims")), params = biome_params)
+)
+
+## draws <- read_csv(file.path("analysis", "data",
+##                             "derived_data", "parameter-draws.csv"))
+## if (!run_all) {
+##   draws <- head(draws, 100)
+## }
+## params <- rlang::syms(colnames(draws))

### 2019-06-03 analysis/drake/02-biome-setup.R

@@ -1,7 +1,7 @@
-biome_draws <- read_csv(file.path("analysis", "data",
-                                  "derived_data", "biome-parameter-draws.csv"),
-                        col_types = cols(.default = "d"))
-if (!run_all) {
-  biome_draws <- head(biome_draws, 100)
-}
-biome_params <- rlang::syms(colnames(biome_draws))
+## biome_draws <- read_csv(file.path("analysis", "data",
+##                                   "derived_data", "biome-parameter-draws.csv"),
+##                         col_types = cols(.default = "d"))
+## if (!run_all) {
+##   biome_draws <- head(biome_draws, 100)
+## }
+## biome_params <- rlang::syms(colnames(biome_draws))

### 2019-06-03 analysis/drake/biome-simulations.R

@@ -1,36 +1,0 @@
-## Print the hector_with_params call args to console:
-## writeLines(sprintf("%1$s = %1$s,", as.character(biome_params)))
-
-## plan <- bind_plans(plan, drake_plan(
-##   biome_draws_plot = GGally::ggpairs(draws),
-##   biome_sims = target(
-##     hector_with_params(
-##       global.beta = global.beta,
-##       global.q10_rh = global.q10_rh,
-##       global.f_litterd = global.f_litterd,
-##       permafrost.beta = permafrost.beta,
-##       permafrost.q10_rh = permafrost.q10_rh,
-##       permafrost.f_litterd = permafrost.f_litterd,
-##       global.f_nppv = global.f_nppv,
-##       global.f_nppd = global.f_nppd,
-##       permafrost.f_nppv = permafrost.f_nppv,
-##       permafrost.f_nppd = permafrost.f_nppd,
-##       frac_veg = frac_veg,
-##       frac_soil = frac_soil,
-##       frac_detritus = frac_detritus,
-##       biome_name = "permafrost"
-##     ),
-##     transform = map(!!!biome_draws)
-##   ),
-##   biome_all_sims = target(
-##     as_tibble(bind_rows(biome_sims, .id = "id")),
-##     transform = combine(biome_sims)
-##   ),
-##   biome_sims_out = write_csv(biome_all_sims, file_out("analysis/data/output/biome-sims.csv"))
-## ))
-
-plan <- bind_plans(plan, drake_plan(
-  biome_all_sims = fst::read_fst(file_in(!!here::here(
-    "analysis", "data", "output", "global-sims.csv"
-  ))) %>% as_tibble()
-))

### 2019-06-03 analysis/drake/scatter_plot.R

@@ -1,22 +1,0 @@
-plan <- bind_plans(plan, drake_plan(
-  # Scatter plot of values in 2100
-  lastyear = filter(all_sims, year == 2100),
-  scatter_common = ggplot(lastyear) +
-    aes(y = value) +
-    facet_wrap(vars(variable), scales = "free_y") +
-    geom_point() +
-    ylab("") +
-    theme_bw(),
-  scatter = target(
-    scatter_common + aes(x = .x),
-    transform = map(.x = !!params)
-  ),
-  scatter_both = target(
-    plot_grid(scatter),
-    transform = combine(scatter)
-  ),
-  scatter_both_png = save_plot(
-    file_out(!!here::here("analysis", "figures", "scatter_both.png")),
-    scatter_both
-  )
-))

### 2019-06-03 analysis/drake/scatter_sensitivity.R

@@ -1,0 +1,52 @@
+scatter_grid <- function(dat) {
+  params <- setdiff(colnames(dat), c("id", "scenario", "year",
+                                     "variable", "value", "units")) %>%
+    rlang::syms()
+  baseplot <- ggplot(dat) +
+    aes(y = value) +
+    facet_wrap(vars(variable), scales = "free_y") +
+    ## geom_point() +
+    geom_hex() +
+    geom_smooth(method = "gam", color = "black") +
+    ylab("") +
+    theme_bw() +
+    guides(fill = FALSE) +
+    scale_fill_gradient(low = "grey80", high = "red4")
+  plot_list <- lapply(params, function(X) baseplot + aes(x = !!X))
+  do.call(plot_grid, plot_list)
+}
+
+plan <- bind_plans(plan, drake_plan(
+  # Scatter plot of values in 2100
+  lastyear = target(
+    filter(.sims, year == 2100),
+    transform = map(.sims = c(global_sims, biome_sims))
+  ),
+  scatter = target(
+    scatter_grid(lastyear),
+    transform = map(lastyear)
+  ),
+  sensitivity = target(
+    lastyear %>%
+      group_by(variable) %>%
+      sensitivity_analysis(!!!.params) %>%
+      tidy_sensitivity(),
+    transform = map(lastyear, .params = !!param_types)
+  ),
+  sensitivity_plot = target(
+    sensitivity %>%
+      filter(
+        parameter != "total",
+        stat != "sens",
+        stat != "var"
+      ) %>%
+      ggplot() +
+      aes(x = parameter, y = value) +
+      geom_segment(aes(x = parameter, y = 0, xend = parameter, yend = value)) +
+      geom_point() +
+      coord_flip() +
+      facet_grid(cols = vars(stat), rows = vars(variable), scales = "free_x") +
+      theme_bw(),
+    transform = map(sensitivity)
+  )
+))

### 2019-06-03 analysis/drake/sensitivity.R

@@ -1,23 +1,0 @@
-plan <- bind_plans(plan, drake_plan(
-  sensitivity = lastyear %>%
-    group_by(variable) %>%
-    sensitivity_analysis(!!!params) %>%
-    tidy_sensitivity(),
-  sensitivity_plot = sensitivity %>%
-    filter(
-      parameter != "total",
-      stat != "sens",
-      stat != "var"
-    ) %>%
-    ggplot() +
-    aes(x = parameter, y = value) +
-    geom_segment(aes(x = parameter, y = 0, xend = parameter, yend = value)) +
-    geom_point() +
-    coord_flip() +
-    facet_grid(cols = vars(stat), rows = vars(variable), scales = "free_x") +
-    theme_bw(),
-  sensitivity_plot_png = save_plot(
-    file_out(!!here::here("analysis", "figures", "sensitivity_plot.png")),
-    sensitivity_plot
-  )
-))

### 2019-06-03 analysis/drake/simulations.R

@@ -1,3 +1,14 @@
+plan <- bind_plans(plan, drake_plan(
+  global_sims = file_in(!!here("analysis", "data",
+                            "output", "global-sims.fst")) %>%
+    fst::read_fst() %>%
+    as_tibble(),
+  biome_sims = file_in(!!here::here("analysis", "data",
+                                        "output", "biome-sims.fst")) %>%
+    fst::read_fst() %>%
+    as_tibble()
+))
+
 ## plan <- bind_plans(plan, drake_plan(
 ##   draws_plot = GGally::ggpairs(draws),
 ##   sims = target(

### 2019-06-03 analysis/drake/ts_plot.R

@@ -1,24 +1,26 @@
 plan <- bind_plans(plan, drake_plan(
-  ts_common = ggplot(all_sims) +
-    aes(x = year, y = value, group = id) +
-    geom_line() +
-    facet_wrap(vars(variable), scales = "free_y") +
-    scale_color_gradient(low = "grey", high = "red") +
-    ylab("") +
-    theme_bw() +
-    theme(
-      legend.position = "bottom"
-    ),
   ts = target(
-    ts_common + aes(color = .y),
-    transform = map(.y = !!params)
-  ),
-  ts_both = target(
-    plot_grid(ts),
-    transform = combine(ts)
-  ),
-  ts_both_png = save_plot(
-    file_out(!!here::here("analysis", "figures", "ts_both.png")),
-    ts_both
+    .sims %>%
+      group_by(year, variable, units) %>%
+      summarize(Mean = mean(value),
+                Median = median(value),
+                SD = sd(value),
+                lo1 = quantile(value, 0.05),
+                hi1 = quantile(value, 0.95),
+                lo2 = quantile(value, 0.025),
+                hi2 = quantile(value, 0.975)) %>%
+      ggplot() +
+      aes(x = year) +
+      geom_ribbon(aes(ymin = lo2, ymax = hi2, fill = "95% CI")) +
+      geom_ribbon(aes(ymin = lo1, ymax = hi1, fill = "90% CI")) +
+      scale_fill_manual(values = c("deepskyblue3", "deepskyblue1")) +
+      geom_line(aes(y = Mean)) +
+      facet_wrap(vars(variable), scales = "free_y") +
+      ylab("") +
+      theme_bw() +
+      theme(
+        legend.position = "bottom"
+      ),
+    transform = cross(.sims = c(global_sims, biome_sims))
   )
 ))

### 2019-06-03 analysis/scratch.R

@@ -163,3 +163,6 @@ plot(pd$beta, y$fit, ylim = range(y$fit + c(1, -1) * y$se.fit))
 ## points(dat$beta, dat$value, col = "red")
 lines(pd$beta, y$fit + y$se.fit, lty = "dashed")
 lines(pd$beta, y$fit - y$se.fit, lty = "dashed")
+
+plan %>% pull(target)
+readd(sensitivity_plot_sensitivity_lastyear_)

### 2019-06-03 analysis/paper/paper.Rmd
Add parameter draws to paper. Cleanup.

@@
 
-```{r fig-parameter-distribution, fig.cap = param_dist_cap}
-param_dist_cap <- paste(
-  "Input parameter distributions for global Hector."
-)
-readd(draws_plot)
-```
-
 # Results
@@
 ## Global Hector
+
+```{r fig-global-dist, fig.cap = global_dist_cap}
+global_dist_cap <- paste(
+  "Input parameter distributions for global Hector."
+)
+readd(draws_global_sims)
+```
 
@@
 ## Permafrost as a biome
+
+```{r fig-biome-dist, fig.cap = biome_dist_cap, fig.dim = c(9, 9)}
+biome_dist_cap <- paste(
+  "Input parameter distributions for global Hector."
+)
+readd(draws_biome_sims)
+```
 

### 2019-06-03 analysis/drake/01-global-setup.R

@@ -19,10 +19,3 @@ biomes_params <- bind_rows(
   tibble(model = list(rlang::sym("global_sims")), params = global_params),
   tibble(model = list(rlang::sym("biome_sims")), params = biome_params)
 )
-
-## draws <- read_csv(file.path("analysis", "data",
-##                             "derived_data", "parameter-draws.csv"))
-## if (!run_all) {
-##   draws <- head(draws, 100)
-## }
-## params <- rlang::syms(colnames(draws))

### 2019-06-03 analysis/drake/02-biome-setup.R

@@ -1,7 +1,0 @@
-## biome_draws <- read_csv(file.path("analysis", "data",
-##                                   "derived_data", "biome-parameter-draws.csv"),
-##                         col_types = cols(.default = "d"))
-## if (!run_all) {
-##   biome_draws <- head(biome_draws, 100)
-## }
-## biome_params <- rlang::syms(colnames(biome_draws))

### 2019-06-03 analysis/drake/markdown.R

@@ -1,5 +1,5 @@
 plan <- bind_plans(plan, drake_plan(
-  paper_md = rmarkdown::render(knitr_in(!!paper_file), "github_document"),
+  ## paper_md = rmarkdown::render(knitr_in(!!paper_file), "github_document"),
   paper_pdf = rmarkdown::render(knitr_in(!!paper_file), "pdf_document"),
   paper_html = rmarkdown::render(knitr_in(!!paper_file), "html_document")
 ))

### 2019-06-03 analysis/drake/simulations.R

@@ -1,12 +1,25 @@
+ggpairs_density <- function(data, mapping, ..., low = "grey80", high = "red4") {
+  ggplot(data = data, mapping = mapping) +
+    geom_hex(...) +
+    scale_fill_gradient(low = low, high = high)
+}
+
 plan <- bind_plans(plan, drake_plan(
   global_sims = file_in(!!here("analysis", "data",
                             "output", "global-sims.fst")) %>%
     fst::read_fst() %>%
     as_tibble(),
   biome_sims = file_in(!!here::here("analysis", "data",
-                                        "output", "biome-sims.fst")) %>%
+                                    "output", "biome-sims.fst")) %>%
     fst::read_fst() %>%
-    as_tibble()
+    as_tibble(),
+  draws = target(
+    .sims %>%
+      select(-(scenario:units)) %>%
+      GGally::ggpairs(lower = list(continuous = ggpairs_density)) +
+      theme_bw(),
+    transform = map(.sims = c(global_sims, biome_sims))
+  )
 ))
 
 ## plan <- bind_plans(plan, drake_plan(

### 2019-06-03 analysis/scratch.R

@@ -165,4 +165,4 @@ lines(pd$beta, y$fit + y$se.fit, lty = "dashed")
 lines(pd$beta, y$fit - y$se.fit, lty = "dashed")
 
 plan %>% pull(target)
-readd(sensitivity_plot_sensitivity_lastyear_)
+readd(draws_biome_sims)

### 2019-06-04 analysis/paper/paper.Rmd
Add some methods text to manuscript

@@
 
-Several studies have attempted to evaluate the potential economic impacts of current and future permafrost C emissions.
-@lawrence_2012_simulation project declines in near-surface permafrost by 33% and 72% in RCP 2.6 and 8.5, respectively, and permafrost C feedback will stabilize at 2100 in RCP 2.6.
-More recently, @hope_2015_economic projected permafrost C emissions using the SiBCASA (TODO: write in full) land surface model and used these projections as inputs to the PAGE09 integrated assessment model (IAM).
-They found that XXX.
-@chen_2019_economic also used PAGE09.
-Similarly, @gonzalez-eguino_2016_significant used the DICE (TODO: write in full) model to XXX.
-@kessler_2017_estimating also used DICE.
-
-@hope_2015_economic (and others?) rely on climate damage functions of the general form $\delta GDP = f(\delta T)$.
-Such functions are, at best, highly uncertain (TODO REF), and possibly conceptually flawed (TODO REF).
-(TODO More on this).
-GCAM (TODO full name) is an alternative (TODO REF).
-(TODO More on GCAM).
-
 (Paragraph on simple climate models.)
@@
 Simple climate model.
+
+(TODO: More details on terrestrial C cycle in Hector).
+The default heterotrophic respiration ($R$) scheme for a pool $p$ (detritus or soil) in Hector:
+
+$$ R_{p} = C_p \times f_p \times Q_{10} ^ \frac{T}{10} $$
+
+## Modeling workflow
+
+We evaluate the sensitivity and parametric uncertainty of three different versions of Hector.
+
+The simplest is the "global" version, which corresponds to the standard version of Hector described in [@hartin_2015_simple].
+The global land carbon sink is modeled as a single entity, with parameterizations corresponding to global averages (originally tuned to match global outputs from CMIP5; TODO REF?).
+In this configuration, we vary the following parameters:
+CO~2~ fertilization effect ($\beta$);
+temperature sensitivity of heterotrophic respiration, ($Q^{10}_{RH}$);
+fraction of net primary production C that goes to vegetation (`f_nppv`), detritus (`f_nppd`), and soil (`1 - f_nppv - f_nppd`);
+and
+fraction of vegetation litter C that goes to detritus (`f_litterd`).
+
+The second is the "biome" version, which is the same as the "global" version except that the land carbon sink is divided into two "biomes": "non-permafrost" and "permafrost".
+Each of these biomes has its own C pools, fluxes, and parameters.
+For this configuration, we varied all of the "global" parameters (previous paragraph) for _both_ biomes ($\beta_{permafrost}$, $\beta_{non-permafrost}$, $Q^10_{RH,permafrost}$, $Q^10_{RH,non-permafrost}$, etc.), as well as the distribution of global pools (vegetation, detritus, soil) and initial primary productivity (`npp_flux0`) aross these two biomes.
+
+The third is the "biome + methane" version.
+This version is the same as the "biome" version, but with an additional process that partitions total heterotrophic respiration ($R_{p,tot}$) into a CO~2~ flux ($R_{p, CO_2}$) and a CH~4~ flux ($R_p, CH_4$) as follows:
+
+$$ R_{p,tot} = C_p \times f_p \times Q_{10} ^ \frac{T}{10} $$
+
+$$ R_{p, CH_4} = \alpha R_{p, tot} $$
+
+$$ R_{p, CO_2} = (1 - \alpha) ^ \phi R_{p, tot} $$
+
+This process is controlled by two new biome-specific parameters:
+$\alpha$ can be interpreted as the fraction of heterotrophic respiration C that is given off as CH~4~ assuming a perfect trade-off between CO~2~ and CH~4~ emissions.
+$\phi$ defines the plasticity of this relationship;
+at $\phi$ = 1, plasticity is perfect, such that $R_{p,tot}$ is evenly split between CO~2~ and CH~4~;
+at $\phi$ = 0, CO~2~ emissions are entirely independent of CH~4~ emissions, such that total C emissions increase linearly with $\alpha$;
+at $\phi$ > 1, CO~2~ emissions decline faster than CH~4~ emissions increase.
+As such, the total C emissions attributable to heterotrophic respiration are expected to decrease as $\phi$ increases.
+In this version of Hector, we varied all of the parameters in the "biome" verson, as well as $\alpha$ and $\phi$ for each biome.
 
@@
 to generate estimates (but can back-calculate from 500 Gt C above?)
-
-## Modeling permafrost emissions
-
-Default heterotrophic respiration ($R$) scheme for a pool $p$ (detritus or soil) in Hector:
-
-$$ R_{p} = C_p \times f_p \times Q_{10} ^ \frac{T}{10} $$
-
-This is case 1.
-
-Case 2 uses the same formulation, but splits permafrost C into its own "biome", with its own photosynthesis ($\beta$) and respiration ($Q_{10}$) parameters.
-
-Case 3 extends Case 2 by calculating CO~2~ and CH~4~ emissions as a fraction of the total respiration, controlled by two new parameters $\alpha$ and $\phi$:
-
-$$ R_{p,tot} = C_p \times f_p \times Q_{10} ^ \frac{T}{10} $$
-
-$$ R_{p, CH_4} = \alpha R_{p, tot} $$
-
-$$ R_{p, CO_2} = (1 - \alpha) ^ \phi R_{p, tot} $$
 

### 2019-06-04 analysis/paper/paper.Rmd
Add more methods information

@@
 
-## Modeling workflow
+## Model configurations
 
@@
 Each of these biomes has its own C pools, fluxes, and parameters.
-For this configuration, we varied all of the "global" parameters (previous paragraph) for _both_ biomes ($\beta_{permafrost}$, $\beta_{non-permafrost}$, $Q^10_{RH,permafrost}$, $Q^10_{RH,non-permafrost}$, etc.), as well as the distribution of global pools (vegetation, detritus, soil) and initial primary productivity (`npp_flux0`) aross these two biomes.
+For this configuration, we varied all of the "global" parameters (previous paragraph) for _both_ biomes ($\beta_{permafrost}$, $\beta_{non-permafrost}$, $Q^10_{RH,permafrost}$, $Q^10_{RH,non-permafrost}$, etc.), as well as the distribution of global pools (vegetation, detritus, soil) and initial primary productivity (`npp_flux0`) across these two biomes and the relative warming factor (compared to the global average) of the permafrost biome.
 
@@
 
-## Permafrost emissions scenarios
+## Parameter distributions
 
-For all scenarios, we added the emission as a combination of fossil fuel CO2 emissions (`ffi_emissions` column in Hector) and methane emissions (`CH4_emissions` column in Hector).
-Where these were explicitly separated, we added them accordingly.
-Where they were presented only in terms of C, we did ???
+We drew parameters from loosely informative distributions that span the range of physical plausibility.
+The distributions are shown in the following table (unless otherwise specified, we used the same parameter distributions for global and biome cases):
 
-Digitized scenarios from [@schaefer_2011_amount].
-SiBCASA model predictions of CO2 emissions (permafrost respiration; $R_{pc}$; note -- no methane!) through 2300.
-These results were digitized using WebPlotDigitizer (https://apps.automeris.io/wpd/), and interpolated to annual resolution (using R `stats::spline` function).
+| Parameter                        | Distribution                    | Hector default value |
+|----------------------------------|---------------------------------|----------------------|
+| $\beta$                          | Uniform(0, 1)                   | 0.36                 |
+| Q^10^~RH~                        | Uniform(0, 10)                  | 2.0                  |
+| Litter-detritus fraction         | Beta(3.92, 0.08)                | 0.98                 |
+| NPP vegetation fraction          | Dirichlet(**0.35**, 0.60, 0.05) | 0.35                 |
+| NPP detritus fraction            | Dirichlet(0.35, **0.60**, 0.05) | 0.60                 |
+|----------------------------------|---------------------------------|----------------------|
+| Permafrost vegetation C fraction | Beta(1, 2)                      | --                   |
+| Permafrost soil C fraction       | Beta(1.1, 1.1)                  | --                   |
+| Permafrost detritus C fraction   | Beta(1.1, 1.1)                  | --                   |
+| Permafrost warming factor        | 1 + 3 * Beta(5, 5)              | --                   |
+|----------------------------------|---------------------------------|----------------------|
+| Non-permafrost $\alpha$          | TODO                            | --                   |
+| Non-permafrost $\phi$            | TODO                            |                      |
+| Permafrost $\alpha$              | TODO                            |                      |
+| Permafrost $\phi$                | TODO                            |                      |
 
-Digitized scenarios from [@hope_2015_economic].
-CO2 and CH4 emissions from SiBCASA model.
+We used 5000 Hector simulations for the global case and 10,000 times for the biome case (biome + methane case is TODO).
 
-[@schuur_2009_effect] -- Estimate 0.8 - 1.1 Pg C yr-1.
+## Analysis of results
 
-Back-of-the-envelope estimates from [@zimov_2006_climate_change]:
-- 500 Gt C in loess that could be completely emitted by 2100 (plus other C sources).
-- 10-40 g C m-3 day-1 off the bat, slowing down to equilibrium (?) rate of 0.5-5 g C m-3 day-1 for several years.
-Combine with data on permafrost spatial extent, density, etc.
-to generate estimates (but can back-calculate from 500 Gt C above?)
+We looked at four output variables:
+Atmospheric CO~2~ concentration (`Ca`),
+total radiative forcing (`Ftot`),
+radiative forcing from CO~2~ alone (`FCO2`),
+and
+global mean atmospheric temperature (`Tgav`).
 
-## Parameter calibration
+First, as a metric of overall parametric uncertainty for each case, we calculated the mean, standard deviation, and 90% and 95% quantiles of the time series for each variable for each simulation.
+Second, we evaluated the impact of parameters by plotting the values of each variable at 2100 as a function of input parameter value.
 
-We used the `BayesianTools` R package [@R-BayesianTools] for all parameter calibration.
-The outputs of these calibrations are joint posterior distributions of parameters and their covariances, from which we sample for the sensitivity analysis.
+Third, we performed a more rigorous sensitivity and uncertainty analysis based on @lebauer_2013_facilitating.
+As a metric of parameter uncertainty, we calculated the coefficient of variation (`cv`) for each input parameter as the ratio of its variance to its mean.
+We evaluated sensitivity as follows:
+First, we fit a multivariate Generalized Additive Model (GAM; `mgcv::gam` function in R) for each variable as a function of all the input parameters.
+Then, we calculate the sensitivity ($s$) to each parameter ($x_i$) as the slope (partial derivative) of the resulting GAM at that parameter's median ($\hat{x}$; discretized at +/- 1% of the median), holding all other parameters ($x_{i+1}..x_n$) constant:
 
-For global parameters, we used the following likelihood:
+$$ s = \frac{d(GAM(x_i | x_{i + 1} .. x_n))}{d(x_i)} = \frac{GAM(1.01 x_i | ...) - GAM(0.99 x_i | ...)}{1.01 x_i - 0.99 x_i}$$
 
-$$ \log(L) = \sum_s Normal(Hector(\beta, Q_{10}, s) | CMIP5(s), \sigma) $$
+We then define the "elasticity" ($\epsilon$) (or normalized sensitivity) of the model to that parameter as:
 
-where $s$ is one of the four representative carbon pathways (RCPs), $CMIP5(s)$ are the CMIP5 global mean outputs for the corresponding variables, and $\sigma$ is the model error (estimated during the fit).
-We also used the resulting distributions for $\beta$ and $Q_{10}$ for the non-permafrost biome in cases 2 and 3.
-We feel this is appropriate because the CMIP5 models against which these parameters are calibrated do not include permafrost C feedbacks.
+$$ \epsilon = s \div \frac{\hat{y}} {\hat{x}} $$
 
-For case 2, we calibrated the permafrost-specific $\beta$ and $Q_{10}$ against various literature sources, including:
+Finally, we calculate the "partial variance" of each parameter (the fractional contribution of each parameter's uncertainty to overall model predictive uncertainty) by passing the distribution of each parameter through the GAM while holding all other parameters constant, calculating the resulting predictive variance for each parameter, and then normalizing these values relative to the sum of all the variances:
 
-- Land surface model simulations [@hope_2015_economic; @schaefer_2011_amount; @burke_2017_quantifying].
-    - Try to calibrate against NPP and soil respiration if possible
-- Literature surveys [@schaefer_2014_impact] 
-- Warming experiments [@wickland_2006_effects]
-- TODO: Others?
+$$ Var(x_i) \approx Var[GAM(x_i | x_{i+1} .. x_n)] $$
 
-Some of these are time series, while others are individual estimates at particular points in time.
-To give them equal weight in the likelihood, we down-weight the time series likelihoods by the number of time steps.
-
-We derived a distribution for the Arctic warming factor from TODO.
-
-TODO: Table and multi-panel figure of input datasets.
-
-For the $\alpha$ and $\phi$ parameters in case 3, we looked at the literature on permafrost methane emissions [e.g., @wickland_2006_effects].
-
-## Initial conditions
-
-For case 1, we use the Hector defaults [@hartin_2015_simple].
-
-For cases 2 and 3, we assumed the same total pool sizes as case 1, but divided them into the permafrost and non-permafrost biomes according to TODO.
-
-## Other notes
-
-Frozen carbon residence time (FCRt) from [@burke_2017_quantifying]:
-
-$$ FCRt = FCRt0 * exp(-\Delta T / \Gamma) (for \Delta T > 0.2 °C) $$
-
-- $FCRt_0$ (years) reflects the stability of permafrost C (length of time that permafrost C is stable when $\Delta T = 0$)
-- $\Gamma$ -- decay term (°C); temperature change at which "the number of years taken for all of the old permafrost C to be emitted reduces to 1/e of its initial value"
-
-Other Hector parameters to consider.
-
-| Variable | INI name | Description                                 | Value |
-|----------|----------|---------------------------------------------|-------|
-| $f_{nv}$ | `f_nppv` | Fraction of NPP C transferred to vegetation | 0.35  |
-| $f_{nd}$ | `f_nppd` | Fraction of NPP C transferred to detritus   | 0.60  |
-| $f_{nd}$ |          | Fraction of NPP C transferred to soil       | 0.05  |
-| $f_{lv}$ | `f_lucv` | Fraction of LUC change flux from vegetation | 0.10  |
-| $f_{ld}$ | `f_lucd` | Fraction of LUC change flux from detritus   | 0.01  |
-| $f_{ls}$ |          | Fraction of LUC change flux from soil       | 0.89  |
-| $f_{ds}$ |          | Fraction of detritus C that goes to soil    | 0.60  |
-| $f_{rd}$ |          | Fraction of respiration C to detritus       | 0.25  |
-| $f_{rs}$ |          | Fraction of respriation C to soil           | 0.02  |
-
-According to [@hartin_2015_simple], these were selected to be "generally consistent with previous simple earth system models [e.g., @meinshausen_2011_emulating]".
+$$ pvar(x_i) = \frac{Var(x_i)}{\sum_{i}^n {Var(x_i)}} $$
 
@@
 
-## Global Hector
+## Global case
 
@@
 
-## Permafrost as a biome
+## Biome case
 
@@
 
+# Miscellaneous notes
+
+## Permafrost emissions scenarios
+
+Digitized scenarios from [@schaefer_2011_amount].
+SiBCASA model predictions of CO2 emissions (permafrost respiration; $R_{pc}$; note -- no methane!) through 2300.
+These results were digitized using WebPlotDigitizer (https://apps.automeris.io/wpd/), and interpolated to annual resolution (using R `stats::spline` function).
+
+Digitized scenarios from [@hope_2015_economic].
+CO2 and CH4 emissions from SiBCASA model.
+
+[@schuur_2009_effect] -- Estimate 0.8 - 1.1 Pg C yr-1.
+
+Back-of-the-envelope estimates from [@zimov_2006_climate_change]:
+- 500 Gt C in loess that could be completely emitted by 2100 (plus other C sources).
+- 10-40 g C m-3 day-1 off the bat, slowing down to equilibrium (?) rate of 0.5-5 g C m-3 day-1 for several years.
+Combine with data on permafrost spatial extent, density, etc.
+to generate estimates (but can back-calculate from 500 Gt C above?)
+
+## Parameter calibration
+
+We used the `BayesianTools` R package [@R-BayesianTools] for all parameter calibration.
+The outputs of these calibrations are joint posterior distributions of parameters and their covariances, from which we sample for the sensitivity analysis.
+
+For global parameters, we used the following likelihood:
+
+$$ \log(L) = \sum_s Normal(Hector(\beta, Q_{10}, s) | CMIP5(s), \sigma) $$
+
+where $s$ is one of the four representative carbon pathways (RCPs), $CMIP5(s)$ are the CMIP5 global mean outputs for the corresponding variables, and $\sigma$ is the model error (estimated during the fit).
+We also used the resulting distributions for $\beta$ and $Q_{10}$ for the non-permafrost biome in cases 2 and 3.
+We feel this is appropriate because the CMIP5 models against which these parameters are calibrated do not include permafrost C feedbacks.
+
+For case 2, we calibrated the permafrost-specific $\beta$ and $Q_{10}$ against various literature sources, including:
+
+- Land surface model simulations [@hope_2015_economic; @schaefer_2011_amount; @burke_2017_quantifying].
+    - Try to calibrate against NPP and soil respiration if possible
+- Literature surveys [@schaefer_2014_impact] 
+- Warming experiments [@wickland_2006_effects]
+- TODO: Others?
+
+Some of these are time series, while others are individual estimates at particular points in time.
+To give them equal weight in the likelihood, we down-weight the time series likelihoods by the number of time steps.
+
+We derived a distribution for the Arctic warming factor from TODO.
+
+TODO: Table and multi-panel figure of input datasets.
+
+For the $\alpha$ and $\phi$ parameters in case 3, we looked at the literature on permafrost methane emissions [e.g., @wickland_2006_effects].
+
+## Other notes
+
+Frozen carbon residence time (FCRt) from [@burke_2017_quantifying]:
+
+$$ FCRt = FCRt0 * exp(-\Delta T / \Gamma) (for \Delta T > 0.2 °C) $$
+
+- $FCRt_0$ (years) reflects the stability of permafrost C (length of time that permafrost C is stable when $\Delta T = 0$)
+- $\Gamma$ -- decay term (°C); temperature change at which "the number of years taken for all of the old permafrost C to be emitted reduces to 1/e of its initial value"
+
+Other Hector parameters to consider.
+
+| Variable | INI name | Description                                 | Value |
+|----------|----------|---------------------------------------------|-------|
+| $f_{nv}$ | `f_nppv` | Fraction of NPP C transferred to vegetation | 0.35  |
+| $f_{nd}$ | `f_nppd` | Fraction of NPP C transferred to detritus   | 0.60  |
+| $f_{nd}$ |          | Fraction of NPP C transferred to soil       | 0.05  |
+| $f_{lv}$ | `f_lucv` | Fraction of LUC change flux from vegetation | 0.10  |
+| $f_{ld}$ | `f_lucd` | Fraction of LUC change flux from detritus   | 0.01  |
+| $f_{ls}$ |          | Fraction of LUC change flux from soil       | 0.89  |
+| $f_{ds}$ |          | Fraction of detritus C that goes to soil    | 0.60  |
+| $f_{rd}$ |          | Fraction of respiration C to detritus       | 0.25  |
+| $f_{rs}$ |          | Fraction of respriation C to soil           | 0.02  |
+
+According to [@hartin_2015_simple], these were selected to be "generally consistent with previous simple earth system models [e.g., @meinshausen_2011_emulating]".
+
 <!-- The following line inserts a page break when the output is MS Word. For page breaks in PDF, use \newpage on its own line.  -->

### 2019-06-04 analysis/paper/paper.Rmd
Render large figures separately
This saves them having to be redrawn every time I fix something in the
paper file. The redrawing process takes forever.

@@
   - Ben Bond-Lamberty
-  - Ben Kravitz?
-  - Kate Calvin?
+  - Ben Kravitz
   - Corinne Hartin
@@
 )
-readd(draws_global_sims)
+# readd(draws_global_sims)
+knitr::include_graphics(file_in(here::here("analysis", "figures", "global_draws.png")))
 ```
@@
 )
-readd(draws_biome_sims)
+# readd(draws_biome_sims)
+knitr::include_graphics(file_in(here::here("analysis", "figures", "biome_draws.png")))
 ```

### 2019-06-04 analysis/drake/simulations.R

@@ -6,7 +6,7 @@ ggpairs_density <- function(data, mapping, ..., low = "grey80", high = "red4") {
 
 plan <- bind_plans(plan, drake_plan(
   global_sims = file_in(!!here("analysis", "data",
-                            "output", "global-sims.fst")) %>%
+                               "output", "global-sims.fst")) %>%
     fst::read_fst() %>%
     as_tibble(),
   biome_sims = file_in(!!here::here("analysis", "data",
@@ -19,24 +19,12 @@ plan <- bind_plans(plan, drake_plan(
       GGally::ggpairs(lower = list(continuous = ggpairs_density)) +
       theme_bw(),
     transform = map(.sims = c(global_sims, biome_sims))
+  ),
+  draws_png = target(
+    file_out(!!here("analysis", "figures", fname)) %>%
+      ggplot2::ggsave(plot = draws, width = width, height = height),
+    transform = map(draws, fname = c("global_draws.png", "biome_draws.png"),
+                    width = c(7, 9), height = c(7, 9),
+                    .id = draws)
   )
 ))
-
-## plan <- bind_plans(plan, drake_plan(
-##   draws_plot = GGally::ggpairs(draws),
-##   sims = target(
-##     hector_with_params(
-##       beta = beta,
-##       q10_rh = q10_rh,
-##       f_nppv = f_nppv,
-##       f_nppd = f_nppd,
-##       f_litterd = f_litterd
-##     ),
-##     transform = map(!!!draws)
-##   ),
-##   all_sims = target(
-##     as_tibble(bind_rows(sims, .id = "id")),
-##     transform = combine(sims)
-##   ),
-##   all_sims_out = write_csv(all_sims, file_out("analysis/data/output/s1_sims.csv"))
-## ))

### 2019-06-04 analysis/scripts/getcitations.R

@@ -1,0 +1,35 @@
+#!/usr/bin/env Rscript
+
+stopifnot(
+  requireNamespace("stringr", quietly = TRUE),
+  requireNamespace("bibtex", quietly = TRUE)
+)
+
+arg <- commandArgs(trailingOnly = TRUE)
+infile <- arg[1]
+outfile <- arg[2]
+reffile <- arg[3]
+
+if (is.na(infile)) infile <- "analysis/paper/paper.Rmd"
+if (is.na(outfile)) outfile <- "analysis/paper/references.bib"
+if (is.na(reffile)) reffile <- "~/Dropbox/references/library.bib"
+
+stopifnot(
+  file.exists(infile),
+  file.exists(reffile)
+)
+
+message("Searching for citations in file ", infile)
+raw_file <- readLines(infile)
+rxp <- "@[[:alnum:]_]+"
+results_raw <- stringr::str_match_all(raw_file, rxp)
+citations <- sort(unique(do.call(c, results_raw)))
+citations <- stringr::str_remove(citations, "@")
+
+# Read bibtex file
+message("Reading bibfile...")
+bibfile <- bibtex::read.bib(reffile)
+selected_references <- bibfile[citations]
+message("Writing ", length(selected_references), " citations to file ", outfile)
+bibtex::write.bib(selected_references, outfile)
+message("Done!")

### 2019-08-09 R/hector_with_params.R
Replace `split_biome` with `hector::split_biome`.

@@ -1,6 +1,6 @@
 #' Run Hector with parameters set to specific values
 #'
-#' To specify units, set the `unit` attribute of the parameter. 
+#' To specify units, set the `unit` attribute of the parameter.
 #'
 #' @param ... Named list of parameter values (e.g. `beta = 0.3`).
 #'   Names must match Hector parameter names, or arguments to
@@ -15,44 +15,43 @@
 #' @examples
 #' hector_with_params(beta = 0.5, q10_rh = 1.8)
 #' hector_with_params(
-#'   global.beta = 0.6,
+#'   default.beta = 0.6,
 #'   permafrost.beta = 0.8,
-#'   biome_name = "permafrost",
-#'   frac_veg = 0.2
+#'   fveg_c = 0.2
 #' )
 #' @importFrom magrittr %>%
 #' @export
 hector_with_params <- function(..., .dots = list(), rcp = "45", core = NULL) {
-  raw_params <- modifyList(.dots, list(...))
-  all_names <- names(raw_params)
-  sb_names_all <- names(formals(split_biome))
-  sb_names <- intersect(all_names, sb_names_all)
   if (is.null(core)) {
     ini_file <- system.file(
       "input",
       paste0("hector_rcp", rcp, ".ini"),
       package = "hector"
     )
-    if (length(sb_names) > 0) {
-      sb_params <- raw_params[sb_names]
-      if (is.null(sb_params[["biome_name"]])) biome_name <- "permafrost"
-      core <- tryCatch(
-        do.call(split_biome, sb_params),
-        error = function(e) {
-          message("Hit the following error while creating core:\n",
-                  conditionMessage(e), ".\n",
-                  "Returning `NULL`.")
-          return(NULL)
-        })
-      if (is.null(core)) return(NULL)
-    } else {
-      core <- hector::newcore(
-        ini_file,
-        name = "sensitivity",
-        suppresslogging = TRUE
-      )
-    }
+    stopifnot(file.exists(ini_file))
+    core <- hector::newcore(
+      ini_file,
+      name = "sensitivity",
+      suppresslogging = TRUE
+    )
   }
+
+  raw_params <- modifyList(.dots, list(...))
+  all_names <- names(raw_params)
+  sb_names_all <- names(formals(hector::split_biome))
+  sb_names <- intersect(all_names, sb_names_all)
+  sb_vars <- raw_params[sb_names]
+  if (length(sb_names) > 0) {
[…]

### 2019-08-09 R/split_biome.R

@@ -1,68 +1,0 @@
-#' Create a Hector core with a split biome
-#'
-#' @param biome_name (Character) Name of new biome
-#' @param frac_veg Fraction of global vegetation C in new biome.
-#'   Default is 0.5.
-#' @param frac_soil Fraction of global soil C in new biome. Default is `frac_veg`.
-#' @param frac_detritus Fraction of global detritus C in new biome.
-#'   Default is `frac_soil` (which in turn defaults to `frac_veg`)
-#' @param frac_npp_flux0 Fraction if global initial NPP flux in new
-#'   biome. Default is `frac_veg`.
-#' @inheritParams hector_with_params
-#' @return Hector core with biomes `"global"` and `biome_name`
-#' @author Alexey Shiklomanov
-#' @export
-split_biome <- function(biome_name,
-                        frac_veg = 0.5,
-                        frac_soil = frac_veg,
-                        frac_detritus = frac_soil,
-                        frac_npp_flux0 = frac_veg,
-                        rcp = "45") {
-  ini_file <- system.file("input", paste0("hector_rcp", rcp, ".ini"),
-                          package = "hector")
-  ini <- hectortools::read_ini(ini_file)
-  frac_tbl <- tibble::tribble(
-    ~variable, ~frac,
-    hector::VEGC(), frac_veg,
-    hector::SOILC(), frac_soil,
-    hector::DETRITUSC(), frac_detritus,
-    hector::NPP_FLUX0(), frac_npp_flux0
-  )
-  inits <- ini[["simpleNbox"]][gsub("global\\.", "", frac_tbl[["variable"]])] %>%
-    unlist() %>%
-    tibble::enframe("variable", "value") %>%
-    dplyr::mutate(variable = paste0("global.", variable))
-  new_biome <- inits %>%
-    dplyr::left_join(frac_tbl, by = "variable") %>%
-    dplyr::mutate(variable = gsub("global", biome_name, variable),
-                  value = value * frac) %>%
-    dplyr::select(-frac)
-  orig_biome <- inits %>%
-    dplyr::left_join(frac_tbl, by = "variable") %>%
-    dplyr::mutate(value = value * (1 - frac)) %>%
-    dplyr::select(-frac)
-  # Also, clone parameters
-  param_names <- c(
-    hector::BETA(),
-    hector::Q10_RH(),
-    hector::F_NPPD(),
-    hector::F_NPPV(),
[…]

### 2019-08-09 tests/testthat/test.split_biome.R

@@ -1,22 +1,0 @@
-context("Run Hector with split biome")
-
-test_that("Hector split_biome works.", {
-  result <- hector_with_params(
-    global.beta = 0.9,
-    global.q10_rh = 1.8,
-    core = split_biome("boreal", frac_veg = 0.2)
-  )
-  expect_s3_class(result, "data.frame")
-  expect_true(!any(is.na(result[["value"]])))
-})
-
-test_that("hector_with_params args passed to split_biome", {
-  result <- hector_with_params(
-    global.beta = 0.9,
-    global.q10_rh = 1.8,
-    biome_name = "boreal",
-    frac_veg = 0.2
-  )
-  expect_s3_class(result, "data.frame")
-  expect_true(!any(is.na(result[["value"]])))
-})

### 2019-08-09 analysis/scripts/biome-draw-params.R
Update biome-draw-params to use new system

@@ -10,32 +10,34 @@ logdir <- here("logs")
 dir.create(logdir, recursive = TRUE, showWarnings = FALSE)
 outdir <- here("analysis", "data", "output")
 dir.create(outdir, recursive = TRUE, showWarnings = FALSE)
+datadir <- here("analysis", "data", "derived_data")
+dir.create(datadir)
 
 set.seed(8675309)
 ndraws <- 10000
 
 npp_alpha <- c(f_nppv = 0.35, f_nppd = 0.60, f_npps = 0.05)
 gl_npp_draws <- rdirichlet(ndraws, npp_alpha)[, 1:2] %>%
-  `colnames<-`(., paste0("global.", colnames(.)))
+  `colnames<-`(., paste0("default.", colnames(.)))
 pf_npp_draws <- rdirichlet(ndraws, npp_alpha)[, 1:2] %>%
   `colnames<-`(., paste0("permafrost.", colnames(.)))
 
 draws <- tibble(
-  global.beta = runif(ndraws, 0, 1),
-  global.q10_rh = runif(ndraws, 0, 10),
-  global.f_litterd = rbeta(ndraws, 0.98 * 4, 0.02 * 4),
+  default.beta = runif(ndraws, 0, 1),
+  default.q10_rh = runif(ndraws, 0, 10),
+  default.f_litterd = rbeta(ndraws, 0.98 * 4, 0.02 * 4),
   permafrost.beta = runif(ndraws, 0, 1),
   permafrost.q10_rh = runif(ndraws, 0, 10),
   permafrost.f_litterd = rbeta(ndraws, 0.98 * 4, 0.02 * 4),
-  frac_veg = rbeta(ndraws, 1, 2),
-  frac_soil = rbeta(ndraws, 1.1, 1.1),
-  frac_detritus = rbeta(ndraws, 1.1, 1.1),
-  warmingfactor = 1 + 3 * rbeta(ndraws, 5, 5)
+  fveg_c = rbeta(ndraws, 1, 2),
+  fsoil_c = rbeta(ndraws, 1.1, 1.1),
+  fdetritus_c = rbeta(ndraws, 1.1, 1.1),
+  ## fnpp_flux0 = rbeta(ndraws, 1.1, 1.1),
+  permafrost.warmingfactor = 1 + 3 * rbeta(ndraws, 5, 5)
 ) %>%
   bind_cols(as_tibble(gl_npp_draws), as_tibble(pf_npp_draws))
 
-write_csv(draws, here("analysis", "data",
[…]

### 2019-08-09 analysis/paper/paper.Rmd
Add some notes to paper

@@
 Total northern soil C pool is estimated at 1672 PgC, of which approximately 1466 Pg (88%) is in permafrost [@tarnocai_2009_soil].
+More recent estimates suggest that total high-latitude soil organic matter pool is 1300 Pg, with roughly 800 Pg in permafrost [@hugelius_2014_estimated].
 This frozen C can be released into the atmosphere by several processes related to warming, including aerobic respiration in thawed soil (REF) and anaeribic respiration in thermokarst lakes and wetlands [@turetsky_2002_boreal; @wickland_2006_effects].
@@
 On the other hand, warming and CO2 fertilization may increase vegetation productivity, which could increase soil C storage through enhanced litterfall; the balance of these two processes is uncertain [@jones_2005_global].
-Several modeling studies generally predict increases in soil C in high latitudes [@burke_2017_quantifying; @qian_2010_enhanced; @ito_2016_impacts].
-However, this increase is dampened by including permafrost C [@burke_2017_quantifying].
+There are additional uncertainties associated with vegetation composition (spatially variable peat accumulation) [@camill_2001_changes].
+Other large uncertainties related to soil properties and climate model structure [@harp_2016_effect].
+Several modeling studies generally predict increases in soil C sequestration at high latitudes [@burke_2017_quantifying; @qian_2010_enhanced; @ito_2016_impacts].
+However, this increase is dampened or even reversed when C emissions from permafrost thaw are included [@schaefer_2011_amount; @burke_2017_quantifying].
 
 (Previous work on modeling permafrost)
+@schaefer_2011_amount -- SiBCASA estimates 190 +/- 64 Gt C, but this does not include warming feedback or discontinuous permafrost regions.
+@harp_2016_effect -- Community Earth System Model (CESM) simulations over 100 years.
 [@burke_2017_quantifying] -- JULES and ORCHIDEE, with new permafrost scheme, combined with intermediate complexity climate-ocean model (IMOGEN), to look at climate sensitivity to permafrost C emissions.
 Models are highly sensitive to representation of soil processes, which can be more important than differences in scenario and/or climate drivers [@burke_2017_quantifying].
+But @harp_2016_effect argue that model structural uncertainty is larger than soil property uncertainty.
 
@@
 Hector [@hartin_2015_simple].
-However, Hector does not have an explicit representation of permafrost C emissions.
+In this study, we use Hector to evaluate the global sensitivty of climate to terrestrial biosphere processes, and how this uncertainty 
+The current version of Hector does not have an explicit representation of permafrost C emissions.
 In this study, we investigate whether the additional complexity and parametric and structural uncertainty of an explicit representation of permafrost may be warranted in Hector.

### 2019-08-09 analysis/scripts/draw_params.R
Update draw_params.R

@@ -10,6 +10,8 @@ logdir <- here("logs")
 dir.create(logdir, recursive = TRUE, showWarnings = FALSE)
 outdir <- here("analysis", "data", "output")
 dir.create(outdir, recursive = TRUE, showWarnings = FALSE)
+datadir <- here("analysis", "data", "derived_data")
+dir.create(datadir, recursive = TRUE, showWarnings = FALSE)
 
 set.seed(8675309)
 ndraws <- 5000
@@ -24,8 +26,7 @@ draws <- tibble(
 ) %>%
   bind_cols(as_tibble(npp_draws))
 
-write_csv(draws, here("analysis", "data",
-                      "derived_data", "parameter-draws.csv"))
+write_csv(draws, file.path(datadir, "parameter-draws.csv"))
 
 hector_fun <- function(...) {
   library(hector.permafrost.emit)

### 2019-08-09 R/hector_with_params.R
Explicit error message in hector_with_params

@@ -60,9 +60,8 @@ hector_with_params <- function(..., .dots = list(), rcp = "45", core = NULL) {
       ~hector::setvar(core, NA, .y, .x, NA)
     )
   }, error = function(e) {
-    message("Failed to set parameter value. ",
-            "Probably hit this error during `reset`. ",
-            "Returning `NULL`.")
+    message("Failed to set parameter value. Returning NULL. ",
+            "Error was:\n", conditionMessage(e))
     return(NULL)
   })
   if (is.null(.iout)) return(NULL)

### 2019-08-09 analysis/scripts/biome-draw-params.R
Setup clustermq with ssh
Can now use PIC for execution, but running from a local R session. Super cool!

@@ -6,12 +6,10 @@ library(fst)
 library(here)
 library(hector.permafrost.emit)
 
-logdir <- here("logs")
-dir.create(logdir, recursive = TRUE, showWarnings = FALSE)
 outdir <- here("analysis", "data", "output")
 dir.create(outdir, recursive = TRUE, showWarnings = FALSE)
 datadir <- here("analysis", "data", "derived_data")
-dir.create(datadir)
+dir.create(datadir, recursive = TRUE, showWarnings = FALSE)
 
 set.seed(8675309)
 ndraws <- 10000
@@ -41,13 +39,23 @@ write_csv(draws, file.path(datadir, "biome-parameter-draws.csv"))
 
 hector_fun <- function(...) {
   library(hector.permafrost.emit)
-  hector_with_params(...)
+  tryCatch(
+    hector_with_params(...),
+    error = function(e) {
+      message("Run failed, returning NULL. ",
+              "Hit the following error:\n",
+              conditionMessage(e))
+      return(NULL)
+    }
+  )
 }
 
 result <- Q_rows(draws, hector_fun,
-                 const = list(biome_name = "permafrost"),
                  n_jobs = 50,
-                 template = list(log_file = file.path(logdir, "biome.log")))
+                 template = list(
+                   account = "epa-ccd",
+                   log_file = "/people/shik544/.cmq_logs/biome-%a.log" #nolint
+                 ))
 result_df <- result %>%
   bind_rows() %>%
   as_tibble()

### 2019-08-09 analysis/scripts/draw_params.R

@@ -30,11 +30,22 @@ write_csv(draws, file.path(datadir, "parameter-draws.csv"))
 
 hector_fun <- function(...) {
   library(hector.permafrost.emit)
-  hector_with_params(...)
+  tryCatch(
+    hector_with_params(...),
+    error = function(e) {
+      message("Run failed, returning NULL. ",
+              "Hit the following error:\n",
+              conditionMessage(e))
+      return(NULL)
+    }
+  )
 }
 
 result <- Q_rows(draws, hector_fun, n_jobs = 50,
-                 template = list(log_file = file.path(logdir, "global.log")))
+                 template = list(
+                   account = "epa-ccd",
+                   log_file = "/people/shik544/.cmq_logs/global-%a.log" #nolint
+                 ))
 result_df <- result %>%
   bind_rows() %>%
   as_tibble()

### 2019-08-12 analysis/paper/paper.Rmd
Clarify permafrost emissions in introduction

@@
 Hector [@hartin_2015_simple].
-In this study, we use Hector to evaluate the global sensitivty of climate to terrestrial biosphere processes, and how this uncertainty 
+In this study, we use Hector to evaluate the global sensitivty of climate to terrestrial biosphere processes, and how this uncertainty is affected by permafrost emissions.
 The current version of Hector does not have an explicit representation of permafrost C emissions.

### 2019-08-12 analysis/scratch.R
Scratch code for multivariate sensitivity analysis

@@ -166,3 +166,74 @@ lines(pd$beta, y$fit - y$se.fit, lty = "dashed")
 
 plan %>% pull(target)
 readd(draws_biome_sims)
+
+##################################################
+#
+##################################################
+loadd(lastyear_biome_sims)
+loadd(lastyear_global_sims)
+
+mvfit <- function(dat) {
+  dat <- filter(dat, variable == "Tgav")
+  dat %>%
+    select(-scenario, -year, -variable, -units, -value) %>%
+    colnames() %>%
+    paste(collapse = " + ") %>%
+    sprintf(fmt = "value ~ (%s) ^ 2") %>%
+    as.formula() %>%
+    lm(data = dat) %>%
+    step()
+}
+gfit <- mvfit(lastyear_global_sims)
+summary(gfit)
+
+corrplot::corrplot(cov2cor(vcov(gfit)), method = "number", col = cp)
+
+bfit <- mvfit(lastyear_biome_sims)
+summary(bfit)[["coefficients"]][,c(1,4)] %>%
+  as_tibble(rownames = "term") %>%
+  ## filter(`Pr(>|t|)` < 0.01) %>%
+  arrange(desc(abs(Estimate))) %>%
+  select(term, Estimate) %>%
+  print(n = Inf)
+
+lm(value ~ permafrost.warmingfactor, data = filter(lastyear_biome_sims, variable == "Tgav"))
+
+lastyear_biome_sims %>%
+  filter(variable == "Tgav") %>%
+  ggplot() +
+  aes(x = permafrost.warmingfactor, y = value) +
+  geom_point() +
+  geom_smooth()
+
+corrplot::corrplot(cov2cor(vcov(bfit)), method = "number", col = cp)
+
+summary(filter(lastyear_global_sims, variable == "Tgav"))
+
+cp <- colorRampPalette(c("red4", "grey80", "blue4"))(200)
+
+ggplot(dat) +
+  aes(x = default.beta, y = default.q10_rh, color = value) +
+  geom_point() +
+  scale_color_gradient(low = "grey80", high = "red4")
+
+ggp2 <- function(data, mapping, ...) {
+  ggplot(data = data, mapping = mapping) +
+    geom_hex(stat = )
+}
+
+dat %>%
+  select(-scenario, -year, -variable, -units) %>%
+  GGally::ggpairs(aes(color = value),
+                  diag = list(continuous = "blank"),
+                  upper = list(continuous = "blank"),
[…]

### 2019-08-12 analysis/drake/01-global-setup.R
Add global setup comments

@@ -1,3 +1,5 @@
+# Get the list of global parameter names, for use in drake plan. Note that only
+# the first row of the output file is loaded.
 global_sims_head <- fst::read_fst(here::here(
   "analysis", "data", "output", "global-sims.fst"
 ), to = 1)
@@ -6,6 +8,7 @@ global_params <- global_sims_head %>%
   colnames() %>%
   rlang::syms()
 
+# Similarly, get the list of biome parameter names.
 biome_sims_head <- here::here("analysis", "data", "output", "biome-sims.fst") %>%
   fst::read_fst(to = 1)
 biome_params <- biome_sims_head %>%

### 2019-08-12 analysis/drake.R
Update drake system to use make

@@ -26,6 +26,7 @@ expose_imports("hector.permafrost.emit")
 args <- commandArgs(trailingOnly = TRUE)
 run_all <- "--all" %in% args
 cluster <- "--cluster" %in% args
+make <- "make" %in% args
 
 paper_file <- here("analysis", "paper", "paper.Rmd")
 
@@ -34,34 +35,34 @@ plan <- drake_plan()
 drake_files <- list.files(here("analysis", "drake"), full.names = TRUE)
 for (drake_file in drake_files) source(drake_file)
 
-if (interactive()) {
-  ## callr::rscript(here::here("analysis", "drake.R"), cmdargs = "--all")
-  callr::rscript(here::here("analysis", "drake.R"))
+if (cluster) {
+  parallelism <- "clustermq"
+  jobs <- 50
 } else {
-  if (cluster) {
-    parallelism <- "clustermq"
-    jobs <- 50
-  } else {
-    parallelism <- "future"
-    jobs <- parallel::detectCores()
-  }
-  dconf <- drake_config(
-    plan,
-    parallelism = parallelism,
-    jobs = jobs,
-    prework = quote({
-      pkgload::load_all(".", attach_testthat = FALSE, quiet = TRUE)
-    })
-  )
+  parallelism <- "future"
+  jobs <- parallel::detectCores()
+}
+dconf <- drake_config(
+  plan,
+  parallelism = parallelism,
+  jobs = jobs,
+  prework = quote({
+    pkgload::load_all(".", attach_testthat = FALSE, quiet = TRUE)
+  })
+)
 
-  # Set number of cores depending on number of outdated tasks
-  dout <- outdated(dconf)
-  if (!cluster && length(dout) > 10) {
-    dconf[["jobs"]] <- parallel::detectCores()
-    message("Number of outdated targets (", length(dout), ") ",
-            "is greater than 10. ",
-            "Running in parallel across ", dconf[["jobs"]], " cores.")
-  }
+# Set number of cores depending on number of outdated tasks
+dout <- outdated(dconf)
+if (!cluster && length(dout) > 10) {
+  dconf[["jobs"]] <- parallel::detectCores()
+  message("Number of outdated targets (", length(dout), ") ",
+          "is greater than 10. ",
+          "Running in parallel across ", dconf[["jobs"]], " cores.")
+}
 
-  make(config = dconf)
+if (make) {
+  message("Running drake::make.")
+  drake::make(config = dconf)
+} else if (!interactive()) {
+  dconf
 }

### 2019-08-12 analysis/scratch.R
Scratch code for testing sensitivity analysis

@@ -237,3 +237,9 @@ dat <- lastyear_biome_sims
 plot_list[[6]]
 for (i in 7:14) print(plot_list[[i]])
 do.call(plot_grid, plot_list)
+
+##################################################
+s <- lastyear_global_sims %>%
+  group_by(variable) %>%
+  sensitivity_analysis(!!!global_params)
+param_types

### 2019-09-11 analysis/paper/paper.Rmd
Add some notes to paper

@@
 
-Climate change is a problem (CITE: IPCC, NCA, etc.).
+Climate change is a problem [@ipcc_ar5_wg1].
 Important to reduce carbon emissions to meet temperature targets.
@@
 Permafrost is an important C reservoir.
-Total northern soil C pool is estimated at 1672 PgC, of which approximately 1466 Pg (88%) is in permafrost [@tarnocai_2009_soil].
+The total northern soil C pool is estimated at 1672 PgC, of which approximately 1466 Pg (88%) is in permafrost [@tarnocai_2009_soil].
 More recent estimates suggest that total high-latitude soil organic matter pool is 1300 Pg, with roughly 800 Pg in permafrost [@hugelius_2014_estimated].
@@
 [@schuur_2009_effect] estimate 0.8 - 1.1 Pg C yr-1.
-Back-of-the-envelope estimates from [@zimov_2006_climate_change]: 10-40 g C m-3 day-1 off the bat, slowing down to equilibrium (?) rate of 0.5-5 g C m-3 day-1 for several years.
+Back-of-the-envelope estimates from [@zimov_2006_climate_change]:
+10-40 g C m-3 day-1 off the bat, slowing down to equilibrium (?) rate of 0.5-5 g C m-3 day-1 for several years.
+
+Rapid methane release from arctic permafrost could potentially cause trillions in economic damage [@whiteman_2013_vast].
 
@@
 
-(Previous work on modeling permafrost)
+The importance of Arctic climate feedbacks has led to efforts to incorporate these processes---particularly, permafrost thaw---into land surface models.
+Early efforts derived permafrost extent and C based purely on soil temperature [@lawrence_2005_projection].
+(More recent models have done this better...)
 @schaefer_2011_amount -- SiBCASA estimates 190 +/- 64 Gt C, but this does not include warming feedback or discontinuous permafrost regions.
@@
 
-(Paragraph on simple climate models.)
+Land surface models are expensive to run, making it challening to use them for uncertainty quantification and exploration of alternative policy scenarios.
+Simple climate models are an alternative.
+(More on simple climate models).
+
+(Previous attempts to incorporate permafrost into simple climate models.)
+
 Hector [@hartin_2015_simple].

### 2019-09-13 R/sensitivity_analysis.R
Dramatically simplify sensitivity analysis code
Don't bother with the crazy tidyeval. Can do everything with mostly
base-R much more transparently, reliably, and concisely.

@@ -31,123 +31,57 @@
 #' derivatives.
 #'
 #' @param df `data.frame` of Hector results.
-#' @param ... Unquoted names of columns describing parameters to be
-#'   used in sensitivity analysis.
+#' @param xcols Character vector of parameter column names
+#' @param ycol Name of column containing response variable
 #' @param .type Whether the multivariate fit is "additive" (`Y ~ a + b +
 #'   c`, default) or "interactive" (`Y ~ a * b * c`). This argument
 #'   supports partial matching via [base::match.arg()].
 #' @return `data.frame` of sensitivity analysis results. See Details.
 #' @author Alexey Shiklomanov
 #' @export
-sensitivity_analysis <- function(df, ..., .type = "additive") {
-  .type <- match.arg(.type, c("additive", "interactive"))
-  .collapse <- switch(.type,
-                      "additive" = " + ",
-                      "interactive" = " * ")
+sensitivity_analysis <- function(dat, xcols, ycol, .type = "additive") {
+  modtype <- match.arg(.type, c("additive", "interactive"))
+  .collapse <- switch(modtype, additive = " + ", interactive = " * ")
+  rhs <- paste(sprintf("s(%s)", xcols), collapse = .collapse)
+  form <- paste(ycol, rhs, sep = " ~ ")
+  fit <- mgcv::gam(as.formula(form), data = dat)
+  xmed <- apply(dat[, xcols], 2, median)
+  xcv <- (apply(dat[, xcols], 2, var) / xmed) %>%
+    tibble::enframe("param", "cv")
+  pred <- apply(dat[, xcols], 2, quantile, probs = c(0.49, 0.51))
+  ymed <- median(dat[[ycol]])
 
-  params_q <- rlang::enquos(...)
-  params_s <- purrr::map_chr(params_q, rlang::quo_name)
-  pqs <- purrr::map(params_q, pq) %>% purrr::reduce(c)
-  rhs <- paste(sprintf("s(%s)", params_s), collapse = .collapse)
-  form <- paste0("value ~ ", rhs)
-  pred_qs <- purrr::map(seq_along(params_q), pred_q, x = params_q) %>%
-    purrr::reduce(c)
-  name_vars <- purrr::map(params_s, mksym, "..var")
-  partial_vars <- purrr::map(name_vars, ~rlang::quo(!!.x / total..var))
[…]

### 2019-09-13 analysis/drake/scatter_sensitivity.R
Use new sensitivity analysis in drake plan

@@ -29,19 +29,22 @@ plan <- bind_plans(plan, drake_plan(
   sensitivity = target(
     lastyear %>%
       group_by(variable) %>%
-      sensitivity_analysis(!!!.params) %>%
-      tidy_sensitivity(),
+      nest() %>%
+      mutate(sens = purrr::map(
+        data,
+        sensitivity_analysis,
+        xcols = as.character(.params),
+        ycol = "value"
+      )) %>%
+      select(-data) %>%
+      unnest(sens),
     transform = map(lastyear, .params = !!param_types)
   ),
   sensitivity_plot = target(
     sensitivity %>%
-      filter(
-        parameter != "total",
-        stat != "sens",
-        stat != "var"
-      ) %>%
+      gather(stat, value, vars(cv:partial_var)) %>%
       ggplot() +
-      aes(x = parameter, y = value) +
+      aes(x = param, y = value) +
       geom_segment(aes(x = parameter, y = 0, xend = parameter, yend = value)) +
       geom_point() +
       coord_flip() +

### 2019-09-13 analysis/drake/scatter_sensitivity.R
Fix drake bugs

@@ -42,10 +42,11 @@ plan <- bind_plans(plan, drake_plan(
   ),
   sensitivity_plot = target(
     sensitivity %>%
-      gather(stat, value, vars(cv:partial_var)) %>%
+      gather(stat, value, cv:partial_var) %>%
+      filter(stat != "pred_var") %>%
       ggplot() +
       aes(x = param, y = value) +
-      geom_segment(aes(x = parameter, y = 0, xend = parameter, yend = value)) +
+      geom_segment(aes(x = param, y = 0, xend = param, yend = value)) +
       geom_point() +
       coord_flip() +
       facet_grid(cols = vars(stat), rows = vars(variable), scales = "free_x") +

### 2019-09-13 analysis/drake/simulations.R

@@ -16,6 +16,7 @@ plan <- bind_plans(plan, drake_plan(
   draws = target(
     .sims %>%
       select(-(scenario:units)) %>%
+      distinct() %>%
       GGally::ggpairs(lower = list(continuous = ggpairs_density)) +
       theme_bw(),
     transform = map(.sims = c(global_sims, biome_sims))

### 2019-09-13 analysis/paper/paper.Rmd
Add note about permafrost surface C fraction
Add markdown captions

@@
 library(hector.permafrost.emit) # Or use devtools::load_all('.', quiet = T) if your code is in script files, rather than as functions in the `/R` diretory
+
+n <- 0
+mdcap <- function(text) {
+  n <<- n + 1
+  out <- sprintf("**Figure %d**: %s", n, text)
+  if (grepl("gfm", knitr::opts_knit$get("rmarkdown.pandoc.to"), ignore.case = TRUE)) {
+    out
+  } else {
+    invisible(NULL)
+  }
+}
+
+fig <- function(...) {
+  filepath <- here("analysis", "figures", ...)
+  knitr::include_graphics(file_in(filepath))
+}
 ```
@@
 
+Actual permafrost surface C fraction is around 30% [@schuur_2015_climate].
+
 We used 5000 Hector simulations for the global case and 10,000 times for the biome case (biome + methane case is TODO).
@@
 
+`r mdcap(global_ts_cap)`
+
 ```{r fig-global-scatter, fig.cap = global_scatter_cap, fig.dim = c(8, 5)}
@@
 ```
+
+`r mdcap(global_scatter_cap)`
 
@@
 
+`r mdcap(biome_ts_cap)`
+
 ```{r fig-biome-scatter, fig.cap = biome_scatter_cap, fig.dim = c(8, 8)}
@@
 
+`r mdcap(biome_scatter_cap)`
+
 ```{r fig-biome-sensitivity, fig.cap = biome_sensitivity_cap, fig.dim = c(5, 7)}
@@
 ```
+
+`r mdcap(biome_sensitivity_cap)`
 

### 2019-09-13 analysis/paper/paper.Rmd
Add univariate param density figure

@@
 
-library(hector.permafrost.emit) # Or use devtools::load_all('.', quiet = T) if your code is in script files, rather than as functions in the `/R` diretory
+library(hector.permafrost.emit)
 
@@
 
-fig <- function(...) {
-  filepath <- here("analysis", "figures", ...)
-  knitr::include_graphics(file_in(filepath))
+f <- function(...) {
+  here::here("analysis", "figures", ...)
 }
@@
 
+These distributions are also shown in Figure 1.
+
+```{r param-density, echo = FALSE, fig.cap = param_density_cap}
+param_density_cap <- paste(
+  "Input parameter distributions for this analysis."
+)
+knitr::include_graphics(file_in(f("param-density.png")))
+```
+
+`r mdcap(param_density_cap)`
+
 Actual permafrost surface C fraction is around 30% [@schuur_2015_climate].
@@
 
-```{r fig-global-dist, fig.cap = global_dist_cap}
-global_dist_cap <- paste(
-  "Input parameter distributions for global Hector."
-)
-# readd(draws_global_sims)
-knitr::include_graphics(file_in(here::here("analysis", "figures", "global_draws.png")))
-```
-
 ```{r fig-global-timeseries, fig.cap = global_ts_cap, fig.dim = c(10, 8)}
@@
 
+`r mdcap(global_sensitivity_cap)`
+
 ## Biome case
-
-```{r fig-biome-dist, fig.cap = biome_dist_cap, fig.dim = c(9, 9)}
-biome_dist_cap <- paste(
-  "Input parameter distributions for global Hector."
-)
-# readd(draws_biome_sims)
-knitr::include_graphics(file_in(here::here("analysis", "figures", "biome_draws.png")))
-```
 
@@
 
+# Appendix
+
+```{r fig-global-dist, fig.cap = global_dist_cap}
+global_dist_cap <- paste(
+  "Input parameter distributions for global Hector."
+)
+# readd(draws_global_sims)
+knitr::include_graphics(file_in(f("global_draws.png")))
+```
+
+`r mdcap(global_dist_cap)`
+
+```{r fig-biome-dist, fig.cap = biome_dist_cap, fig.dim = c(9, 9)}
+biome_dist_cap <- paste(
+  "Input parameter distributions for Hector with biomes."
+)
+# readd(draws_biome_sims)
+knitr::include_graphics(file_in(fig("biome_draws.png")))
+```
+
+`r mdcap(biome_dist_cap)`
+
 ### Colophon

### 2019-09-13 analysis/drake.R

@@ -1,7 +1,7 @@
 #!/usr/bin/env Rscript
 if (getRversion() >= "3.6.0") {
   library(drake, exclude = c("expand", "gather", "plan"))
-  library(ggplot2, exclude = "ggsave")
+  library(ggplot2)
 } else {
   library(drake)
   library(ggplot2)

### 2019-09-13 analysis/drake/parameter-densities.R

@@ -1,0 +1,33 @@
+plan <- bind_plans(plan, drake_plan(
+  global_inparams = read_csv(file_in(!!here::here(
+    "analysis", "data", "derived_data", "parameter-draws.csv"
+  ))),
+  biome_inparams = read_csv(file_in(!!here::here(
+    "analysis", "data", "derived_data", "biome-parameter-draws.csv"
+  ))) %>%
+    select(matches("warmingfactor|_c")),
+  densities = target(
+    purrr::map(.inparams, density, cut = 0) %>%
+      purrr::map_dfr(~bind_cols(.x[c("x", "y")]), .id = "param"),
+    transform = map(.inparams = c(global_inparams, biome_inparams))
+  ),
+  both_densities = target(bind_rows(densities), transform = combine(densities)),
+  param_density_plot = both_densities %>%
+    mutate(param = forcats::fct_inorder(param)) %>%
+    group_by(param) %>%
+    mutate(y = y / max(y)) %>%
+    ungroup() %>%
+    filter(!(param == "f_litterd" & x < 0.95)) %>%
+    ggplot() +
+    aes(x = x, y = y) +
+    geom_area() +
+    lims(y = c(0, 1)) +
+    labs(x = "Value", y = "Normalized probability density") +
+    facet_wrap(vars(param), scales = "free_x") +
+    cowplot::theme_cowplot(),
+  param_density_plot_file = ggsave(
+    file_out(!!here::here("analysis", "figures", "param-density.png")),
+    param_density_plot,
+    width = 10.3, height = 7.3
+  )
+))

### 2019-09-13 analysis/paper/paper.Rmd
Clean up captions, etc. in RMArkdown doc

@@
 
-```{r, setup, echo = FALSE}
-knitr::opts_chunk$set(
+```{r setup, echo = FALSE}
+library(drake)
+library(knitr)
+library(here)
+
+opts_chunk$set(
   collapse = TRUE,
@@
 library(hector.permafrost.emit)
+
+fig <- function(...) {
+  filepath <- here("analysis", "figures", ...)
+  knitr::include_graphics(file_in(filepath))
+}
 
@@
 
-f <- function(...) {
-  here::here("analysis", "figures", ...)
+m <- 0
+htmlfig <- function() {
+  m <<- m + 1
+  out <- paste0("**Figure ", m, "**:")
+  if (grepl("html", knitr::opts_knit$get("rmarkdown.pandoc.to"), ignore.case = TRUE)) {
+    out
+  } else {
+    invisible(NULL)
+  }
 }
@@
 param_density_cap <- paste(
+  htmlfig(),
   "Input parameter distributions for this analysis."
 )
-knitr::include_graphics(file_in(f("param-density.png")))
+fig("param-density.png")
 ```
@@
 global_ts_cap <- paste(
+  htmlfig(),
   "Time series of Hector outputs for different parameter combinations."
@@
 global_scatter_cap <- paste(
+  htmlfig(),
   "Hector outputs at 2100 as a function of input parameter values."
@@
 global_sensitivity_cap <- paste(
+  htmlfig(),
   "Sensitivity and variance decomposition analysis for global Hector."
@@
 biome_scatter_cap <- paste(
+  htmlfig(),
   "Hector outputs (with permafrost biome) at 2100 as a function of input parameter values."
@@
 biome_sensitivity_cap <- paste(
+  htmlfig(),
   "Sensitivity and variance decomposition analysis for Hector with permafrost biome."
@@
 global_dist_cap <- paste(
+  htmlfig(),
   "Input parameter distributions for global Hector."
@@
 # readd(draws_global_sims)
-knitr::include_graphics(file_in(f("global_draws.png")))
+fig("global_draws.png")
 ```
@@
 biome_dist_cap <- paste(
+  htmlfig(),
   "Input parameter distributions for Hector with biomes."
@@
 # readd(draws_biome_sims)
-knitr::include_graphics(file_in(fig("biome_draws.png")))
+fig("biome_draws.png")
 ```

### 2019-09-13 analysis/paper/paper.Rmd
Add summary time-series figure

@@
 
-## Global case
-
-```{r fig-global-timeseries, fig.cap = global_ts_cap, fig.dim = c(10, 8)}
-global_ts_cap <- paste(
+```{r fig-summary-ts, fig.cap = summary_ts_cap}
+summary_ts_cap <- paste(
   htmlfig(),
-  "Time series of Hector outputs for different parameter combinations."
+  "Mean and 95% confidence interval on Hector simulations ",
+  "with different model structures and parameter combinations."
 )
-readd(ts_global_sims)
+fig("ts-summary.png")
 ```
 
-`r mdcap(global_ts_cap)`
+`r mdcap(summary_ts_cap)`
+
+## Global case
 
@@
 ## Biome case
-
-```{r fig-biome-timeseries, fig.cap = biome_ts_cap, fig.dim = c(10, 8)}
-biome_ts_cap <- paste(
-  "Overall parameter uncertainty of Hector simulations with permafrost biome."
-)
-readd(ts_biome_sims)
-```
-
-`r mdcap(biome_ts_cap)`
 
@@
 
+```{r fig-global-timeseries, fig.cap = global_ts_cap, fig.dim = c(10, 8)}
+global_ts_cap <- paste(
+  htmlfig(),
+  "Time series of Hector outputs for different parameter combinations."
+)
+readd(ts_global_sims)
+```
+
+`r mdcap(global_ts_cap)`
+
+```{r fig-biome-timeseries, fig.cap = biome_ts_cap, fig.dim = c(10, 8)}
+biome_ts_cap <- paste(
+  htmlfig(),
+  "Overall parameter uncertainty of Hector simulations with permafrost biome."
+)
+readd(ts_biome_sims)
+```
+
+`r mdcap(biome_ts_cap)`
+
 ### Colophon

### 2019-09-13 analysis/drake/ts_plot.R

@@ -1,15 +1,45 @@
 plan <- bind_plans(plan, drake_plan(
-  ts = target(
+  ts_summary = target(
     .sims %>%
       group_by(year, variable, units) %>%
-      summarize(Mean = mean(value),
-                Median = median(value),
-                SD = sd(value),
-                lo1 = quantile(value, 0.05),
-                hi1 = quantile(value, 0.95),
-                lo2 = quantile(value, 0.025),
-                hi2 = quantile(value, 0.975)) %>%
-      ggplot() +
+      summarize(
+        Mean = mean(value),
+        Median = median(value),
+        SD = sd(value),
+        lo1 = quantile(value, 0.05),
+        hi1 = quantile(value, 0.95),
+        lo2 = quantile(value, 0.025),
+        hi2 = quantile(value, 0.975)
+      ),
+    transform = map(.sims = c(global_sims, biome_sims))
+  ),
+  sims_summary = target(
+    bind_rows(ts_summary, .id = "simulation_type") %>%
+      mutate(simulation_type = forcats::fct_recode(
+        simulation_type,
+        "global" = "1",
+        "biome" = "2"
+      )),
+    transform = combine(ts_summary)
+  ),
+  sims_summary_plot = ggplot(sims_summary) +
+    aes(x = year, color = simulation_type) +
+    geom_line(aes(y = Mean)) +
+    ## geom_line(aes(y = lo2), linetype = "dashed") +
+    ## geom_line(aes(y = hi2), linetype = "dashed") +
+    geom_ribbon(aes(ymin = lo2, ymax = hi2, fill = simulation_type),
+                alpha = 0.2, color = NA) +
+    ## geom_ribbon(aes(ymin = lo1, ymax = hi1), alpha = 0.7) +
+    facet_wrap(vars(variable), scales = "free_y") +
+    labs(x = "Year", y = "Mean +/- 95% CI") +
+    cowplot::theme_cowplot(),
+  sims_summary_plot_png = ggsave(
+    file_out(!!here::here("analysis", "figures", "ts-summary.png")),
+    sims_summary_plot,
+    width = 6.62, height = 4.66
+  ),
+  ts = target(
+    ggplot(ts_summary) +
       aes(x = year) +
       geom_ribbon(aes(ymin = lo2, ymax = hi2, fill = "95% CI")) +
       geom_ribbon(aes(ymin = lo1, ymax = hi1, fill = "90% CI")) +
[…]

### 2019-09-13 analysis/paper/paper.Rmd
Save more figures as PNG to speed up build

@@
 
-```{r fig-global-scatter, fig.cap = global_scatter_cap, fig.dim = c(8, 5)}
+```{r fig-global-scatter, fig.cap = global_scatter_cap}
 global_scatter_cap <- paste(
@@
 )
-readd(scatter_lastyear_global_sims)
+fig("global-scatter.png")
 ```
@@
 )
-readd(scatter_lastyear_biome_sims)
+fig("biome-scatter.png")
 ```
@@
 )
-# readd(draws_global_sims)
 fig("global_draws.png")
@@
 )
-# readd(draws_biome_sims)
 fig("biome_draws.png")

### 2019-09-13 analysis/drake/scatter_sensitivity.R

@@ -26,6 +26,15 @@ plan <- bind_plans(plan, drake_plan(
     scatter_grid(lastyear),
     transform = map(lastyear)
   ),
+  scatter_png = target(
+    ggsave(
+      file_out(!!here::here("analysis", "figures", .fname)),
+      scatter,
+      width = .width, height = .height
+    ),
+    transform = map(scatter, .fname = c("global-scatter.png", "biome-scatter.png"),
+                    .width = c(8, 10), .height = c(5, 8))
+  ),
   sensitivity = target(
     lastyear %>%
       group_by(variable) %>%

### 2019-09-13 analysis/drake/ts_plot.R

@@ -52,5 +52,13 @@ plan <- bind_plans(plan, drake_plan(
         legend.position = "bottom"
       ),
     transform = map(ts_summary)
+  ),
+  ts_png = target(
+    ggsave(
+      file_out(!!here::here("analysis", "figures", .fname)),
+      ts,
+      width = 10, height = 8
+    ),
+    transform = map(ts, .fname = c("global-ts.png", "biome-ts.png"))
   )
 ))

### 2019-09-13 analysis/paper/paper.Rmd
Bugfix sensitivity analysis, and save as PNG

@@
 )
-readd(sensitivity_plot_sensitivity_global_params_lastyear_global_sims)
+fig("sensitivity-global.png")
 ```
@@
 
-```{r fig-biome-scatter, fig.cap = biome_scatter_cap, fig.dim = c(8, 8)}
+```{r fig-biome-scatter, fig.cap = biome_scatter_cap}
 biome_scatter_cap <- paste(
@@
 
-```{r fig-biome-sensitivity, fig.cap = biome_sensitivity_cap, fig.dim = c(5, 7)}
+```{r fig-biome-sensitivity, fig.cap = biome_sensitivity_cap}
 biome_sensitivity_cap <- paste(
@@
 )
-readd(sensitivity_plot_sensitivity_biome_params_lastyear_biome_sims)
+fig("sensitivity-biome.png")
 ```

### 2019-09-13 R/sensitivity_analysis.R

@@ -49,6 +49,7 @@ sensitivity_analysis <- function(dat, xcols, ycol, .type = "additive") {
   xcv <- (apply(dat[, xcols], 2, var) / xmed) %>%
     tibble::enframe("param", "cv")
   pred <- apply(dat[, xcols], 2, quantile, probs = c(0.49, 0.51))
+  dpred <- apply(pred, 2, diff)
   ymed <- median(dat[[ycol]])
 
   # Elasticity
@@ -63,7 +64,8 @@ sensitivity_analysis <- function(dat, xcols, ycol, .type = "additive") {
   )
   names(el_inputs) <- xcols
   el_results <- lapply(el_inputs, predict, object = fit)
-  el_slope <- (vapply(el_results, diff, numeric(1)) / (ymed / xmed)) %>%
+  el_slope <- ((vapply(el_results, diff, numeric(1)) / dpred) /
+                 (ymed / xmed)) %>%
     tibble::enframe("param", "elasticity")
 
   # Prediction variance

### 2019-09-13 analysis/drake/scatter_sensitivity.R

@@ -49,17 +49,33 @@ plan <- bind_plans(plan, drake_plan(
       unnest(sens),
     transform = map(lastyear, .params = !!param_types)
   ),
-  sensitivity_plot = target(
+  sensitivity_wide = target(
     sensitivity %>%
       gather(stat, value, cv:partial_var) %>%
-      filter(stat != "pred_var") %>%
-      ggplot() +
+      filter(stat != "pred_var"),
+    transform = map(sensitivity),
+  ),
+  sensitivity_plot = target(
+    ggplot(sensitivity_wide) +
       aes(x = param, y = value) +
       geom_segment(aes(x = param, y = 0, xend = param, yend = value)) +
       geom_point() +
       coord_flip() +
       facet_grid(cols = vars(stat), rows = vars(variable), scales = "free_x") +
       theme_bw(),
-    transform = map(sensitivity)
+    transform = map(sensitivity_wide)
+  ),
+  sensitivity_plot_png = target(
+    ggsave(
+      file_out(!!here::here("analysis", "figures", .fname)),
+      sensitivity_plot,
+      width = .width, height = .height
+    ),
+    transform = map(
+      sensitivity_plot,
+      .fname = c("sensitivity-global.png",
+                 "sensitivity-biome.png"),
+      .width = c(5, 5), .height = c(5, 7)
+    )
   )
 ))

### 2019-09-26 analysis/drake.R
Dummy target to make markdown depend on fig pngs

@@ -30,6 +30,8 @@ make <- "make" %in% args
 
 paper_file <- here("analysis", "paper", "paper.Rmd")
 
+figfile <- function(...) here::here("analysis", "figures", ...)
+
 # Construct plan from components in "analysis/drake" directory
 plan <- drake_plan()
 drake_files <- list.files(here("analysis", "drake"), full.names = TRUE)

### 2019-09-26 analysis/drake/99-markdown.R

@@ -1,5 +1,17 @@
+dig <- function(...) digest::digest(..., file = TRUE)
+
 plan <- bind_plans(plan, drake_plan(
-  ## paper_md = rmarkdown::render(knitr_in(!!paper_file), "github_document"),
-  paper_pdf = rmarkdown::render(knitr_in(!!paper_file), "pdf_document"),
-  paper_html = rmarkdown::render(knitr_in(!!paper_file), "html_document")
+  figs_dir_t = list(
+    dig(file_in(!!figfile("param-density.png"))),
+    dig(file_in(!!figfile("ts-summary.png"))),
+    dig(file_in(!!figfile("global-scatter.png"))),
+    dig(file_in(!!figfile("biome-scatter.png"))),
+    dig(file_in(!!figfile("sensitivity-biome.png"))),
+    dig(file_in(!!figfile("global_draws.png"))),
+    dig(file_in(!!figfile("biome_draws.png")))
+  ),
+  paper = target(
+    rmarkdown::render(knitr_in(!!paper_file), .format),
+    transform = map(.format = c("pdf_document", "html_document"))
+  )
 ))

### 2019-09-26 analysis/drake/parameter-densities.R

@@ -20,14 +20,15 @@ plan <- bind_plans(plan, drake_plan(
     filter(!(param == "f_litterd" & x < 0.95)) %>%
     ggplot() +
     aes(x = x, y = y) +
-    geom_area() +
+    geom_area(fill = "gray85", color = "black") +
     lims(y = c(0, 1)) +
     labs(x = "Value", y = "Normalized probability density") +
     facet_wrap(vars(param), scales = "free_x") +
+    theme(strip.background = element_blank()) +
     cowplot::theme_cowplot(),
   param_density_plot_file = ggsave(
-    file_out(!!here::here("analysis", "figures", "param-density.png")),
+    file_out(!!figfile("param-density.png")),
     param_density_plot,
-    width = 10.3, height = 7.3
+    width = 10.3, height = 7.30
   )
 ))

### 2020-01-21 analysis/paper/paper.Rmd
Dummy target to make markdown depend on fig pngs
Draft permafrost (sigmoid) submodel

@@
 library(hector.permafrost.emit)
+
+# Dummy target to force-rebuild this when figures change
+invisible(readd(figs_dir_t))
 
@@
 
-Hector [@hartin_2015_simple,@hartin_2016_ocean].
+Hector [@hartin_2015_simple,@hartin_2016_ocean,].
 Simple climate model.
@@
 $$ R_{p} = C_p \times f_p \times Q_{10} ^ \frac{T}{10} $$
+
+## Hector permafrost sub-model
+
+The general principle is that permafrost constitutes an additional reserve of soil carbon that, because it is frozen, is inaccessible to microbes.
+As permafrost thaws and some fraction of this carbon becomes accessible to microbes, it is transferred from the frozen permafrost pool to the standard soil C pool, where it is decomposed by Hector's standard decomposition routine.
+Let $C[t]_{pf}$ be the permafrost C pool and $C[t]_{s}$ be the soil C pool at time $t$ (both in units of Pg C), and let $\Delta C[t]_{pf}$ be the change in the permafrost C pool at time $t$.
+The C consequences of permafrost thaw can therefore be represented as:
+
+\[ C[t]_{pf} = C[t-1]_{pf} - \Delta C[t]_{pf} \]
+
+\[ C[t]_{s} = C[t]_{s} + \Delta C[t]_{pf} \]
+
+Let $C[0]_{pf}$ be the initial size of the permafrost C pool (Pg C) and $\Phi[t]$ be the fraction of permafrost remaining (in arbitrary area or volume units) at time $t$.
+Assuming a uniform permafrost C density, $\Delta C[t]_{pf}$ can be expressed as:
+
+\[ \Delta C[t]_{pf} = \Phi[t] C[0]_{pf} \]
+
+To a first approximation, $\Phi[t]$ is a function of air temperature ($T[t]_{air}$).
+[Kessler (2015)][kessler] assume this relationship is linear.
+However, because the permafrost area fraction is, by definition, bounded by zero (and 1 [^1]), and because deeper permafrost thaws more slowly than shallow permafrost, we use a negative logistic relationship instead:
+
+[^1]: Technically, permafrost area could increase in the case of cooling temperatures, and therefore the area fraction could be greater than 1.
+However, because even the most aggressive climate action scenarios show temperatures that stabilize above year 2000, we assume that permafrost area will never grow more than the starting value.
+
+\[
+\Phi[t] = \frac{1}{1 + \exp\left( m T[t]_{air} + b \right)}
+\]
+
+where $m$ and $b$ are model parameters.
+For globally-averaged permafrost, $m = 0.829$ and $b = 4.078$ most closely reproduce the rate of 0.172 K$^{-1}$ reported by [Kessler (2015)] over the range of 2 to 6 K$^{-1}$.
+
+```{r sigmoid, echo = FALSE}
+sigmoid_cap <- paste(
+  "Fraction of permafrost thaw as a function of change in",
+  "global annual mean air temperature since pre-industrial (1750)."
+)
+m <- 0.829
+b <- 4.078
+kessler_m <- 0.172
+kessler_b <- 2 * kessler_m + 1
+plot(0, 0, type = "n", xlim = c(0, 10), ylim = c(0, 1),
+     xlab = expression(Delta * T[air] ~ (K)),
+     ylab = "Frac. permafrost remaining")
+curve(1 / (1 + exp(m * x - b)), 0, 10, col = 1, add = TRUE)
+abline(a = kessler_b, b = -kessler_m, lty = "dashed")
+abline(h = 1, lty = "dotted")
+legend("topright", c("Hector", "Kessler (2015)"), lty = c("solid","dashed"),
+       bg = "white")
+```
+
+[kessler]: https://econpapers.repec.org/RePEc:lsg:lsgwps:wp219
 

### 2020-01-21 analysis/scratch.R

@@ -340,3 +340,20 @@ max(m, na.rm = TRUE)
 sum(m, na.rm = TRUE) / sum(!is.na(m))
 
 raster::plot(map1_r)
+
+##################################################
+permfunc <- function(pars) {
+  m <- pars[1]
+  b <- pars[2]
+  x <- seq(2, 6, 0.1)
+  yexp <- 1 / (1 + exp(m * x - b))
+  yline <- -0.172 * (x - 2) + 1
+  sum((yline - yexp) ^ 2)
+}
+fit <- optim(c(1.7, 8), permfunc)
+m <- fit$par[1]
+b <- fit$par[2]
+curve(1 / (1 + exp(m * x - b)), 0, 10, ylim = c(0, 1))
+abline(a = 2 * 0.172 + 1, b = -0.172)
+abline(v = c(2, 6), lty = "dashed")
+abline(h = 1, lty = "dashed")

### 2020-01-21 analysis/paper/old-sensitivity.Rmd
Move old sensitivity text out of main paper

--- 2019-09-13 analysis/paper/paper.Rmd
@@
----
-title: "Hector Permafrost"
-author:
-  - Alexey N. Shiklomanov
-  - Ben Bond-Lamberty
-  - Ben Kravitz
-  - Corinne Hartin
-date: "`r format(Sys.time(), '%d %B, %Y')`"
-output: 
-    bookdown::html_document2:
-      fig_caption: yes
-      reference_docx: "../templates/template.docx"
-    rmarkdown::github_document:
-      html_preview: false
-bibliography: ["references.bib", "packages.bib"]
-csl: "../templates/biogeosciences.csl"
-abstract: |
-  Text of abstract
-keywords: |
-  keyword 1; keyword 2; keyword 3
-highlights: |
-  These are the highlights. 
----
-
-```{r setup, echo = FALSE}
-library(drake)
-library(knitr)
-library(here)
-
-opts_chunk$set(
-  collapse = TRUE,
-  warning = FALSE,
-  message = FALSE,
-  echo = FALSE,
-  comment = "#>"
-)
-
-library(hector.permafrost.emit)
-
-fig <- function(...) {
-  filepath <- here("analysis", "figures", ...)
-  knitr::include_graphics(file_in(filepath))
-}
-
-n <- 0
-mdcap <- function(text) {
-  n <<- n + 1
-  out <- sprintf("**Figure %d**: %s", n, text)
-  if (grepl("gfm", knitr::opts_knit$get("rmarkdown.pandoc.to"), ignore.case = TRUE)) {
-    out
-  } else {
-    invisible(NULL)
-  }
-}
-
-m <- 0
-htmlfig <- function() {
-  m <<- m + 1
-  out <- paste0("**Figure ", m, "**:")
-  if (grepl("html", knitr::opts_knit$get("rmarkdown.pandoc.to"), ignore.case = TRUE)) {
-    out
-  } else {
-    invisible(NULL)
-  }
-}
-```
-
-# Introduction
-
-Climate change is a problem [@ipcc_ar5_wg1].
-Important to reduce carbon emissions to meet temperature targets.
-The extent of anthropogenic emissions reduction depends on what how much help (or push-back) we get from the physical and biological Earth systems.
-
-Permafrost is an important C reservoir.
-The total northern soil C pool is estimated at 1672 PgC, of which approximately 1466 Pg (88%) is in permafrost [@tarnocai_2009_soil].
-More recent estimates suggest that total high-latitude soil organic matter pool is 1300 Pg, with roughly 800 Pg in permafrost [@hugelius_2014_estimated].
-This frozen C can be released into the atmosphere by several processes related to warming, including aerobic respiration in thawed soil (REF) and anaeribic respiration in thermokarst lakes and wetlands [@turetsky_2002_boreal; @wickland_2006_effects].
-But emissions from wet soils may be offset by high organic matter accumulation rates [@camill_2001_changes].
-Permafrost thaw in boreal peatlands in north-central Saskatchewan increased CO2 and CH4 fluxes from soil to atmosphere by 1.6 and 30 times, respectively [@turetsky_2002_boreal].
-These impacts are exacerbated by the fact that the Arctic is warming roughly 2.5 (TODO ???) times faster than the global average (TODO REF).
-Projections of permafrost C emissions vary.
-[@schuur_2009_effect] estimate 0.8 - 1.1 Pg C yr-1.
-Back-of-the-envelope estimates from [@zimov_2006_climate_change]:
-10-40 g C m-3 day-1 off the bat, slowing down to equilibrium (?) rate of 0.5-5 g C m-3 day-1 for several years.
-
-Rapid methane release from arctic permafrost could potentially cause trillions in economic damage [@whiteman_2013_vast].
-
-On the other hand, warming and CO2 fertilization may increase vegetation productivity, which could increase soil C storage through enhanced litterfall; the balance of these two processes is uncertain [@jones_2005_global].
-There are additional uncertainties associated with vegetation composition (spatially variable peat accumulation) [@camill_2001_changes].
-Other large uncertainties related to soil properties and climate model structure [@harp_2016_effect].
-Several modeling studies generally predict increases in soil C sequestration at high latitudes [@burke_2017_quantifying; @qian_2010_enhanced; @ito_2016_impacts].
-However, this increase is dampened or even reversed when C emissions from permafrost thaw are included [@schaefer_2011_amount; @burke_2017_quantifying].
-
-The importance of Arctic climate feedbacks has led to efforts to incorporate these processes---particularly, permafrost thaw---into land surface models.
-Early efforts derived permafrost extent and C based purely on soil temperature [@lawrence_2005_projection].
-(More recent models have done this better...)
-@schaefer_2011_amount -- SiBCASA estimates 190 +/- 64 Gt C, but this does not include warming feedback or discontinuous permafrost regions.
-@harp_2016_effect -- Community Earth System Model (CESM) simulations over 100 years.
-[@burke_2017_quantifying] -- JULES and ORCHIDEE, with new permafrost scheme, combined with intermediate complexity climate-ocean model (IMOGEN), to look at climate sensitivity to permafrost C emissions.
-Models are highly sensitive to representation of soil processes, which can be more important than differences in scenario and/or climate drivers [@burke_2017_quantifying].
-But @harp_2016_effect argue that model structural uncertainty is larger than soil property uncertainty.
-
-Land surface models are expensive to run, making it challening to use them for uncertainty quantification and exploration of alternative policy scenarios.
-Simple climate models are an alternative.
-(More on simple climate models).
-
-(Previous attempts to incorporate permafrost into simple climate models.)
-
-Hector [@hartin_2015_simple].
-In this study, we use Hector to evaluate the global sensitivty of climate to terrestrial biosphere processes, and how this uncertainty is affected by permafrost emissions.
-The current version of Hector does not have an explicit representation of permafrost C emissions.
-In this study, we investigate whether the additional complexity and parametric and structural uncertainty of an explicit representation of permafrost may be warranted in Hector.
-To do this, we evaluate the sensitivity of climate variables (as predicted by Hector) to several different exogenous scenarios of permafrost C emissions.
-(TODO Modify this to be about economic impact)
-
-Some more relevant references:
-- [@zimov_2006_climate_change]
-- [@treat_2013_permafrost_carbon_bomb]
-- [@schuur_2015_climate; @schuur_2011_climate_change; @schaefer_2014_impact; @schaefer_2011_amount; @lee_2011_rate; @kuhry_2010_potential; @kessler_2017_estimating; @hope_2015_economic; @drake_2015_ancient; @burke_2017_quantifying]
-
-# Methods
-
-## Model description: Hector
-
-Hector [@hartin_2015_simple,@hartin_2016_ocean].
-Simple climate model.
-
-(TODO: More details on terrestrial C cycle in Hector).
-The default heterotrophic respiration ($R$) scheme for a pool $p$ (detritus or soil) in Hector:
-
-$$ R_{p} = C_p \times f_p \times Q_{10} ^ \frac{T}{10} $$
-
 ## Model configurations
@@
 **Figure**: PEcAn-like variance decomposition of parameters (for each parameter: sensitivity, uncertainty, and partial variance)
-
-# Discussion
-
-Rate of permafrost C release also depends on soil moisture conditions -- drier soils release C much faster ("carbon bomb") than wetter soils ("carbon fizz") [@elberling_2013_long].
-Moisture will also affect the balance of aerobic (CO2 release) vs.
-anaerobic (CH4) C release [@turetsky_2002_boreal], to the extent that unclear if anaerobic (wet) areas are C sources or sinks [@wickland_2006_effects].
-Effects of permafrost thaw on soil moisture are a complex hydrological problem -- drainage very sensitive to local (micro-)topography [@wickland_2006_effects].
-So will vegetation cover [@wickland_2006_effects].
-
-Temperature amplification of permafrost carbon feedback (by 2100) 0.02 to 0.36 °C [@burke_2013_estimating; @schneider_2012_estimating; @schneider_2015_observation], or 0.1 to 0.8 °C in [@macdougall_2012_significant; @macdougall_2013_if], or 10-40% of peak temperature change [@crichton_2016_permafrost], or 0.2 to 12% [@burke_2017_quantifying].
-
-Permafrost carbon has greater impact at lower emissions scenarios [@burke_2017_quantifying; @macdougall_2012_significant; @macdougall_2013_if] .
-
-# Conclusion
-
-# Acknowledgments
-
-Funded by EPA grant XXX.
-Cyberinfrastructure support from Pacific Northwest National Laboratory (PNNL).
-
-# Miscellaneous notes
-
-## Permafrost emissions scenarios
-
-Digitized scenarios from [@schaefer_2011_amount].
-SiBCASA model predictions of CO2 emissions (permafrost respiration; $R_{pc}$; note -- no methane!) through 2300.
-These results were digitized using WebPlotDigitizer (https://apps.automeris.io/wpd/), and interpolated to annual resolution (using R `stats::spline` function).
-
-Digitized scenarios from [@hope_2015_economic].
-CO2 and CH4 emissions from SiBCASA model.
-
-[@schuur_2009_effect] -- Estimate 0.8 - 1.1 Pg C yr-1.
-
-Back-of-the-envelope estimates from [@zimov_2006_climate_change]:
-- 500 Gt C in loess that could be completely emitted by 2100 (plus other C sources).
-- 10-40 g C m-3 day-1 off the bat, slowing down to equilibrium (?) rate of 0.5-5 g C m-3 day-1 for several years.
-Combine with data on permafrost spatial extent, density, etc.
-to generate estimates (but can back-calculate from 500 Gt C above?)
-
-## Parameter calibration
-
-We used the `BayesianTools` R package [@R-BayesianTools] for all parameter calibration.
-The outputs of these calibrations are joint posterior distributions of parameters and their covariances, from which we sample for the sensitivity analysis.
-
-For global parameters, we used the following likelihood:
-
-$$ \log(L) = \sum_s Normal(Hector(\beta, Q_{10}, s) | CMIP5(s), \sigma) $$
-
-where $s$ is one of the four representative carbon pathways (RCPs), $CMIP5(s)$ are the CMIP5 global mean outputs for the corresponding variables, and $\sigma$ is the model error (estimated during the fit).
-We also used the resulting distributions for $\beta$ and $Q_{10}$ for the non-permafrost biome in cases 2 and 3.
-We feel this is appropriate because the CMIP5 models against which these parameters are calibrated do not include permafrost C feedbacks.
-
-For case 2, we calibrated the permafrost-specific $\beta$ and $Q_{10}$ against various literature sources, including:
-
-- Land surface model simulations [@hope_2015_economic; @schaefer_2011_amount; @burke_2017_quantifying].
-    - Try to calibrate against NPP and soil respiration if possible
-- Literature surveys [@schaefer_2014_impact] 
-- Warming experiments [@wickland_2006_effects]
-- TODO: Others?
-
-Some of these are time series, while others are individual estimates at particular points in time.
-To give them equal weight in the likelihood, we down-weight the time series likelihoods by the number of time steps.
-
-We derived a distribution for the Arctic warming factor from TODO.
-
-TODO: Table and multi-panel figure of input datasets.
-
-For the $\alpha$ and $\phi$ parameters in case 3, we looked at the literature on permafrost methane emissions [e.g., @wickland_2006_effects].
-
-## Other notes
-
-Frozen carbon residence time (FCRt) from [@burke_2017_quantifying]:
-
-$$ FCRt = FCRt0 * exp(-\Delta T / \Gamma) (for \Delta T > 0.2 °C) $$
-
-- $FCRt_0$ (years) reflects the stability of permafrost C (length of time that permafrost C is stable when $\Delta T = 0$)
-- $\Gamma$ -- decay term (°C); temperature change at which "the number of years taken for all of the old permafrost C to be emitted reduces to 1/e of its initial value"
-
-Other Hector parameters to consider.
-
-| Variable | INI name | Description                                 | Value |
-|----------|----------|---------------------------------------------|-------|
-| $f_{nv}$ | `f_nppv` | Fraction of NPP C transferred to vegetation | 0.35  |
-| $f_{nd}$ | `f_nppd` | Fraction of NPP C transferred to detritus   | 0.60  |
-| $f_{nd}$ |          | Fraction of NPP C transferred to soil       | 0.05  |
-| $f_{lv}$ | `f_lucv` | Fraction of LUC change flux from vegetation | 0.10  |
-| $f_{ld}$ | `f_lucd` | Fraction of LUC change flux from detritus   | 0.01  |
-| $f_{ls}$ |          | Fraction of LUC change flux from soil       | 0.89  |
-| $f_{ds}$ |          | Fraction of detritus C that goes to soil    | 0.60  |
-| $f_{rd}$ |          | Fraction of respiration C to detritus       | 0.25  |
-| $f_{rs}$ |          | Fraction of respriation C to soil           | 0.02  |
-
-According to [@hartin_2015_simple], these were selected to be "generally consistent with previous simple earth system models [e.g., @meinshausen_2011_emulating]".
-
-<!-- The following line inserts a page break when the output is MS Word. For page breaks in PDF, use \newpage on its own line.  -->
-##### pagebreak
-
-# References 
-<!-- The following line ensures the references appear here for the MS Word or HTML output files, rather than right at the end of the document (this will not work for PDF files):  -->
-<div id="refs"></div>
-
-##### pagebreak
-
-# Appendix
-
-```{r fig-global-dist, fig.cap = global_dist_cap}
-global_dist_cap <- paste(
-  htmlfig(),
-  "Input parameter distributions for global Hector."
-)
-fig("global_draws.png")
-```
-
-`r mdcap(global_dist_cap)`
-
-```{r fig-biome-dist, fig.cap = biome_dist_cap, fig.dim = c(9, 9)}
-biome_dist_cap <- paste(
-  htmlfig(),
-  "Input parameter distributions for Hector with biomes."
-)
-fig("biome_draws.png")
-```
-
-`r mdcap(biome_dist_cap)`
-
-```{r fig-global-timeseries, fig.cap = global_ts_cap, fig.dim = c(10, 8)}
-global_ts_cap <- paste(
-  htmlfig(),
-  "Time series of Hector outputs for different parameter combinations."
-)
-readd(ts_global_sims)
-```
-
-`r mdcap(global_ts_cap)`
-
-```{r fig-biome-timeseries, fig.cap = biome_ts_cap, fig.dim = c(10, 8)}
-biome_ts_cap <- paste(
-  htmlfig(),
-  "Overall parameter uncertainty of Hector simulations with permafrost biome."
-)
-readd(ts_biome_sims)
-```
-
-`r mdcap(biome_ts_cap)`
-
-### Colophon
-
-This report was generated on `r Sys.time()` using the following computational environment and dependencies:
-
-```{r colophon, cache = FALSE}
-# which R packages and versions?
-devtools::session_info()
-```
-
-The current Git commit details are:
-
-```{r}
-# what commit is this file at? You may need to change the path value
-# if your Rmd is not in analysis/paper/
-git2r::repository("../..")
-```

### 2020-01-21 analysis/paper/paper.Rmd
Move old sensitivity text out of main paper

@@
 
-```{r sigmoid, echo = FALSE}
+```{r sigmoid, fig.cap = sigmoid_cap}
 sigmoid_cap <- paste(
@@
 
-## Model configurations
+## Configuration
 
-We evaluate the sensitivity and parametric uncertainty of three different versions of Hector.
-
-The simplest is the "global" version, which corresponds to the standard version of Hector described in [@hartin_2015_simple].
-The global land carbon sink is modeled as a single entity, with parameterizations corresponding to global averages (originally tuned to match global outputs from CMIP5; TODO REF?).
-In this configuration, we vary the following parameters:
-CO~2~ fertilization effect ($\beta$);
-temperature sensitivity of heterotrophic respiration, ($Q^{10}_{RH}$);
-fraction of net primary production C that goes to vegetation (`f_nppv`), detritus (`f_nppd`), and soil (`1 - f_nppv - f_nppd`);
-and
-fraction of vegetation litter C that goes to detritus (`f_litterd`).
-
-The second is the "biome" version, which is the same as the "global" version except that the land carbon sink is divided into two "biomes": "non-permafrost" and "permafrost".
-Each of these biomes has its own C pools, fluxes, and parameters.
-For this configuration, we varied all of the "global" parameters (previous paragraph) for _both_ biomes ($\beta_{permafrost}$, $\beta_{non-permafrost}$, $Q^10_{RH,permafrost}$, $Q^10_{RH,non-permafrost}$, etc.), as well as the distribution of global pools (vegetation, detritus, soil) and initial primary productivity (`npp_flux0`) across these two biomes and the relative warming factor (compared to the global average) of the permafrost biome.
-
-The third is the "biome + methane" version.
-This version is the same as the "biome" version, but with an additional process that partitions total heterotrophic respiration ($R_{p,tot}$) into a CO~2~ flux ($R_{p, CO_2}$) and a CH~4~ flux ($R_p, CH_4$) as follows:
-
-$$ R_{p,tot} = C_p \times f_p \times Q_{10} ^ \frac{T}{10} $$
-
-$$ R_{p, CH_4} = \alpha R_{p, tot} $$
-
-$$ R_{p, CO_2} = (1 - \alpha) ^ \phi R_{p, tot} $$
-
-This process is controlled by two new biome-specific parameters:
-$\alpha$ can be interpreted as the fraction of heterotrophic respiration C that is given off as CH~4~ assuming a perfect trade-off between CO~2~ and CH~4~ emissions.
-$\phi$ defines the plasticity of this relationship;
-at $\phi$ = 1, plasticity is perfect, such that $R_{p,tot}$ is evenly split between CO~2~ and CH~4~;
-at $\phi$ = 0, CO~2~ emissions are entirely independent of CH~4~ emissions, such that total C emissions increase linearly with $\alpha$;
-at $\phi$ > 1, CO~2~ emissions decline faster than CH~4~ emissions increase.
-As such, the total C emissions attributable to heterotrophic respiration are expected to decrease as $\phi$ increases.
-In this version of Hector, we varied all of the parameters in the "biome" verson, as well as $\alpha$ and $\phi$ for each biome.
-
-## Parameter distributions
-
-We drew parameters from loosely informative distributions that span the range of physical plausibility.
-The distributions are shown in the following table (unless otherwise specified, we used the same parameter distributions for global and biome cases):
-
-| Parameter                        | Distribution                    | Hector default value |
-|----------------------------------|---------------------------------|----------------------|
-| $\beta$                          | Uniform(0, 1)                   | 0.36                 |
-| Q^10^~RH~                        | Uniform(0, 10)                  | 2.0                  |
-| Litter-detritus fraction         | Beta(3.92, 0.08)                | 0.98                 |
-| NPP vegetation fraction          | Dirichlet(**0.35**, 0.60, 0.05) | 0.35                 |
-| NPP detritus fraction            | Dirichlet(0.35, **0.60**, 0.05) | 0.60                 |
-|----------------------------------|---------------------------------|----------------------|
-| Permafrost vegetation C fraction | Beta(1, 2)                      | --                   |
-| Permafrost soil C fraction       | Beta(1.1, 1.1)                  | --                   |
-| Permafrost detritus C fraction   | Beta(1.1, 1.1)                  | --                   |
-| Permafrost warming factor        | 1 + 3 * Beta(5, 5)              | --                   |
-|----------------------------------|---------------------------------|----------------------|
-| Non-permafrost $\alpha$          | TODO                            | --                   |
-| Non-permafrost $\phi$            | TODO                            |                      |
-| Permafrost $\alpha$              | TODO                            |                      |
-| Permafrost $\phi$                | TODO                            |                      |
-
-These distributions are also shown in Figure 1.
-
-```{r param-density, echo = FALSE, fig.cap = param_density_cap}
-param_density_cap <- paste(
-  htmlfig(),
-  "Input parameter distributions for this analysis."
-)
-fig("param-density.png")
-```
-
-`r mdcap(param_density_cap)`
-
-Actual permafrost surface C fraction is around 30% [@schuur_2015_climate].
-
-We used 5000 Hector simulations for the global case and 10,000 times for the biome case (biome + methane case is TODO).
-
-## Analysis of results
-
-We looked at four output variables:
-Atmospheric CO~2~ concentration (`Ca`),
-total radiative forcing (`Ftot`),
-radiative forcing from CO~2~ alone (`FCO2`),
-and
-global mean atmospheric temperature (`Tgav`).
-
-First, as a metric of overall parametric uncertainty for each case, we calculated the mean, standard deviation, and 90% and 95% quantiles of the time series for each variable for each simulation.
-Second, we evaluated the impact of parameters by plotting the values of each variable at 2100 as a function of input parameter value.
-
-Third, we performed a more rigorous sensitivity and uncertainty analysis based on @lebauer_2013_facilitating.
-As a metric of parameter uncertainty, we calculated the coefficient of variation (`cv`) for each input parameter as the ratio of its variance to its mean.
-We evaluated sensitivity as follows:
-First, we fit a multivariate Generalized Additive Model (GAM; `mgcv::gam` function in R) for each variable as a function of all the input parameters.
-Then, we calculate the sensitivity ($s$) to each parameter ($x_i$) as the slope (partial derivative) of the resulting GAM at that parameter's median ($\hat{x}$; discretized at +/- 1% of the median), holding all other parameters ($x_{i+1}..x_n$) constant:
-
-$$ s = \frac{d(GAM(x_i | x_{i + 1} .. x_n))}{d(x_i)} = \frac{GAM(1.01 x_i | ...) - GAM(0.99 x_i | ...)}{1.01 x_i - 0.99 x_i}$$
-
-We then define the "elasticity" ($\epsilon$) (or normalized sensitivity) of the model to that parameter as:
-
-$$ \epsilon = s \div \frac{\hat{y}} {\hat{x}} $$
-
-Finally, we calculate the "partial variance" of each parameter (the fractional contribution of each parameter's uncertainty to overall model predictive uncertainty) by passing the distribution of each parameter through the GAM while holding all other parameters constant, calculating the resulting predictive variance for each parameter, and then normalizing these values relative to the sum of all the variances:
-
-$$ Var(x_i) \approx Var[GAM(x_i | x_{i+1} .. x_n)] $$
-
-$$ pvar(x_i) = \frac{Var(x_i)}{\sum_{i}^n {Var(x_i)}} $$
+Initial permafrost C is set to 1035 Pg C.
 
@@
 
-```{r fig-summary-ts, fig.cap = summary_ts_cap}
-summary_ts_cap <- paste(
-  htmlfig(),
-  "Mean and 95% confidence interval on Hector simulations ",
-  "with different model structures and parameter combinations."
+## Effect of permafrost C
+
+```{r permafrost-c, fig.cap = }
+permafrost_c_cap <- paste(
+  "Effect of permafrost C emissions on scenarios."
 )
-fig("ts-summary.png")
+
+devtools::load_all("~/Projects/hector_project/hector-permafrost-submodel")
+run_rcp <- function(rcp) {
+  inidir <- file.path("~", "Projects", "hector_project",
+                      "hector-permafrost-submodel", "inst", "input")
+  inifile <- normalizePath(file.path(inidir, paste0("hector_rcp", rcp, ".ini")))
+  stopifnot(file.exists(inifile))
+  hc <- newcore(inifile)
+  run(hc)
+  dates <- seq(1750, 2100)
+  outvars <- c(ATMOSPHERIC_CO2(), GLOBAL_TEMP())
+  no_pf <- fetchvars(hc, dates, outvars, scenario = "Original")
+  reset(hc)
+  setvar(hc, NA, PERMAFROST_C(), 1035, "PgC")
+  run(hc)
+  yes_pf <- fetchvars(hc, dates, outvars, scenario = "Permafrost")
+  return(tibble::as_tibble(dplyr::bind_rows(no_pf, yes_pf)))
+}
+r26 <- run_rcp("26")
+library(ggplot2)
+ggplot(r26) +
+  aes(x = year, y = value, color = scenario) +
+  geom_line() +
+  facet_wrap(vars(variable), scales = "free_y")
+
 ```
-
-`r mdcap(summary_ts_cap)`
-
-## Global case
-
-```{r fig-global-scatter, fig.cap = global_scatter_cap}
-global_scatter_cap <- paste(
-  htmlfig(),
-  "Hector outputs at 2100 as a function of input parameter values."
-)
-fig("global-scatter.png")
-```
-
-`r mdcap(global_scatter_cap)`
-
-```{r fig-global-sensitivity, fig.cap = global_sensitivity_cap, fig.dim = c(5, 5)}
-global_sensitivity_cap <- paste(
-  htmlfig(),
-  "Sensitivity and variance decomposition analysis for global Hector."
-)
-fig("sensitivity-global.png")
-```
-
-`r mdcap(global_sensitivity_cap)`
-
-## Biome case
-
-```{r fig-biome-scatter, fig.cap = biome_scatter_cap}
-biome_scatter_cap <- paste(
-  htmlfig(),
-  "Hector outputs (with permafrost biome) at 2100 as a function of input parameter values."
-)
-fig("biome-scatter.png")
-```
-
-`r mdcap(biome_scatter_cap)`
-
-```{r fig-biome-sensitivity, fig.cap = biome_sensitivity_cap}
-biome_sensitivity_cap <- paste(
-  htmlfig(),
-  "Sensitivity and variance decomposition analysis for Hector with permafrost biome."
-)
-fig("sensitivity-biome.png")
-```
-
-`r mdcap(biome_sensitivity_cap)`
-
-## Permafrost methane emissions
-
-**Figure**: Hector projections of parameter sensitivity (CI ribbon, or light/transparent lines), colored (faceted?) by case.
-
-**Figure**: PEcAn-like variance decomposition of parameters (for each parameter: sensitivity, uncertainty, and partial variance)
 

### 2020-01-21 analysis/scratch.R

@@ -357,3 +357,27 @@ curve(1 / (1 + exp(m * x - b)), 0, 10, ylim = c(0, 1))
 abline(a = 2 * 0.172 + 1, b = -0.172)
 abline(v = c(2, 6), lty = "dashed")
 abline(h = 1, lty = "dashed")
+
+devtools::load_all("~/Projects/hector_project/hector-permafrost-submodel")
+run_rcp <- function(rcp) {
+  inidir <- file.path("~", "Projects", "hector_project",
+                      "hector-permafrost-submodel", "inst", "input")
+  inifile <- normalizePath(file.path(inidir, paste0("hector_rcp", rcp, ".ini")))
+  stopifnot(file.exists(inifile))
+  hc <- newcore(inifile)
+  run(hc)
+  dates <- seq(1750, 2100)
+  outvars <- c(ATMOSPHERIC_CO2(), GLOBAL_TEMP())
+  no_pf <- fetchvars(hc, dates, outvars, scenario = "Original")
+  reset(hc)
+  setvar(hc, NA, PERMAFROST_C(), 1035, "PgC")
+  run(hc)
+  yes_pf <- fetchvars(hc, dates, outvars, scenario = "Permafrost")
+  return(tibble::as_tibble(dplyr::bind_rows(no_pf, yes_pf)))
+}
+r26 <- run_rcp("26")
+library(ggplot2)
+ggplot(r26) +
+  aes(x = year, y = value, color = scenario) +
+  geom_line() +
+  facet_wrap(vars(variable), scales = "free_y")

### 2020-01-21 analysis/scratch/debug.R

@@ -1,0 +1,26 @@
+devtools::load_all("~/Projects/hector_project/hector-permafrost-submodel")
+
+rcp <- "26"
+inidir <- file.path("~", "Projects", "hector_project",
+                    "hector-permafrost-submodel", "inst", "input")
+inifile <- normalizePath(file.path(inidir, paste0("hector_rcp", rcp, ".ini")))
+stopifnot(file.exists(inifile))
+hc <- newcore(inifile)
+run(hc)
+dates <- seq(1750, 2100)
+outvars <- c(ATMOSPHERIC_CO2(), GLOBAL_TEMP())
+no_pf <- fetchvars(hc, dates, outvars, scenario = "Original")
+idate <- 1745
+reset(hc, idate)
+setvar(hc, idate, PERMAFROST_C(), 1035, "PgC")
+fetchvars(hc, idate, PERMAFROST_C())
+run(hc)
+fetchvars(hc, idate, PERMAFROST_C())
+yes_pf <- fetchvars(hc, dates, outvars, scenario = "Permafrost")
+r26 <- rbind(no_pf, yes_pf)
+
+library(ggplot2)
+ggplot(r26) +
+  aes(x = year, y = value, color = scenario) +
+  geom_line() +
+  facet_wrap(vars(variable), scales = "free_y")

### 2020-01-23 analysis/paper/paper.Rmd
Update permafrost submodel documentation
Now, it reflects what Hector actually does.

@@
 library(here)
+library(dplyr)
+library(ggplot2)
+library(hector.permafrost.emit)
 
@@
 
-library(hector.permafrost.emit)
+devtools::load_all("~/Projects/hector_project/hector-permafrost-submodel")
 
@@
 As permafrost thaws and some fraction of this carbon becomes accessible to microbes, it is transferred from the frozen permafrost pool to the standard soil C pool, where it is decomposed by Hector's standard decomposition routine.
-Let $C[t]_{pf}$ be the permafrost C pool and $C[t]_{s}$ be the soil C pool at time $t$ (both in units of Pg C), and let $\Delta C[t]_{pf}$ be the change in the permafrost C pool at time $t$.
+Let $C_{pf}[t]$ be the permafrost C pool and $C_{s}[t]$ be the soil C pool at time $t$ (both in units of Pg C), and let $\Delta C_{pf}[t]$ be the change in the permafrost C pool at time $t$.
 The C consequences of permafrost thaw can therefore be represented as:
 
-\[ C[t]_{pf} = C[t-1]_{pf} - \Delta C[t]_{pf} \]
+\[ C_{pf}[t] = C_{pf}[t-1] - \Delta C_{pf}[t] \]
 
-\[ C[t]_{s} = C[t]_{s} + \Delta C[t]_{pf} \]
+\[ C_{s}[t] = C_{s}[t] + \Delta C_{pf}[t] \]
 
 Let $C[0]_{pf}$ be the initial size of the permafrost C pool (Pg C) and $\Phi[t]$ be the fraction of permafrost remaining (in arbitrary area or volume units) at time $t$.
-Assuming a uniform permafrost C density, $\Delta C[t]_{pf}$ can be expressed as:
+Assuming a uniform permafrost C density, $\Delta C_{pf}[t]$ can be expressed as:
 
-\[ \Delta C[t]_{pf} = \Phi[t] C[0]_{pf} \]
+\[ \Delta C_{pf}[t] = \Delta \Phi[t] C_{pf}[t-1] \]
 
@@
 \[
-\Phi[t] = \frac{1}{1 + \exp\left( m T[t]_{air} + b \right)}
+\Phi[t] = 1 - \left( 1 + a \exp\left( b * T_{air}[t] \right) \right) ^ c
 \]
 
-where $m$ and $b$ are model parameters.
-For globally-averaged permafrost, $m = 0.829$ and $b = 4.078$ most closely reproduce the rate of 0.172 K$^{-1}$ reported by [Kessler (2015)] over the range of 2 to 6 K$^{-1}$.
+where $a$, $b$, and $c$ are model parameters.
+The change in frozen fraction at a given time step, $\Delta \Phi[t]$, is given by:
+
+\[
+\Delta \Phi[t] = \max(\Phi[t] - \Phi[t-1], 0)
+\]
+
+In other words, permafrost thaw is permanent -- once it thaws, it does not re-freeze, even if the temperature drops.
+
+For globally-averaged permafrost, $a = 2.371$, $b = -0.676$, and $c = -3.685$ to most closely reproduce the rate of 0.172 K$^{-1}$ reported by [Kessler (2015)][kessler] over the range of 0.8 [^2] to 4 K above the pre-industrial baseline.
+
+[^2]: [Kessler (2015)][kessler] report this as temperature change from year 2000. 0.8 K is the warming since pre-industrial as estimated by the default Hector configuration.
 
 ```{r sigmoid, fig.cap = sigmoid_cap}
+
 sigmoid_cap <- paste(
@@
 )
-m <- 0.829
+pfa <- 2.371
+pfb <- -0.676
+pfc <- -3.685
 b <- 4.078
 kessler_m <- 0.172
-kessler_b <- 2 * kessler_m + 1
-plot(0, 0, type = "n", xlim = c(0, 10), ylim = c(0, 1),
+kessler_b <- 0.8 * kessler_m + 1
+plot(0, 0, type = "n", xlim = c(0, 8), ylim = c(0, 1),
      xlab = expression(Delta * T[air] ~ (K)),
      ylab = "Frac. permafrost remaining")
-curve(1 / (1 + exp(m * x - b)), 0, 10, col = 1, add = TRUE)
+curve(1 - (1 + pfa * exp(pfb * x)) ^ pfc, 0, 10, col = 1, add = TRUE)
 abline(a = kessler_b, b = -kessler_m, lty = "dashed")
@@
        bg = "white")
+
 ```
@@
 
-```{r permafrost-c, fig.cap = }
+```{r permafrost-c, fig.cap = permafrost_c_cap}
+
 permafrost_c_cap <- paste(
@@
 
-devtools::load_all("~/Projects/hector_project/hector-permafrost-submodel")
 run_rcp <- function(rcp) {
@@
   hc <- newcore(inifile)
-  run(hc)
   dates <- seq(1750, 2100)
-  outvars <- c(ATMOSPHERIC_CO2(), GLOBAL_TEMP())
+  run(hc, max(dates))
+  outvars <- c(ATMOSPHERIC_CO2(), GLOBAL_TEMP(), PERMAFROST_C(), SOIL_C())
   no_pf <- fetchvars(hc, dates, outvars, scenario = "Original")
   reset(hc)
-  setvar(hc, NA, PERMAFROST_C(), 1035, "PgC")
-  run(hc)
+  setvar(hc, 0, PERMAFROST_C(), 1035, "PgC")
+  run(hc, max(dates))
   yes_pf <- fetchvars(hc, dates, outvars, scenario = "Permafrost")
@@
 }
-r26 <- run_rcp("26")
-library(ggplot2)
-ggplot(r26) +
-  aes(x = year, y = value, color = scenario) +
+
+rcps <- c("26", "45", "60", "85")
+names(rcps) <- paste0("RCP", rcps)
+results <- purrr::map_dfr(rcps, run_rcp, .id = "RCP")
+
+results %>%
+  mutate(variable = recode(
+    variable,
+    Ca = "Atmospheric CO2 (ppm)",
+    Tgav = "Global mean temperature anomaly (K)",
+    permafrost_c = "Permafrost C pool (PgC)",
+    soil_c = "Soil C pool (PgC)"
+  )) %>%
+  ggplot() +
+  aes(x = year, y = value, linetype = scenario, color = RCP) +
   geom_line() +
-  facet_wrap(vars(variable), scales = "free_y")
+  facet_wrap(vars(variable), scales = "free_y") +
+  theme_bw() +
+  theme(axis.title = element_blank())
 

### 2020-01-23 analysis/scratch.R
Random tweaks to scratch and debug code

@@ -342,21 +342,76 @@ sum(m, na.rm = TRUE) / sum(!is.na(m))
 raster::plot(map1_r)
 
 ##################################################
+t2000 <- 0.812
+tmax <- 6
+kessler_line <- function(x) -0.172 * (x - t2000) + 1
+logistic <- function(x, pars) {
+  r <- pars[1]
+  p0 <- 0.01
+  if (length(pars) == 2) p0 <- pars[2]
+  k <- 1
+  n <- k * p0 * exp(x * r)
+  d <- k + p0 * (exp(x * r) - 1)
+  n / d
+}
+permfunc <- function(pars) {
+  x <- seq(t2000, tmax, 0.1)
+  yexp <- 1 - logistic(x, pars)
+  yline <- kessler_line(x)
+  sum((yline - yexp) ^ 2 * (1 / x^2))
+}
+fit <- optim(c(r = 1), permfunc)
+plot(0, 0, type = "n", xlim = c(0, 10), ylim = c(0, 1))
+curve(kessler_line(x), 0, 10, add = TRUE, lty = "dashed")
+curve(1 - logistic(x, fit$par), 0, 10, add = TRUE, col = "red")
+abline(h = 1.0, lty = "dotted")
+fit$par
+
+curve(1 - gen_log(x, c(0.037, 0.91, 1)), 0, 10)
+curve(gen_log(x, c(0.037, 0.91, 5)), 0, 10, add = TRUE, col = 2)
+curve(gen_log(x, c(0.037, 0.91, 10)), 0, 10, add = TRUE, col = 3)
+
+curve(1 - gen_log(x, c(1, -1, 1)), 0, 10)
+curve(gen_log(x, c(0.037, 0.91, 5)), 0, 10, add = TRUE, col = 2)
+curve(gen_log(x, c(0.037, 0.91, 10)), 0, 10, add = TRUE, col = 3)
+
+
 permfunc <- function(pars) {
   m <- pars[1]
   b <- pars[2]
-  x <- seq(2, 6, 0.1)
+  x <- seq(t2000, tmax, 0.1)
   yexp <- 1 / (1 + exp(m * x - b))
-  yline <- -0.172 * (x - 2) + 1
+  yline <- kessler_line(x)
   sum((yline - yexp) ^ 2)
 }
 fit <- optim(c(1.7, 8), permfunc)
 m <- fit$par[1]
 b <- fit$par[2]
[…]

### 2020-01-23 analysis/scratch/debug.R

@@ -1,21 +1,18 @@
 devtools::load_all("~/Projects/hector_project/hector-permafrost-submodel")
-
-rcp <- "26"
 inidir <- file.path("~", "Projects", "hector_project",
                     "hector-permafrost-submodel", "inst", "input")
+rcp <- "45"
 inifile <- normalizePath(file.path(inidir, paste0("hector_rcp", rcp, ".ini")))
 stopifnot(file.exists(inifile))
 hc <- newcore(inifile)
-run(hc)
-dates <- seq(1750, 2100)
-outvars <- c(ATMOSPHERIC_CO2(), GLOBAL_TEMP())
+yearz <- 2100
+run(hc, yearz)
+dates <- seq(1750, yearz)
+outvars <- c(ATMOSPHERIC_CO2(), GLOBAL_TEMP(), PERMAFROST_C(), SOIL_C())
 no_pf <- fetchvars(hc, dates, outvars, scenario = "Original")
-idate <- 1745
-reset(hc, idate)
-setvar(hc, idate, PERMAFROST_C(), 1035, "PgC")
-fetchvars(hc, idate, PERMAFROST_C())
-run(hc)
-fetchvars(hc, idate, PERMAFROST_C())
+reset(hc)
+setvar(hc, 0, PERMAFROST_C(), 1035, "PgC")
+run(hc, yearz)
 yes_pf <- fetchvars(hc, dates, outvars, scenario = "Permafrost")
 r26 <- rbind(no_pf, yes_pf)
 

### 2020-01-23 analysis/paper/paper.Rmd
Show permafrost frozen fraction output

@@
   run(hc, max(dates))
-  outvars <- c(ATMOSPHERIC_CO2(), GLOBAL_TEMP(), PERMAFROST_C(), SOIL_C())
+  outvars <- c(ATMOSPHERIC_CO2(), GLOBAL_TEMP(), "f_frozen", SOIL_C())
   no_pf <- fetchvars(hc, dates, outvars, scenario = "Original")
@@
 results %>%
-  mutate(variable = recode(
+  mutate(variable = recode_factor(
     variable,
+    Tgav = "Global mean temperature anomaly (K)",
     Ca = "Atmospheric CO2 (ppm)",
-    Tgav = "Global mean temperature anomaly (K)",
+    soil_c = "Active soil C pool (PgC)",
     permafrost_c = "Permafrost C pool (PgC)",
-    soil_c = "Soil C pool (PgC)"
+    f_frozen = "Frozen permafrost fraction"
   )) %>%

### 2020-01-23 analysis/paper/paper.Rmd
Proper identifer for F_FROZEN

@@
   run(hc, max(dates))
-  outvars <- c(ATMOSPHERIC_CO2(), GLOBAL_TEMP(), "f_frozen", SOIL_C())
+  outvars <- c(ATMOSPHERIC_CO2(), GLOBAL_TEMP(), F_FROZEN(), SOIL_C())
   no_pf <- fetchvars(hc, dates, outvars, scenario = "Original")

### 2020-02-05 analysis/scratch/gruber.R
Scratch code for Gruber permafrost area

@@ -1,0 +1,10 @@
+# Based on:
+# Gruber, S. (2012). Derivation and analysis of a high-resolution estimate of
+# global permafrost zonation. The Cryosphere, 6(1), 221–233.
+# http://dx.doi.org/https://doi.org/10.5194/tc-6-221-2012
+ 
+mu <- 4.38
+sig2 <- 6.43
+curve(1 - pnorm(x, -mu, sqrt(sig2)), -12, 2,
+      xlab = expression(MAAT ~ (degree * C)),
+      ylab = "Permafrost area fraction")

### 2020-02-05 analysis/paper/paper.Rmd
New first intro paragraph

@@
 # Introduction
+
+Permafrost---soil that continuously remains below 0°C for at least two consecutive years---underlies an area of 22 (± 3) million km~2~, roughly 17% of the Earth's exposed land surface [@gruber_2012_derivation], and contains an estimated 1300 (1100-1500) Pg of organic carbon [@hugelius_2014_estimated].
+Recent increases in global air temperature [@ipcc_ar5_wg1], which are amplified at high latitudes [@pithan_2014_arctic], have resulted in widespread permafrost thaw [@romanovsky_2010_permafrost], and simulations from variety of climate and land surface models across a wide range of scenarios suggest that this trend will continue into the future [@koven_2013_analysis; @chadburn_2017_observation].
+Permafrost thaw can dramatically alter surface terrain and hydrology [@jones_2013_quantifying; @godin_2014_effects; @necsoiu_2013_multitemporal], with adverse consequences for human infrastructure in permafrost regions [@anisimov_1997_permafrost; @nelson_2002_climate; @larsen_2008_estimating].
+Moreover, as permafrost thaws, this carbon becomes available to microbes for decomposition, resulting in the production of CO~2~ and CH~4~ [@brown_2008_report; @romanovsky_2010_thermal; @bond-lamberty_2016_temperature] that could lead to further warming [@koven_2011_permafrost; @schuur_2015_climate].
+Accounting for this permafrost carbon-climate feedback generally increases projections of greenhouse gas concentrations and global temperatures (REFS) and increases estimates of the economic impact of climate change [@hope_2015_economic; @yumashev_2019_climate; @chen_2019_economic].
 

### 2020-02-05 analysis/paper/old-sensitivity.Rmd
More revisions to introduction

@@
+# Introduction
+
+Climate change is a problem [@ipcc_ar5_wg1].
+Important to reduce carbon emissions to meet temperature targets.
+The extent of anthropogenic emissions reduction depends on what how much help (or push-back) we get from the physical and biological Earth systems.
+
+Permafrost is an important C reservoir.
+The total northern soil C pool is estimated at 1672 PgC, of which approximately 1466 Pg (88%) is in permafrost [@tarnocai_2009_soil].
+More recent estimates suggest that total high-latitude soil organic matter pool is 1300 Pg, with roughly 800 Pg in permafrost [@hugelius_2014_estimated].
+This frozen C can be released into the atmosphere by several processes related to warming, including aerobic respiration in thawed soil (REF) and anaeribic respiration in thermokarst lakes and wetlands [@turetsky_2002_boreal; @wickland_2006_effects].
+But emissions from wet soils may be offset by high organic matter accumulation rates [@camill_2001_changes].
+Permafrost thaw in boreal peatlands in north-central Saskatchewan increased CO2 and CH4 fluxes from soil to atmosphere by 1.6 and 30 times, respectively [@turetsky_2002_boreal].
+These impacts are exacerbated by the fact that the Arctic is warming roughly 2.5 (TODO ???) times faster than the global average (TODO REF).
+Projections of permafrost C emissions vary.
+[@schuur_2009_effect] estimate 0.8 - 1.1 Pg C yr-1.
+Back-of-the-envelope estimates from [@zimov_2006_climate_change]:
+10-40 g C m-3 day-1 off the bat, slowing down to equilibrium (?) rate of 0.5-5 g C m-3 day-1 for several years.
+
+Rapid methane release from arctic permafrost could potentially cause trillions in economic damage [@whiteman_2013_vast].
+
+On the other hand, warming and CO2 fertilization may increase vegetation productivity, which could increase soil C storage through enhanced litterfall; the balance of these two processes is uncertain [@jones_2005_global].
+There are additional uncertainties associated with vegetation composition (spatially variable peat accumulation) [@camill_2001_changes].
+Other large uncertainties related to soil properties and climate model structure [@harp_2016_effect].
+Several modeling studies generally predict increases in soil C sequestration at high latitudes [@burke_2017_quantifying; @qian_2010_enhanced; @ito_2016_impacts].
+However, this increase is dampened or even reversed when C emissions from permafrost thaw are included [@schaefer_2011_amount; @burke_2017_quantifying].
+
+The importance of Arctic climate feedbacks has led to efforts to incorporate these processes---particularly, permafrost thaw---into land surface models.
+Early efforts derived permafrost extent and C based purely on soil temperature [@lawrence_2005_projection].
+(More recent models have done this better...)
+@schaefer_2011_amount -- SiBCASA estimates 190 +/- 64 Gt C, but this does not include warming feedback or discontinuous permafrost regions.
+@harp_2016_effect -- Community Earth System Model (CESM) simulations over 100 years.
+[@burke_2017_quantifying] -- JULES and ORCHIDEE, with new permafrost scheme, combined with intermediate complexity climate-ocean model (IMOGEN), to look at climate sensitivity to permafrost C emissions.
+Models are highly sensitive to representation of soil processes, which can be more important than differences in scenario and/or climate drivers [@burke_2017_quantifying].
+But @harp_2016_effect argue that model structural uncertainty is larger than soil property uncertainty.
+
+Land surface models are expensive to run, making it challening to use them for uncertainty quantification and exploration of alternative policy scenarios.
+Simple climate models are an alternative.
+(More on simple climate models).
+
+(Previous attempts to incorporate permafrost into simple climate models.)
+
+Hector [@hartin_2015_simple].
+In this study, we use Hector to evaluate the global sensitivty of climate to terrestrial biosphere processes, and how this uncertainty is affected by permafrost emissions.
+The current version of Hector does not have an explicit representation of permafrost C emissions.
+In this study, we investigate whether the additional complexity and parametric and structural uncertainty of an explicit representation of permafrost may be warranted in Hector.
+To do this, we evaluate the sensitivity of climate variables (as predicted by Hector) to several different exogenous scenarios of permafrost C emissions.
+(TODO Modify this to be about economic impact)
+
+Some more relevant references:
+- [@zimov_2006_climate_change]
+- [@treat_2013_permafrost_carbon_bomb]
+- [@schuur_2015_climate; @schuur_2011_climate_change; @schaefer_2014_impact; @schaefer_2011_amount; @lee_2011_rate; @kuhry_2010_potential; @kessler_2017_estimating; @hope_2015_economic; @drake_2015_ancient; @burke_2017_quantifying]
+
+# Methods
+
 ## Model configurations

### 2020-02-05 analysis/paper/paper.Rmd
More revisions to introduction

@@
 
-Climate change is a problem [@ipcc_ar5_wg1].
-Important to reduce carbon emissions to meet temperature targets.
-The extent of anthropogenic emissions reduction depends on what how much help (or push-back) we get from the physical and biological Earth systems.
+The higher carbon emissions associated with warming-driven permafrost thaw may be offset by increases in primary productivity.
+Some studies based on meteorological tower measurements of carbon flux and optical remote sensing suggest that high-latitude ecosystems (mainly tundra and boreal forest) remain carbon-neutral or are even a small net carbon sink [@mcguire_2012_assessment; @welp_2016_increasing].
+However, other studies, based on different sets of flux measurements and airborne gas sampling, suggest that the high-latitude regions are a net carbon source [@belshe_2013_tundra; @commane_2017_carbon; @natali_2019_large].
+The uncertainty in the net high-latitude carbon flux may be driven by the heterogeneity of high-latitude landscapes in terms of vegetation cover, soil properties, topography, and many other factors known to affect both the rate of permafrost thaw and the subsequent carbon flux [@turetsky_2002_boreal; @wickland_2006_effects; @lund_2010_variability; @james_2013_multidecadal; @johnson_2013_permafrost; @grant_2019_modeling1; @grant_2019_modeling2].
+This uncertainty is also present in recent permafrost modeling studies [@burke_2017_quantifying; @qian_2010_enhanced; @ito_2016_impacts; @harp_2016_effect].
 
-Permafrost is an important C reservoir.
-The total northern soil C pool is estimated at 1672 PgC, of which approximately 1466 Pg (88%) is in permafrost [@tarnocai_2009_soil].
-More recent estimates suggest that total high-latitude soil organic matter pool is 1300 Pg, with roughly 800 Pg in permafrost [@hugelius_2014_estimated].
-This frozen C can be released into the atmosphere by several processes related to warming, including aerobic respiration in thawed soil (REF) and anaeribic respiration in thermokarst lakes and wetlands [@turetsky_2002_boreal; @wickland_2006_effects].
-But emissions from wet soils may be offset by high organic matter accumulation rates [@camill_2001_changes].
-Permafrost thaw in boreal peatlands in north-central Saskatchewan increased CO2 and CH4 fluxes from soil to atmosphere by 1.6 and 30 times, respectively [@turetsky_2002_boreal].
-These impacts are exacerbated by the fact that the Arctic is warming roughly 2.5 (TODO ???) times faster than the global average (TODO REF).
-Projections of permafrost C emissions vary.
-[@schuur_2009_effect] estimate 0.8 - 1.1 Pg C yr-1.
-Back-of-the-envelope estimates from [@zimov_2006_climate_change]:
-10-40 g C m-3 day-1 off the bat, slowing down to equilibrium (?) rate of 0.5-5 g C m-3 day-1 for several years.
-
-Rapid methane release from arctic permafrost could potentially cause trillions in economic damage [@whiteman_2013_vast].
-
-On the other hand, warming and CO2 fertilization may increase vegetation productivity, which could increase soil C storage through enhanced litterfall; the balance of these two processes is uncertain [@jones_2005_global].
-There are additional uncertainties associated with vegetation composition (spatially variable peat accumulation) [@camill_2001_changes].
-Other large uncertainties related to soil properties and climate model structure [@harp_2016_effect].
-Several modeling studies generally predict increases in soil C sequestration at high latitudes [@burke_2017_quantifying; @qian_2010_enhanced; @ito_2016_impacts].
-However, this increase is dampened or even reversed when C emissions from permafrost thaw are included [@schaefer_2011_amount; @burke_2017_quantifying].
-
-The importance of Arctic climate feedbacks has led to efforts to incorporate these processes---particularly, permafrost thaw---into land surface models.
-Early efforts derived permafrost extent and C based purely on soil temperature [@lawrence_2005_projection].
-(More recent models have done this better...)
-@schaefer_2011_amount -- SiBCASA estimates 190 +/- 64 Gt C, but this does not include warming feedback or discontinuous permafrost regions.
-@harp_2016_effect -- Community Earth System Model (CESM) simulations over 100 years.
-[@burke_2017_quantifying] -- JULES and ORCHIDEE, with new permafrost scheme, combined with intermediate complexity climate-ocean model (IMOGEN), to look at climate sensitivity to permafrost C emissions.
-Models are highly sensitive to representation of soil processes, which can be more important than differences in scenario and/or climate drivers [@burke_2017_quantifying].
-But @harp_2016_effect argue that model structural uncertainty is larger than soil property uncertainty.
-
-Land surface models are expensive to run, making it challening to use them for uncertainty quantification and exploration of alternative policy scenarios.
+Land surface models are expensive to run, making it challenging to use them for uncertainty quantification and exploration of alternative policy scenarios.
 Simple climate models are an alternative.
 (More on simple climate models).
+(More on permafrost in simple climate models).
 
-(Previous attempts to incorporate permafrost into simple climate models.)
-
-Hector [@hartin_2015_simple].
-In this study, we use Hector to evaluate the global sensitivty of climate to terrestrial biosphere processes, and how this uncertainty is affected by permafrost emissions.
-The current version of Hector does not have an explicit representation of permafrost C emissions.
-In this study, we investigate whether the additional complexity and parametric and structural uncertainty of an explicit representation of permafrost may be warranted in Hector.
-To do this, we evaluate the sensitivity of climate variables (as predicted by Hector) to several different exogenous scenarios of permafrost C emissions.
-(TODO Modify this to be about economic impact)
-
-Some more relevant references:
-- [@zimov_2006_climate_change]
-- [@treat_2013_permafrost_carbon_bomb]
-- [@schuur_2015_climate; @schuur_2011_climate_change; @schaefer_2014_impact; @schaefer_2011_amount; @lee_2011_rate; @kuhry_2010_potential; @kessler_2017_estimating; @hope_2015_economic; @drake_2015_ancient; @burke_2017_quantifying]
+Hector [@hartin_2015_simple; @vega-westhoff_2019_impacts].
+In this study, we implement a simple representation of permafrost thaw and evaluate its consequences for global climate in Hector.
 

### 2020-02-06 analysis/paper/paper.Rmd
Remove deprecated text about parameter calibration

@@
 
-We used the `BayesianTools` R package [@R-BayesianTools] for all parameter calibration.
-The outputs of these calibrations are joint posterior distributions of parameters and their covariances, from which we sample for the sensitivity analysis.
-
-For global parameters, we used the following likelihood:
-
-$$ \log(L) = \sum_s Normal(Hector(\beta, Q_{10}, s) | CMIP5(s), \sigma) $$
-
-where $s$ is one of the four representative carbon pathways (RCPs), $CMIP5(s)$ are the CMIP5 global mean outputs for the corresponding variables, and $\sigma$ is the model error (estimated during the fit).
-We also used the resulting distributions for $\beta$ and $Q_{10}$ for the non-permafrost biome in cases 2 and 3.
-We feel this is appropriate because the CMIP5 models against which these parameters are calibrated do not include permafrost C feedbacks.
-
-For case 2, we calibrated the permafrost-specific $\beta$ and $Q_{10}$ against various literature sources, including:
-
 - Land surface model simulations [@hope_2015_economic; @schaefer_2011_amount; @burke_2017_quantifying].

### 2020-02-06 analysis/paper/paper.Rmd
Move methods text around.

@@
 
-where $a$, $b$, and $c$ are model parameters.
+where $a$, $b$, and $c$ are model parameters whose values are described in the _Configuration_ section below.
 The change in frozen fraction at a given time step, $\Delta \Phi[t]$, is given by:
@@
 
-For globally-averaged permafrost, $a = 2.371$, $b = -0.676$, and $c = -3.685$ to most closely reproduce the rate of 0.172 K$^{-1}$ reported by [Kessler (2015)][kessler] over the range of 0.8 [^2] to 4 K above the pre-industrial baseline.
+[kessler]: https://econpapers.repec.org/RePEc:lsg:lsgwps:wp219
+
+## Methane emissions
+
+In the default configuration of Hector, natural CH~4~ emissions are prescribed at a constant value (300 Tg year^-1^) [@hartin_2015_simple].
+In the new configuration, CH~4~ emissions are a fixed fraction of heterotrophic respiration.
+
+Total heterotrophic respiration ($RH[i,t]$) for biome $i$ at time $t$ is the sum of heterotrophic respiration in detritus ($RH_d$) and soil ($RH_s$).
+
+$$ RH[i,t] = RH_s[i,t] + RH_d[i,t] $$
+
+Detritus and soil heterotrophic respiration are proportional to the sizes of their respective carbon pools ($C_d$ and $C_s$, respectively, both in Pg C),
+with a rate that increases exponentially with temperature according to a biome-specific temperature sensitivity parameter ($Q_{10}$).
+Detritus respiration increases with biome-specific air temperature change since pre-industrial ($T[i,t]$),
+while soil respiraiton increases with the 200 year running mean of air temperature ($T_{200}[i,t]$) because soils warm slowly relative to the atmosphere.
+
+\[ T_{air}[i,t] = \delta \, T_{air}[t] \]
+
+\[ RH_d[i,t] = \frac{1}{4} C_d Q_{10}[i] ^ {T}[i,t] / 10} \]
+
+\[ RH_s[i,t] = \frac{1}{50} C_s Q_{10}[i] ^ {T_{200}[i,t] / 10} \]
+
+In the original formulation, all of the total heterotrophic respiration was assumed to be CO~2~ and natural CH~4~ emissions were entirely separate.
+In the revised version here, total heterotrophic respiration is divided into CO~2~ ($RH_{CO_2}[i,t]$) and CH~4~ ($RH_{CH_4}[i,t]$) based on the biome-specific CH~4~ fraction parameter ($f_{CH_4}[i]$):
+
+\[ RH_{CH_4}[i,t] = f_{CH_4}[i] RH[i,t] \]
+
+\[ RH_{CO_2}[i,t] = (1 - f_{CH_4}[i]) RH[i,t] \]
+
+## Configuration
+
+Initial permafrost C is set to 1035 Pg C based on the estimate for near-surface (<3 m depth) permafrost by @hugelius_2014_estimated.
+
+```{r params}
+# Permafrost exponential model
+pfa <- 2.371
+pfb <- -0.676
+pfc <- -3.685
+
+# Kessler linear model
+kessler_m <- 0.172
+kessler_b <- 0.8 * kessler_m + 1
+```
+
+For globally-averaged permafrost, exponential parameters $a = `r pfa`$, $b = `r pfb`$, and $c = `r pfc`$ to most closely reproduce the rate of `r kessler_m` K^-1^ reported by [Kessler (2015)][kessler] over the range of 0.8 [^2] to 4 K above the pre-industrial baseline.
 
@@
 )
-pfa <- 2.371
-pfb <- -0.676
-pfc <- -3.685
-b <- 4.078
-kessler_m <- 0.172
-kessler_b <- 0.8 * kessler_m + 1
 plot(0, 0, type = "n", xlim = c(0, 8), ylim = c(0, 1),
@@
 ```
-
-[kessler]: https://econpapers.repec.org/RePEc:lsg:lsgwps:wp219
-
-## Configuration
-
-Initial permafrost C is set to 1035 Pg C.
 

### 2020-02-06 analysis/scratch/methane-calibrate.R
Add scratch script for methane calibration

@@ -1,0 +1,52 @@
+devtools::load_all("~/Projects/hector_project/hector-permafrost-submodel")
+library(tidyverse)
+
+outvars_ch4 <- c(RF_CH4(), ATMOSPHERIC_CH4())
+outvars_co2 <- c(ATMOSPHERIC_C(), ATMOSPHERIC_CO2())
+outvars_both <- c(outvars_ch4, outvars_co2, GLOBAL_TEMP())
+yrs <- 1750:2300
+rcp45 <- system.file(
+  "input", "hector_rcp45.ini",
+  package = "hector"
+)
+hc <- newcore(rcp45)
+invisible(run(hc))
+results_default <- fetchvars(hc, yrs, outvars_both, scenario = "default")
+setvar(hc, NA, RH_CH4_FRAC(), 0.004, NA)
+setvar(hc, NA, NATURAL_CH4(), 0, "Tg CH4 year-1")
+invisible(reset(hc))
+invisible(run(hc))
+results_methane <- fetchvars(hc, yrs, outvars_both, scenario = "methane")
+results_both <- as_tibble(bind_rows(results_default, results_methane))
+ggplot(results_both) +
+  aes(x = year, y = value, color = scenario) +
+  geom_line() +
+  geom_vline(xintercept = 2100, linetype = "dashed") +
+  facet_wrap(vars(variable), scales = "free")
+
+ggsave("analysis/figures/dynamic-methane.png")
+
+# Put it together with the permafrost model
+outvars_both2 <- c(outvars_ch4, outvars_co2, GLOBAL_TEMP(),
+                  ATMOSPHERIC_CO2(), GLOBAL_TEMP(), F_FROZEN(), SOIL_C())
+yrs <- 1750:2300
+hc <- newcore(rcp45)
+invisible(run(hc))
+results_nopf <- fetchvars(hc, yrs, outvars_both2, scenario = "default")
+setvar(hc, NA, RH_CH4_FRAC(), 0.004, NA)
+setvar(hc, NA, NATURAL_CH4(), 0, "Tg CH4 year-1")
+invisible(reset(hc))
+setvar(hc, 0, PERMAFROST_C(), 1035, "PgC")
+invisible(run(hc))
+results_pf <- fetchvars(hc, yrs, outvars_both2, scenario = "permafrost")
+results_allpf <- bind_rows(results_nopf, results_pf) %>%
+  as_tibble()
+
+ggplot(results_allpf) +
[…]

### 2020-02-06 analysis/paper/paper.Rmd
Revise permafrost methods for new Hector changes
- Use lognormal CDF
- RH{CH4} is on top of regular RH, not a trade-off

@@
 output: 
+    bookdown::pdf_document2:
     bookdown::html_document2:
@@
 [Kessler (2015)][kessler] assume this relationship is linear.
-However, because the permafrost area fraction is, by definition, bounded by zero (and 1 [^1]), and because deeper permafrost thaws more slowly than shallow permafrost, we use a negative logistic relationship instead:
+However, because the permafrost area fraction is, by definition, bounded by zero (and 1 [^1]), and because deeper permafrost thaws more slowly than shallow permafrost, we use the lognormal cumulative distribution function ($NCDF(\log(x) | \mu, \sigma)$) instead:
 
@@
 \[
-\Phi[t] = 1 - \left( 1 + a \exp\left( b * T_{air}[t] \right) \right) ^ c
+\Phi[t] = 1 - NCDF(\log(x) | \mu, \sigma)
 \]
 
-where $a$, $b$, and $c$ are model parameters whose values are described in the _Configuration_ section below.
+where $\mu$ and $\sigma$ are model parameters whose values are described in the _Configuration_ section below.
 The change in frozen fraction at a given time step, $\Delta \Phi[t]$, is given by:
@@
 \[
-\Delta \Phi[t] = \max(\Phi[t] - \Phi[t-1], 0)
+\Delta \Phi[t] = \Phi[t] - \Phi[t-1]
 \]
-
-In other words, permafrost thaw is permanent -- once it thaws, it does not re-freeze, even if the temperature drops.
 
@@
 
-\[ RH_d[i,t] = \frac{1}{4} C_d Q_{10}[i] ^ {T}[i,t] / 10} \]
+\[ RH_d[i,t] = \frac{1}{4} C_d Q_{10}[i] ^ {{T}[i,t] / 10} \]
 
@@
 
-In the original formulation, all of the total heterotrophic respiration was assumed to be CO~2~ and natural CH~4~ emissions were entirely separate.
-In the revised version here, total heterotrophic respiration is divided into CO~2~ ($RH_{CO_2}[i,t]$) and CH~4~ ($RH_{CH_4}[i,t]$) based on the biome-specific CH~4~ fraction parameter ($f_{CH_4}[i]$):
+In the original formulation, natural CH~4~ emissions were a prescribed constant value.
+In the revised version here, natural CH~4~ emissions are a fraction of heterotrophic respiration ($RH_{CH_4}[i,t]$) based on the biome-specific CH~4~ fraction parameter ($f_{CH_4}[i]$):
 
 \[ RH_{CH_4}[i,t] = f_{CH_4}[i] RH[i,t] \]
-
-\[ RH_{CO_2}[i,t] = (1 - f_{CH_4}[i]) RH[i,t] \]
 

### 2020-02-06 analysis/scripts/calibrate-permafrost.R

@@ -1,0 +1,17 @@
+kessler_m <- 0.172
+kessler_b <- 0.8 * kessler_m + 1
+kessler_line <- function(x) pmin(kessler_b - kessler_m * x, 1)
+cdfun <- function(x, mu, sig) plnorm(x, mu, sig, lower.tail = FALSE)
+pfit <- function(pars) {
+  mu <- pars[1]
+  sig <- pars[2]
+  x <- seq(0.01, 6, 0.1)
+  yline <- kessler_line(x)
+  yfun <- cdfun(x, mu, sig)
+  ## sum((yline - yfun) ^ 2 * (1 / x^1.8))
+  sum((yline - yfun) ^ 2)
+}
+fit <- optim(c(3.5, 2), pfit)
+curve(kessler_line(x), 0, 6, col = "blue", ylim = c(0, 1))
+curve(cdfun(x, fit$par[1], fit$par[2]), 0, 6, add = TRUE)
+abline(h = c(0, 1), lty = 2)

### 2020-02-06 analysis/scratch/methane-calibrate.R
Tweak methane calibration script

@@ -18,17 +18,19 @@ invisible(reset(hc))
 invisible(run(hc))
 results_methane <- fetchvars(hc, yrs, outvars_both, scenario = "methane")
 results_both <- as_tibble(bind_rows(results_default, results_methane))
-ggplot(results_both) +
+p1 <- ggplot(results_both) +
   aes(x = year, y = value, color = scenario) +
   geom_line() +
   geom_vline(xintercept = 2100, linetype = "dashed") +
   facet_wrap(vars(variable), scales = "free")
 
-ggsave("analysis/figures/dynamic-methane.png")
+## ggsave("analysis/figures/dynamic-methane.png")
 
 # Put it together with the permafrost model
 outvars_both2 <- c(outvars_ch4, outvars_co2, GLOBAL_TEMP(),
-                  ATMOSPHERIC_CO2(), GLOBAL_TEMP(), F_FROZEN(), SOIL_C())
+                   ATMOSPHERIC_CO2(), GLOBAL_TEMP(), F_FROZEN(), SOIL_C(),
+                   RH(), RH_SOIL(), RH_DETRITUS(),
+                   VEG_C(), DETRITUS_C(), SOIL_C())
 yrs <- 1750:2300
 hc <- newcore(rcp45)
 invisible(run(hc))
@@ -41,12 +43,13 @@ invisible(run(hc))
 results_pf <- fetchvars(hc, yrs, outvars_both2, scenario = "permafrost")
 results_allpf <- bind_rows(results_nopf, results_pf) %>%
   as_tibble()
-
-ggplot(results_allpf) +
+p2 <- ggplot(results_allpf) +
   aes(x = year, y = value, color = scenario) +
   geom_line() +
-  geom_vline(xintercept = 2100, linetype = "dashed") +
+  ## geom_vline(xintercept = 2100, linetype = "dashed") +
   facet_wrap(vars(variable), scales = "free") +
   ggtitle("RCP 4.5")
+## p2 %+% filter(results_allpf, year < 1900)
+p2
 
-ggsave("analysis/figures/pf-methane-rcp45.png")
+## ggsave("analysis/figures/pf-methane-rcp45.png")

### 2020-03-17 analysis/paper/paper.Rmd
Remove old appendices

@@
 
-# Appendix
-
-```{r fig-global-dist, fig.cap = global_dist_cap}
-global_dist_cap <- paste(
-  htmlfig(),
-  "Input parameter distributions for global Hector."
-)
-fig("global_draws.png")
-```
-
-`r mdcap(global_dist_cap)`
-
-```{r fig-biome-dist, fig.cap = biome_dist_cap, fig.dim = c(9, 9)}
-biome_dist_cap <- paste(
-  htmlfig(),
-  "Input parameter distributions for Hector with biomes."
-)
-fig("biome_draws.png")
-```
-
-`r mdcap(biome_dist_cap)`
-
-```{r fig-global-timeseries, fig.cap = global_ts_cap, fig.dim = c(10, 8)}
-global_ts_cap <- paste(
-  htmlfig(),
-  "Time series of Hector outputs for different parameter combinations."
-)
-readd(ts_global_sims)
-```
-
-`r mdcap(global_ts_cap)`
-
-```{r fig-biome-timeseries, fig.cap = biome_ts_cap, fig.dim = c(10, 8)}
-biome_ts_cap <- paste(
-  htmlfig(),
-  "Overall parameter uncertainty of Hector simulations with permafrost biome."
-)
-readd(ts_biome_sims)
-```
-
-`r mdcap(biome_ts_cap)`
-
 ### Colophon

### 2020-03-17 analysis/paper/paper.Rmd
Use pkgload::load_all instead of devtools

@@
 library(ggplot2)
-library(hector.permafrost.emit)
 
@@
 
-devtools::load_all("~/Projects/hector_project/hector-permafrost-submodel")
+pkgload::load_all(here())
+pkgload::load_all("~/Projects/hector_project/hector-permafrost-submodel")
 

### 2020-03-17 analysis/paper/paper.Rmd
Remove deprecated functions from paper

@@
 ```{r setup, echo = FALSE}
-library(drake)
 library(knitr)
@@
 
-# Dummy target to force-rebuild this when figures change
-invisible(readd(figs_dir_t))
-
 fig <- function(...) {
@@
   knitr::include_graphics(file_in(filepath))
-}
-
-n <- 0
-mdcap <- function(text) {
-  n <<- n + 1
-  out <- sprintf("**Figure %d**: %s", n, text)
-  if (grepl("gfm", knitr::opts_knit$get("rmarkdown.pandoc.to"), ignore.case = TRUE)) {
-    out
-  } else {
-    invisible(NULL)
-  }
-}
-
-m <- 0
-htmlfig <- function() {
-  m <<- m + 1
-  out <- paste0("**Figure ", m, "**:")
-  if (grepl("html", knitr::opts_knit$get("rmarkdown.pandoc.to"), ignore.case = TRUE)) {
-    out
-  } else {
-    invisible(NULL)
-  }
 }

### 2020-03-17 README.Rmd
Update README with renv instructions

@@
 ---
-output: github_document
+output:
+  github_document:
+    html_preview: false
 ---
@@
 This repository contains the source code and raw manuscript files for the Hector permafrost emissions paper.
+The paper source code is located in `analysis/paper/paper.Rmd`.
 
-## Installation
+## `renv`
 
-You can install the development version of `hector.permafrost.emit` from GitHub with:
-
-```r
-# install.packages("devtools")
-devtools::install_github("ashiklom/hector.permafrost.emit")
-```
-
-## Data description
-
-- `analysis/data/raw_data/schaefer_teb_rpc_(min/max/mean).csv` -- Raw, digitized versions of Figure 5c (Ensemble mean and uncertainty of permafrost C respiration) in [@schaefer_2011_amount].
-Original figure is `analysis/data/raw_data/schaefer_teb_timeseries.gif`.
-These results were digitized using WebPlotDigitizer (https://apps.automeris.io/wpd/), and interpolated to annual resolution (using R `stats::spline` function).
-    - [@schaefer_2011_amount] assume only CO2 emissions. So we add this as an extra anthropogenic CO2 flux.
-- `analysis/data/raw_data/hope_2016_Mt{CO2/CH4}_{lo/mean/hi}.csv` -- Raw, digitized versions of Figures 1 (permafrost CO2 emissions from SiBCASA) and 2 (same for CH4) from [@hope_2015_economic].
-Same digitization and interpolation procedure.
-Original figures are `analysis/data/raw_data/hope_2016_nclimate_{co2,ch4}.jpg`.
+This repository uses `renv` to manage R package dependencies in a reproducible way.
+To download and install all packages in the correct versions for this analysis, launch R from the project root directory and run `renv::restore()`.

### 2020-03-17 analysis/paper/paper.Rmd
Add missing packages for paper.Rmd & other fixes

@@
 library(ggplot2)
+library(hector)
 
@@
 pkgload::load_all(here())
-pkgload::load_all("~/Projects/hector_project/hector-permafrost-submodel")
 
@@
 run_rcp <- function(rcp) {
-  inidir <- file.path("~", "Projects", "hector_project",
-                      "hector-permafrost-submodel", "inst", "input")
+  inidir <- system.file("input", package = "hector")
   inifile <- normalizePath(file.path(inidir, paste0("hector_rcp", rcp, ".ini")))
